---
title: "Fig 2"
---
```{r}
library(tidyverse)
library(cowplot)
library(ComplexHeatmap)
library(circlize)
library(ggpubr)
library(ggrepel)
library(speckle)

domcol <- c( "gray13", "gold2", "plum4", "darkorange1", "lightskyblue2", "maroon", "burlywood3")
names(domcol)<-1:7
classcols = c("Eutopic endometrium" = "#c2a5cf", 
              "Peritoneal endometriosis" = "#d9f0d3")
cellcols<-c("#8DB600","#fdae61","#a95f96","#0067A5","#604E97",
            "#f46d43","hotpink","#BE0032","cornflowerblue")
names(cellcols)<-c("B cells","Plasma cells","CD4 T cells","CD8 T cells",
                   "NK cells","Myeloid cells","unknown","Epithelial","Stroma")
dominantcols<-c(Epithelial="#BE0032",Stroma="cornflowerblue",Immune="forestgreen")
files<-list.files()
files<-files[str_detect(files,"bgclus")]
files<-files[-1]
metadata<-lapply(files,data.table::fread)%>%bind_rows()

i<-metadata%>%distinct(Run_Tissue_name,dom_nEndoh1_75)%>%
  count(dom_nEndoh1_75)%>%filter(n==1)%>%
  select(dom_nEndoh1_75)%>%unlist()
metadata<-metadata%>%filter(dom_nEndoh1_75!=i)
#metadata<-metadata%>%mutate(cl_corrected=ifelse(cl_corrected=="",cl_type,cl_corrected))
```

# Supplementary heatmap
```{r}
files<-list.files("docs",full.names = T)
files<-files[str_detect(files,"metag")]
files<-files[str_remove_all(files,"docs/|_metagenes.csv")%in%
               unique(meta$Run_Tissue_name)]
metagenes<-lapply(files,read_csv)
names(metagenes)<-str_remove_all(files,"docs/|_metagenes.csv")
metagenes<-lapply(names(metagenes),function(x) 
  {tab<-metagenes[[x]];
  tab<-tab%>%column_to_rownames("probe");
  colnames(tab)<-paste0(x,"part_",colnames(tab));
  tab})%>%bind_cols()
zscores<-t(scale(t(metagenes)))

temp<-meta%>%
  count(protodomain,merged_corrected)%>%
  group_by(protodomain)%>%
  mutate(total=sum(n),prop=100*n/total)%>%
  select(protodomain,merged_corrected,prop)%>%
  pivot_wider(names_from = merged_corrected,values_from = prop)
temp<-temp%>%mutate(across(-1,function(x) replace_na(x,0)))%>%
  column_to_rownames("protodomain")
temp<-t(temp)[,colnames(zscores)]

annot<-metadata%>%distinct(protodomain,Run_Tissue_name,Class,dom_nEndoh1_75)
annot<-annot[order(match(annot$protodomain,colnames(temp))),]
nobatch<-sva::ComBat(dat = metagenes,batch = annor$Run_Tissue_name,par.prior = T)
cor_dist<-cor(nobatch)
cl<-hclust(as.dist(1-cor_dist),method="ward.D2")
hm1<-Heatmap(temp,show_column_names = F,
        top_annotation = HeatmapAnnotation(Slide=annot$Run_Tissue_name,
                                           Domain=annot$dom_nEndoh1_75,
                                           Class=annot$Class,
                                        col=list(Class=classcols,
                                                 Slide=slidecol,
                                                 Domain=domcol)),
       cluster_rows = F,cluster_columns = cl,use_raster = F,name="%",
       column_dend_height = unit(4, "cm"),
       col=colorRamp2(c(100,50,25,12,6,0), c("#14081D","#651B79","#de7065ff","#FFC28B","#FCFDBF","white")))
hm2<-Heatmap(zscores,show_column_names = F,show_row_dend = F,
             use_raster = F,name="zscore",cluster_columns = cl)

pdf("suppHM.pdf",width = 11,height = 8)
 hm1%v%hm2
dev.off()
```

# Composition heatmap
```{r}
#discard single slide domains & qc failing cells
#i<-metadata%>%distinct(dom_nEndoh1_75,Run_Tissue_name)%>%
#  add_count(dom_nEndoh1_75)%>%filter(n==1)%>%select(dom_nEndoh1_75)%>%unlist()
#metadata<-metadata%>%filter(!dom_nEndoh1_75%in%i)

compo<-metadata%>%
  count(dom_nEndoh1_75,cl_corrected)%>%
  group_by(dom_nEndoh1_75)%>%
  mutate(total=sum(n),prop=100*n/total)
compo<-compo%>%
  mutate(dominant=ifelse(str_detect(cl_corrected,"B|CD|Myel|Plas|NK"),"Immune",
                cl_corrected))%>%
  group_by(dom_nEndoh1_75,dominant)%>%summarise(prop=sum(prop))%>%
  group_by(dom_nEndoh1_75)%>%slice_max(prop)%>%
  select(-prop)%>%right_join(compo)
compo<-compo%>%distinct(dom_nEndoh1_75,total)%>%
  arrange(desc(total))%>%rownames_to_column("index")%>%
  select(-total)%>%right_join(compo)

pA<-compo%>%filter(cl_corrected!="unknown")%>%
    ggplot(aes(x=cl_corrected,y=as.factor(index),fill=prop))+
    geom_tile()+scale_fill_viridis_c(option = "F",direction = -1,trans="sqrt")+
    theme_minimal()+labs(x="Cell type",y="Domain",fill="%")+
    theme(axis.text.x = element_text(angle=45,vjust = 1,hjust = 1))
pA1<-compo%>%ggplot(aes(x=1,y=index,fill=as.factor(dom_nEndoh1_75)))+
    geom_bar(stat="identity",position = "fill")+
    scale_fill_manual(values = domcol)+theme_void()+
    theme(legend.position = "n")
pA2<-compo%>%ggplot(aes(x=1,y=index,fill=dominant))+
  geom_bar(stat="identity",position = "fill")+
  scale_fill_manual(values = dominantcols)+theme_void()+
    theme(legend.position = "bottom",legend.direction = "vertical")
pA3<-compo%>%distinct(index,total)%>%
  ggplot(aes(x=total,y=index))+geom_col()+
  theme_classic()+scale_x_sqrt()+
  theme(axis.ticks.y = element_blank(),axis.text.y = element_blank(),
        plot.margin=grid::unit(c(0,3,0,-6), "mm"),
        axis.text.x = element_text(angle=45,vjust=1,hjust=1))+
  labs(y="")

compo<-metadata%>%count(dom_nEndoh1_75,Class,name = "class_n")%>%right_join(compo)
compo<-compo%>%distinct(index,dom_nEndoh1_75,Class,class_n)%>%group_by(Class)%>%summarise(class_n=sum(class_n),index="Total")%>%bind_rows(compo)
pA4<-compo%>%#mutate(dom=factor(dom,levels=c(6:1,"Total")))%>%
    ggplot(aes(x=class_n,y=index,fill=Class))+
    geom_bar(stat = "identity",position = "fill")+
    theme_classic()+labs(x="Cells",y="")+
    theme(axis.ticks.y = element_blank(),axis.text.y = element_blank(),
          axis.line.y = element_blank(),plot.margin=grid::unit(c(0,3,0,-6), "mm"),
          legend.position = "bottom",legend.direction = "vertical")+
    scale_fill_manual(values = classcols)
plot_grid(pA1,pA2,pA+theme(legend.position = "top"),pA3,pA4,ncol=5,rel_widths = c(.05,.05,.4,.3,.2),align = "h",axis = "tb")


```


# Domain signature 
```{r}
temp<-meta%>%distinct(protodomain,dom_nEndoh1_75)
temp<-temp%>%split(temp$dom_nEndoh1_75)%>%sapply(function(x) 
  rowMeans(metagenes[,x$protodomain]))
temp<-t(scale(t(temp)))
temp<-temp%>%as.data.frame()%>%rownames_to_column("probe")%>%
  pivot_longer(-1,names_to = "domain",values_to = "zscore")
temp<-temp%>%filter(str_detect(probe,"Chann",negate = T))

pB<-temp%>%ggplot(aes(x=domain,y=zscore))+geom_violin()+
  geom_jitter(width=0.1)+geom_text_repel(data=temp%>%group_by(domain)%>%
                                           slice_max(zscore,n=3),
          aes(x=domain,y=zscore,label=probe),nudge_x = .2)+theme_classic()

save(hm,lgd,lgd2,pA,pA0,pB,file = "docs/Fig2.RData")
#plot_grid(pA0,pA,pB,ncol=3,align = 'h',axis = 'tb',rel_widths = c(.2,.2,.6))

```
## complex signature
```{r}
temp<-meta%>%distinct(protodomain,dom_nEndoh1_75,merged_corrected)%>%mutate(gr=paste0(dom_nEndoh1_75,":",merged_corrected))
temp<-temp%>%split(temp$gr)%>%sapply(function(x) 
  rowMeans(metagenes[,x$protodomain]))
temp<-temp[str_detect(rownames(temp),"Chann",negate=T),]

temp1<-as.data.frame(cbind(name=colnames(temp),dom=str_remove(colnames(temp),":.+"),celltype=str_remove(colnames(temp),".+:")))
nobatch<-sva::ComBat(temp,batch = temp1$dom,par.prior = T)
Heatmap(t(scale(t(nobatch))),name="z",column_split = str_remove(colnames(temp),".+:"),column_labels = str_remove(colnames(temp),":.+"))

```
# Example space plots: squidpy
#   Cell composition contrast
```{r}
dombars<-lapply(names(classcols),function(x)
    meta%>%filter(Class==x)%>%count(Class,dom_nEndoh1_75)%>%
        ggplot(aes(y=Class,x=n,fill=as.factor(dom_nEndoh1_75)))+
        geom_bar(stat="identity",position="fill")+
        theme_void(base_size = 14)+scale_fill_manual(values=domcol)+
      labs(x="%",y="",fill="domain")+theme(legend.position="none")+
      ggtitle(x))
cellbars<-lapply(names(classcols),function(x)
    meta%>%filter(Class==x)%>%count(Class,merged_corrected)%>%
        ggplot(aes(y=Class,x=n,fill=merged_corrected))+
        geom_bar(stat="identity",position="fill")+
        theme_void()+scale_fill_manual(values=cellcols)+
      labs(x="%",y="",fill="domain")+theme(legend.position="none"))

temp<-meta%>%count(Class,dom_nEndoh1_75,merged_corrected)
dombars1<-lapply(names(classcols),function(x)
    temp%>%filter(Class==x)%>%ggplot(aes(x=n,y=as.factor(dom_nEndoh1_75),fill=merged_corrected))+
        geom_bar(stat="identity",position="fill")+
        theme_classic(base_size = 14)+scale_fill_manual(values=cellcols)+
        labs(x="%",y="",fill="domain")+
        theme(legend.position="none",axis.line = element_blank(),
              axis.ticks.y = element_blank()))
names(dombars1)<-names(dombars)<-names(cellbars)<-names(classcols)
pC<-plot_grid(dombars$`Eutopic endometrium`,cellbars$`Eutopic endometrium`,
              dombars1$`Eutopic endometrium`,nrow=3,rel_heights = c(.2,.1,.7),
              align="v",axis = "lr")
pC1<-plot_grid(dombars$`Peritoneal endometriosis`,cellbars$`Peritoneal endometriosis`,
               dombars1$`Peritoneal endometriosis`,nrow=3,rel_heights = c(.2,.1,.8),
               align="v",axis = "lr")

#test
temp<-split(meta,f = meta$dom_nEndoh1_75)
props<-lapply(temp,function(x) 
  getTransformedProps(x$merged_corrected,
                      x$Run_Tissue_name))
temp<-lapply(temp,function(x)
  x%>%distinct(Class,Run_Tissue_name)%>%
      mutate(Class=str_replace(Class,".+ e","E")))

designM<-lapply(temp,function(x) model.matrix(~0+Class,data=x))#groups around 
fit<-lapply(1:6,function(x) lmFit(props[[x]]$TransformedProps,designM[[x]]))
cont<-lapply(designM,function(x)
  makeContrasts(ClassEndometriosis-ClassEndometrium,levels = x))
fit<-lapply(1:6,function(x) contrasts.fit(fit[[x]],cont[[x]]))
fit <- lapply(fit,eBayes, robust=TRUE)

res<-lapply(fit,function(x) topTable(x,number = length(cellcols),coef = 1))
res<-lapply(1:6, function(x)
  res[[x]]%>%as.data.frame()%>%
    rownames_to_column("cell_type")%>%
    mutate(dom=x))%>%bind_rows()
```
# Neighbors
```{r}
files<-list.files("docs",full.names = T)
files<-files[str_detect(files,"delau")]

edges<-lapply(files, read_csv)
names(edges)<-str_remove_all(files,"docs/|delaunay.+")
edges<-lapply(names(edges), function(x) 
  edges[[x]]%>%mutate(Run_Tissue_name=x))%>%bind_rows()
edges<-edges%>%filter(distance<=0.1)
edges<-edges%>%mutate(from=paste0(Run_Tissue_name,from),to=paste0(Run_Tissue_name,to))
edges<-meta%>%
    select(cell_ID,merged_corrected,dom_nEndoh1_75)%>%
    right_join(edges,by = c("cell_ID"="from"))
colnames(edges)[1:3]<-c("from","type_from","dom_from")
edges<-meta%>%
    select(cell_ID,merged_corrected,dom_nEndoh1_75)%>%
    right_join(edges,by = c("cell_ID"="to"))
colnames(edges)[1:3]<-c("to","type_to","dom_to")
edges<-meta%>%distinct(Run_Tissue_name,Class)%>%right_join(edges)
edges<-edges%>%mutate(homodom=dom_to==dom_from,
                      homotypic=type_from==type_to)

edges<-edges%>%filter(homodom==TRUE)
temp<-edges%>%count(dom_to,type_from,type_to,homotypic)%>%
  filter(type_from!="",type_to!="")
temp<-temp%>%group_by(dom_to)%>%mutate(total=sum(n),prop=100*n/total)
temp$edge_type<-apply(temp,1,function(x) paste0(sort(x[2:3]),collapse = " - "))
temp<-temp%>%mutate(name_from=str_remove(edge_type," -.+"),name_to=str_remove(edge_type,".+- "))

pF<-temp%>%group_by(type_from,type_to)%>%slice_max(prop)%>%
    mutate(name_to=factor(name_to,levels = c("Plasma cells","NK cells","Myeloid cells","Mesenchymal cells","Epithelial cells","Endothelial cells","CD8 T cells","CD4 T cells","B cells" )))%>%
    ggplot(aes(y=name_from,x=name_to,col=as.factor(dom_to),size=prop))+
    geom_point()+
    scale_color_manual(values=domcol)+
    scale_size(range = c(1,8),breaks = c(80,40,20,10,5))+
    theme_classic()+
    theme(axis.line.y = element_blank(),
          axis.ticks.y = element_blank())+
    labs(x="",y="")+guides(col="none")
temp<-temp%>%group_by(dom_to)%>%mutate(total=sum(n),prop=100*n/total)
pD1<-temp%>%group_by(dom_to)%>%arrange(prop)%>%mutate(i=1:81)%>%
  ggplot(aes(x=i,y=prop,col=factor(dom_to),group = dom_to))+
  geom_point(aes(size=prop))+geom_line()+
  scale_color_manual(values=domcol)+theme_classic()+
  scale_x_sqrt()+scale_y_sqrt()+
  theme(axis.text.x = element_blank(),axis.ticks.x = element_blank())+
  labs(x="interacting partners",y="%",size="%")+guides(col="none")

#density
pE<-edges%>%ggplot(aes(x=as.factor(dom_to),y=distance,col=as.factor(dom_to)))+
  geom_violin()+
  geom_boxplot(width=0.1)+scale_y_log10()+
  scale_colour_manual(values=domcol)+theme_minimal()+
  labs(x="",y="Distance between nuclei")+
  stat_compare_means(method = "wilcox.test",label="p.signif",
                     comparisons = list(c(1,6),c(2,6),c(3,6),c(4,6),c(5,6)),
                     alternative="g")+theme(legend.position="none")

#partA<-plot_grid(hm,plot_grid(pB,NULL,nrow = 2,rel_heights = c(.95,.05)))
#partA<-plot_grid(partA,plot_grid(lgd,lgd2,lgd1,ncol=3),nrow=2,rel_heights = c(.8,.2))
#partA<-plot_grid(partA,pC,pC1,ncol=3,rel_widths = c(.6,.2,.2))
#partB<-plot_grid(hm1,pD,NULL,nrow = 3,rel_heights = c(.3,.5,.2))
#partB<-plot_grid(NULL,partB,rel_widths = c(.6,.4))
#pdf("fig2.pdf",width = 21,height = 15)
#plot_grid(partA,partB,nrow = 2,rel_heights = c(.3,.7))
#dev.off()
```
#######################################################################################
# Domain plot
```{r}
#temp<-split(metadata,by="Run_Tissue_name")%>%
#  lapply(function(x) x[sample(1:nrow(x),10000),])
dom6<-metadata%>%filter(dom_nEndoh1_75==6)
temp<-dom6%>%split(by="Run_Tissue_name")


get_expre<-function(sample_name){
  fname<-paste0("docs/",sample_name,"asinh.csv")
  mat<-data.table::fread(fname)
  mat<-mat%>%as.data.frame()%>%column_to_rownames("probe")
  mat<-mat[,temp[[sample_name]]$cell_id]
  colnames(mat)<-paste0(sample_name,colnames(mat))
  return(mat)}
dom_mat<-lapply(names(temp),function(x) get_expre(x))

dom_mat<-do.call(cbind,dom_mat)
temp<-do.call(rbind,temp)%>%
  mutate(colname=paste0(Run_Tissue_name,cell_id))
dom_mat<-dom_mat<-dom_mat[,temp$colname]

#umap
#options(future.globals.maxSize = 30000 * 1024^2,future.seed=TRUE)#!!!
#Do we want seurat style plots?
#seuobj<-CreateSeuratObject(counts=dom_mat,
#                           meta.data = temp,
#                           project = "cosmx",
#                           assay = "protein")
#seuobj <- NormalizeData(seuobj,normalization.method = "LogNormalize",verbose=T)
#seuobj[["protein"]] <- split(seuobj[["protein"]], f = seuobj$Run_Tissue_name)
#seuobj<- FindVariableFeatures(seuobj, assay = "protein",
#                              nfeatures = 67, layer = "counts.1019F1-PT9")#force it to use all protes
#seuobj <- ScaleData(seuobj,features = rownames(seuobj),verbose=F)
#seuobj <- RunPCA(seuobj,features = rownames(seuobj),npcs=30,verbose=F)
#seuobj <- IntegrateLayers(
#  object = seuobj, method = RPCAIntegration,
#  orig.reduction = "pca", new.reduction = "integrated.rpca",
#  normalization.method ="LogNormalize",
#  k.filter=50,#neighbors to use when filtering anchors, default 200
#  k.anchor=20,#neighbors for picking anchors, default: 5, for 120k cells vignette uses 20
#  k.weight=30,#neighbors for weighting anchors, default 100
#  reference=14,#BEME596
#  features = rownames(seuobj),
#  verbose = T)
#seuobj <- RunUMAP(seuobj, reduction = "integrated.rpca",
#                  dims = 1:30, return.model = F, )

nobatch<-ComBat(dom_mat,batch = temp$Run_Tissue_name,par.prior = T)
scaled<-t(scale(t(nobatch)))
temp<-temp%>%split(by="merged_corrected")
scaled<-lapply(temp, function(x) scaled[,x$cell_ID])
scaled_ClassAvg_ct<-lapply(names(temp), function(x) temp[[x]]%>%
                             split(by="Class")%>%sapply(function(y) 
                               rowMeans(scaled[[x]][,y$cell_ID])))
names(scaled_ClassAvg_ct)<-names(scaled)
hms<-lapply(names(scaled_ClassAvg_ct),function(x) 
  Heatmap(scaled_ClassAvg_ct[[x]],name="sample corrected\nscaled\ncell type\navg per Class",
          cluster_columns = F,column_title = x))
names(hms)<-names(scaled)
```

## Menstrual cycle
```{r}
meta<-meta%>%mutate(cycle=case_when(Run_Tissue_name%in%c("BEME541_Endometriosis-795R3","6480D2-PT7")~"Hormone",Run_Tissue_name%in%c("BEME209-Endometrium","BEME550-Endometrium","6318B1-PT20","BEME518_Endometriosis_1023-A2","BEME523-Endometriosis","1019F1-PT9")~"Follicular",Run_Tissue_name%in%c("BEME228-9152C4-PT14","BEME596-Endometrium","BEME142_1044B1-PT4","BEME005-3042A1-PT11","BEME540-Endometriosis","1194A4-PT5" )~"Luteal"))