---
title: "Figure 1"
author: "Pak Yu"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---
# Figure 1 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Library loading 


```{r ,message = FALSE}
library(Seurat)
library(tidyverse)
library(cowplot)
library(broom)
library(readxl)
library(ggplot2)
library(gdata)
library(ggbeeswarm)
library(ggrepel)
library(beeswarm)
library(pheatmap)
library(MAST)
```


```{r}
getwd()

setwd("/home/harom/Code/sc_endo_la_immune")
```


#Immune cells in all the classes
```{r}
#load Immune cells RDS file
Immune.Cells.TrueV2<- readRDS("data/fixed_annotated_aux.seurat.rds")
```




#Getting the Fold change enrichment accross class
```{r}
table.class.cell = table(Immune.Cells.TrueV2@meta.data$Immune_Cluster_ID, Immune.Cells.TrueV2@meta.data$Class)
table.class.cell.perc = table.class.cell/rowSums(table.class.cell)

obs.class.perc = table(Immune.Cells.TrueV2@meta.data$Class)/length(Immune.Cells.TrueV2@meta.data$Class)
table.fc = table.class.cell
table.fc = is.na(table.fc)

for (class in colnames(table.class.cell.perc)) {
        for (cell.type in rownames(table.class.cell.perc)) {
                expected = table.class.cell.perc[cell.type, class]
                obs = obs.class.perc[class]
                if (obs > expected) {
                        ratio = -(obs/expected)
                } else {
                        ratio = expected/obs
                }
                #ratio = expected/obs
                
                table.fc[cell.type, class] = ratio
        }
}


quantile_breaks <- function(xs, n = 10) {
        breaks <- quantile(xs, probs = seq(0, 1, length.out = n))
        breaks[!duplicated(breaks)]
}
mat_breaks <- quantile_breaks(table.fc, n = 11)

table.fc = table.fc[,c("Endometrioma", "Endometriosis", "Eutopic Endometrium", "Unaffected ovary")]


pheatmap(t(table.fc), 
         display_numbers = signif(t(table.fc), 3),
         color             = colorRampPalette(c("#81B8D5", "white", "#E37F65"))(10),
         breaks            = mat_breaks,
         cluster_rows      = F,
         cluster_cols      = F,
         fontsize          = 14)
```


#Fig1A
```{r}
#UMAP of the immune cell subsets

Final.Immune.Cells <- subset(Immune.Cells.TrueV2,cells=which(!is.na(Immune.Cells.TrueV2$Final_Immune_Cluster_ID))) 
  

cluster.cols = c("plum4", "darkorange1", "lightskyblue2", "firebrick", "burlywood3", "gray51", "springgreen4", "lightpink2", "deepskyblue4", "lightsalmon2", "mediumpurple4", "orange", "maroon", "yellow3", "brown4", "yellow4", "sienna4", "chocolate", "gray", "gold2", "black", "#00441b", "#2166ac", "#fddbc7")

p1 <- DimPlot(object = Final.Immune.Cells, pt.size = 0.1, group.by ="Final_Immune_Cluster_ID", cols = cluster.cols,  split.by = "Class", label = F, raster=FALSE) +
  theme_classic(base_size = 20)  ##this is a umap with the clusters grouped with their cell identity

p1

pdf(file = "Plots/Endometriosis_LA/Immune_SubClusters_PotentialFig1A.pdf", width = 20, height = 10) 
p1
dev.off()

```






#Figure 1B
```{r}
#Proportions of Immune Cells groups  by class first reorder the clusters

Immune.Cells.TrueV2@meta.data$Immune_Cluster_ID <- factor(Immune.Cells.TrueV2@meta.data$Immune_Cluster_ID, 
                            levels=c("Activated CD8 T-cells/Early Exhausted CD8 T-cells",
                                     "Activated CD8 T-cells",
                                     "Central Memory CD8 T-cells (1)",
                                     "Central Memory CD8 T-cells (2)", 
                                     "Central Memory CD8 T-cells (3)",
                                     "Cytotoxic CD8 T-cells",
                                     "Effector Memory CD8",
                                     "Exhausted CD8 T-cells",
                                     "Th2 T-cells",
                                     "Th17 T-cells",
                                     "B-cells",
                                     "Plasma cells",
                                     "Plasmablast/Plasma cells",
                                     "Macrophages M1", 
                                     "Macrophage M1/M2",
                                     "Macrophages M2",
                                     "Mast cells",
                                     "NK cells"))

Immune.Cells.TrueV2@meta.data %>% 
      group_by(Immune_Cluster_ID) %>% 
     summarise(n_tissue=n_distinct(orig.ident), n_patient= n_distinct(Patient.No.), n=n()) %>%
      print(n=18)


Final.Immune.Cells@meta.data <- droplevels(Final.Immune.Cells@meta.data)

Immune.Cells.TrueV2@meta.data <- Immune.Cells.TrueV2@meta.data %>% 
  mutate(has_LA=LA_per_Sample>0)

counts.prop = data.frame(table(Final.Immune.Cells@meta.data$Class, Final.Immune.Cells@meta.data$Final_Immune_Cluster_ID))
counts.prop.perc = group_by(counts.prop, Var2) %>% mutate(percent = Freq/sum(Freq))
head(counts.prop.perc)
aux = aggregate(counts.prop$Freq, by=list(Category=counts.prop$Var1), FUN=sum)
aux$Var2 = "Total"
colnames(aux) <- c("Var1", "Freq", "Var2")
aux$percent = aux$Freq / sum(aux$Freq)
counts.prop.perc = rbind(as.data.frame(counts.prop.perc), aux)

bsize = 20
p1 <- ggplot(counts.prop.perc, aes(x = Var2, y = percent, fill = Var1)) + 
  geom_bar(position="stack",stat = "identity", width=0.8) +
  scale_fill_manual(name="Class", values = c("Endometrioma" = "#7b3294", 
                      "Eutopic endometrium" = "#c2a5cf", 
                      "Peritoneal endometriosis" = "#d9f0d3", 
                      "Unaffected ovary" = "#008837")) +
  scale_x_discrete(limits = rev(levels(counts.prop.perc$Var2))) +
  coord_flip() +
  theme_set(theme_classic(base_size = 25)) +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 270, hjust = 1, color = "Black"),
        axis.text.y = element_text(angle = 0, hjust = 1, colour = "Black"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle('Proportion of Immune cells TrueV2 res0.6 in Major class')

p1

pdf(file = "Plots/Endometriosis_LA/Immune_SubClusters_Proportions_Fig1D.pdf", width = 15, height = 15) 
p1
dev.off()




#Immune cells by LA status (LA status is by sample not patient) in PE

#Final.Immune.Cells <- subset(Immune.Cells.TrueV2,cells=which(!is.na(Immune.Cells.TrueV2$Final_Immune_Cluster_ID))) 

Final.Immune.Cells.PE <- subset(Final.Immune.Cells, subset = Class %in% "Peritoneal endometriosis")

Final.Immune.Cells.PE@meta.data$Final_Immune_Cluster_ID <- factor(Final.Immune.Cells.PE@meta.data$Final_Immune_Cluster_ID, 
                            levels=c("B-cells",
                                     "Effector Memory CD8 T-cells (4)",
                                     "NK cells (2)",
                                     "Th17 T-cells",
                                     "Exhausted CD8 T-cells",
                                     "Resident Memmory CD8 T-cells (2)",
                                     "Early Exhausted CD8 T-cells (2)",
                                     "Macrophages M2",
                                     "NK cells (1)",
                                     "Early Exhaused CD8 T-cells (1)",
                                     "Plasmablast/Plasma cells",
                                     "Dendritic Cells (2)",
                                     "Dendritic Cells (1)",
                                     "Gamma-Delta T-cells (2)",
                                     "Effector CD8 T-cells",
                                     "Effector Memory CD8 T-cells (5)",
                                     "Gamma-Delta T-cells (1)",
                                     "Effector Memory CD8 T-cells (2)",
                                     "Macrophages M1",
                                     "Resident Memmory CD8 T-cells (1)",
                                     "Th2 T-cells",
                                     "Effector Memory CD8 T-cells (3)",
                                     "Effector Memory CD8 T-cells (1)",
                                     "Mast cells"))


Final.Immune.Cells.PE@meta.data <- droplevels(Final.Immune.Cells.PE@meta.data)

Final.Immune.Cells.PE@meta.data <- Final.Immune.Cells.PE@meta.data %>% 
  mutate(has_LA=LA_per_Sample>0)

counts.prop = data.frame(table(Final.Immune.Cells.PE@meta.data$has_LA, Final.Immune.Cells.PE@meta.data$Final_Immune_Cluster_ID))
counts.prop.perc = group_by(counts.prop, Var2) %>% mutate(percent = Freq/sum(Freq))
head(counts.prop.perc)
aux = aggregate(counts.prop$Freq, by=list(Category=counts.prop$Var1), FUN=sum)
aux$Var2 = "Total"
colnames(aux) <- c("Var1", "Freq", "Var2")
aux$percent = aux$Freq / sum(aux$Freq)
counts.prop.perc = rbind(as.data.frame(counts.prop.perc), aux)

bsize = 20
p1 <- ggplot(counts.prop.perc, aes(x = Var2, y = percent, fill = Var1,  
  label = signif(round(percent, digits = 3), digits = 2))) +
  geom_bar(position="stack",stat = "identity", width=0.8) +
  #scale_fill_manual(name="Class", values = c("Endometrioma" = "#7b3294", 
   #                   "Eutopic endometrium" = "#c2a5cf", 
    #                  "Peritoneal endometriosis" = "#d9f0d3", 
     #                 "Unaffected ovary" = "#008837")) +
  scale_x_discrete(limits = rev(levels(counts.prop.perc$Var2))) +
  coord_flip() +
  geom_text(data=subset(counts.prop.perc,Freq != 0), position = position_stack(vjust = 0.5), size = 4, color = "#ffffff") +
  theme_set(theme_classic(base_size = 25)) +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 270, hjust = 1, color = "Black"),
        axis.text.y = element_text(angle = 0, hjust = 1, colour = "Black"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle('Proportion of Immune cells TrueV2 res0.6 in LA Status')



p1

#number of immune cells: first need to reorder cells from alphabetically from last to fisrt

Immune.Cells.TrueV2@meta.data$Immune_Cluster_ID <- factor(Immune.Cells.TrueV2@meta.data$Immune_Cluster_ID, 
                            levels=c("NK cells",
                                     "Mast cells",
                                     "Macrophages M2",
                                     "Macrophage M1/M2",
                                     "Macrophages M1",
                                     "Plasmablast/Plasma cells",
                                     "Plasma cells",
                                     "B-cells",
                                     "Th17 T-cells",
                                     "Th2 T-cells",
                                     "Exhausted CD8 T-cells",
                                     "Effector Memory CD8",
                                     "Cytotoxic CD8 T-cells",
                                     "Central Memory CD8 T-cells (3)",
                                     "Central Memory CD8 T-cells (2)",
                                     "Central Memory CD8 T-cells (1)",
                                     "Activated CD8 T-cells",
                                     "Activated CD8 T-cells/Early Exhausted CD8 T-cells"))


aux = data.frame(table(Immune.Cells.TrueV2@meta.data$Immune_Cluster_ID))
#aux = aux[order(aux$Freq),]
#aux$Var1 <- factor(aux$Var1, levels = aux$Var1)
#order.cell = aux$Var1

p1 <- ggplot(aux, aes(x = Var1, y = Freq)) +
        geom_point(size=2) +
        ylim(0, 300000) +
        scale_y_continuous(breaks=c(0, 5000, 10000, 15000, 20000, 25000, 30000)) +
        #geom_hline(yintercept = c(0, 25000, 50000, 75000, 100000, 125000), linetype="dashed", color = "grey") +
        coord_flip() +
        theme_classic(base_size = 25) +
        theme(plot.title = element_text(hjust = 0.5),
              axis.text.x = element_text(angle = 270, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_blank(),
              legend.position = "left")
p1

pdf(file = "Plots/Endometriosis_LA/Immune_SubClusters_Cell_number_Fig1D.pdf", width = 10, height = 10) 
p1
dev.off()


#number of patients
aux = data.frame(Immune.Cells.TrueV2@meta.data %>% 
      group_by(Immune_Cluster_ID) %>% 
     summarise(n_patient= n_distinct(Patient.No.)))

p1 <- ggplot(aux, aes(x = Immune_Cluster_ID, y = n_patient)) +
        geom_point(size=2) +
        #ylim(0, 30) +
        #scale_y_continuous(breaks=c(0, 15, 20)) +
        #geom_hline(yintercept = c(0, 25000, 50000, 75000, 100000, 125000), linetype="dashed", color = "grey") +
        coord_flip() +
        theme_classic(base_size = 25) +
        theme(plot.title = element_text(hjust = 0.5),
              axis.text.x = element_text(angle = 270, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_blank(),
              legend.position = "left")

p1

pdf(file = "Plots/Endometriosis_LA/Immune_SubClusters_Patient_number_Fig1D.pdf", width = 10, height = 10) 
p1
dev.off()

```


#Figure 1C
```{r}
#DotPlot to ID Immune clusters first ordering cells alphabetically from last to first
Immune.Cells.TrueV2@meta.data$Immune_Cluster_ID <- factor(Immune.Cells.TrueV2@meta.data$Immune_Cluster_ID, 
                            levels=c("NK cells",
                                     "Mast cells",
                                     "Macrophages M2",
                                     "Macrophage M1/M2",
                                     "Macrophages M1",
                                     "Plasmablast/Plasma cells",
                                     "Plasma cells",
                                     "B-cells",
                                     "Th17 T-cells",
                                     "Th2 T-cells",
                                     "Exhausted CD8 T-cells",
                                     "Effector Memory CD8",
                                     "Cytotoxic CD8 T-cells",
                                     "Central Memory CD8 T-cells (3)",
                                     "Central Memory CD8 T-cells (2)",
                                     "Central Memory CD8 T-cells (1)",
                                     "Activated CD8 T-cells",
                                     "Activated CD8 T-cells/Early Exhausted CD8 T-cells"))


p1 <- DotPlot(Immune.Cells.TrueV2, assay = "RNA", features = c("CD3D", "CD8A", "KLRG1", "IL7R", "PRF1", "PDCD1", "CTLA4", "CD4", "GATA3","RORC", "IL17A", "MS4A1", "CD79A", "SDC1", "JCHAIN", "IGHG4", "CD68", "CD163", "IL10", "VEGFA", "KIT", "TPSB2", "KIR2DL4", "NCAM1"), group.by = "Immune_Cluster_ID") + labs(title = "Identifying  cell TrueV2 Sub-cluster by Canonical Markers res0.6") +
 theme_classic(base_size = 20) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.8, colour = "black"), axis.text.y = element_text(colour = "black")) +
  #coord_flip() +
  scale_colour_gradient2(low = "#2166ac", mid = "#f7f7f7", high = "#b2182b")

p1

pdf(file = "Plots/Endometriosis_LA/Immune_SubClusters_Markers_DotPlot_Fig1E.pdf", width = 15, height = 10) 
p1
dev.off()
```


#Figure 1D
#Fold enrichment of Immune clusters in the 4 classes

#Sol's Enrichment/fold change plot
```{r}
library(circlize)  
library(ComplexHeatmap)

plot_df <- Immune.Cells.TrueV2@meta.data %>%
     group_by(Immune_Cluster_ID,Class) %>%
     summarise(count=n())

head(plot_df)

test_enrich<-function(val1,val2){
  res<-plot_df%>%mutate(active=case_when(Immune_Cluster_ID==val1 & Class==val2 ~"pp",
                                    Immune_Cluster_ID==val1 & Class!=val2 ~"pn",
                                    Immune_Cluster_ID!=val1 & Class==val2 ~"np",
                                    Immune_Cluster_ID!=val1 & Class !=val2 ~"nn"))%>%
    group_by(active)%>%
    summarise(n=sum(count,na.rm = T))%>%
    arrange(desc(active))%>%
    select(n)%>%unlist%>%
    matrix(ncol=2,nrow=2)%>%
    fisher.test(alternative = 'g')
  res$p.value}
#plot_df<-get_df("Master_Assigned_Cell_Type")
plot_df$pval=apply(plot_df,1,function(x) test_enrich(x[1],x[2]))
plot_df$padj<-p.adjust(plot_df$pval,"fdr")
observed<-plot_df%>%
  select(Class,Immune_Cluster_ID,count)%>%
  pivot_wider(id_cols=Class,names_from = Immune_Cluster_ID,values_from = count)%>%
  column_to_rownames("Class")
observed[is.na(observed)]<-0
expected<-chisq.test(as.matrix(observed))$expected
fc<-observed/expected
i<-fc<1
fc[i]<--(expected/observed)[i]
pvals<-plot_df%>%
  select(Immune_Cluster_ID,Class,padj)%>%
  pivot_wider(id_cols=Class,names_from = Immune_Cluster_ID,values_from = padj)%>%
  column_to_rownames("Class")
col_fun<-colorRamp2(c(4,0,-4), c("#B2182B","#F7F7F7","#2166AC"))
side <- rowAnnotation(total=anno_barplot(plot_df%>%group_by(Immune_Cluster_ID)%>%summarise(n=sum(count))%>%ungroup%>%select(n)%>%unlist()))
top <- HeatmapAnnotation(total=anno_barplot(plot_df%>%group_by(Class)%>%summarise(n=sum(count))%>%ungroup%>%select(n)%>%unlist))
fc<-t(fc)
pvals<-t(pvals)

p1 <- Heatmap(fc,cluster_rows = F,cluster_columns=F,row_names_side = "left",column_names_rot = 45,col=col_fun,heatmap_legend_param = list(title = "FC"),right_annotation = side, top_annotation = top) 

p1

pdf(file = "Plots/Endometriosis_LA/Immune_SubClusters_Markers_Foldchange_enrichment_Fig1F.pdf", width = 12, height = 10) 
p1
dev.off()

write_csv(plot_df, "data/Immune_cells_FC_Class_Pvalues.csv")

write_csv(fc,"data/Immune_cells_FC_Class_values.csv")


```












#Stattistics for LA High vs Low
```{r}
##stage comparision in Low vs High (not including patient 1 and 2 that have only endometrioma)
Metadata.Stats = read_xlsx("data/LA_Supp_organization.xlsx", sheet = 1)
counts = table(Metadata.Stats$Group[-c(1:2)], Metadata.Stats$Stage.Group[-c(1:2)])
Chisq <- chisq.test(counts)
Chisq

##Low vs High mestral cycle comparision pch = circles, 21= is the type symbol (21 =circle)
Metadata.Stats = read_xlsx("data/LA_Supp_organization.xlsx", sheet = 1)
counts = table(Metadata.Stats$Group[-c(1:2)], Metadata.Stats$Menstral.Cycle.Phase[-c(1:2)])
Chisq <- chisq.test(counts)
Chisq


##Low vs High Age comparision 
Metadata.Stats = read_xlsx("data/LA_Supp_organization.xlsx", sheet = 1)
Metadata.Stats = Metadata.Stats[-c(1:2), c("Group", "Age")]

Low <- c(32, 41, 31)
High <- c(32, 24, 40, 42, 28, 21, 37, 46, 36)

wilcox.test(Low, High)

p1 <- ggplot(Metadata.Stats, aes(x=as.character(Group), y=Age)) +
  geom_boxplot(width = 0.5) + 
  geom_point(pch = 21) +
  theme_classic(base_size = 20) +
  theme(axis.text.x=element_text(angle = -45, vjust = 0.5, colour = "black"), axis.text.y = element_text(colour = "black")) +
  geom_text(x=0.75, y=45, label= "p-value = 1", size = 10)
  
p1


##Low vs High Deep vs superficial comparision by tissue
Metadata.Stats = read_xlsx("data/LA_Supp_organization.xlsx", sheet = 2)
counts = table(Metadata.Stats$Group, Metadata.Stats$Deep_superficial)
Chisq <- chisq.test(counts)
Chisq



##Low vs High mutations status comparision by tissue
Metadata.Stats = read_xlsx("data/LA_Supp_organization.xlsx", sheet = 2)
counts = table(Metadata.Stats$Group[-c(3, 12, 16, 18, 20:22, 28)], Metadata.Stats$ARID1A_Mut[-c(3, 12, 16, 18, 20:22, 28)])
Chisq <- chisq.test(counts)
Chisq


counts = table(Metadata.Stats$Group[-c(3:4, 6:8, 11:14, 16, 18:23, 28)], Metadata.Stats$KRAS_Mut[-c(3:4, 6:8, 11:14, 16, 18:23, 28)])
Chisq <- chisq.test(counts)
Chisq

```













#looking at specific receptors and ligans in EnEpi and Central Memory in endometriosis
```{r}
sel <- subset(d10x.seurat, subset = Master_Assigned_Cell_Type %in% "EnS cells")

sel <- subset(sel, subset = Class %in% c("Peritoneal endometriosis", "Endometrioma"))

Immune.markers = read.delim("data/cell_chat_receptor_ligand_filtered.csv", sep = ",")
head(Immune.markers)

selected.features = unique(as.character(unique(Immune.markers$M2Macs)))


DotPlot(sel, assay = "RNA", features = c("IGF1", "THY1", "ANXA1", "C3", "IGF"), group.by = "Class") + labs(title = "Ligand Receptor in CLASS EnS cells") +
 theme_classic(base_size = 25) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.8, colour="black"), axis.text.y = element_text(color = "black")) +
  coord_flip() +
  scale_colour_gradient2(low = "#2166ac", mid = "#f7f7f7", high = "#b2182b")


sel <- subset(Endometriosis, subset = Immune_Cluster_ID %in% "Plasma cells")

sel <- subset(Immune.Cells.TrueV2, subset = Class %in% c("Peritoneal endometriosis", "Endometrioma"))

sel <- subset(sel, subset = Immune_Cluster_ID %in% "Plasma cells")

DotPlot(sel, assay = "RNA", features = c("IL1B", "NAMPT", "CD69", "GZMA", "HLA-B", "APP", "CLEC2B", "INSR"), group.by = "Class") + labs(title = "Ligand Recipient signals from Activated Plasma cell") +
 theme_classic(base_size = 25) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.8, colour="black"), axis.text.y = element_text(colour = "black")) +
  coord_flip() +
  scale_colour_gradient2(low = "#2166ac", mid = "#f7f7f7", high = "#b2182b")



Endometriosis <- subset(Immune.Cells.TrueV2, subset = Class %in% "Peritoneal endometriosis")

Idents(sel) <- "Immune_Cluster_ID"

DotPlot(sel, assay = "RNA", features = 
          c("NAMPT", "CD69", "GZMA", "HLA-B", "APP", "CLEC2B"), split.by = "Class", cols= "BrBG") + labs(title = "COX expression in Immune cells in endometriosis") +
 theme_classic(base_size = 20) #+
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.8)) +
  #coord_flip() +
  #scale_colour_gradient2(low = "#2166ac", mid = "#f7f7f7", high = "#b2182b")

```






















##Figure1A will be an HE of LA high with big aggregate and a small aggregate





## Fig 1B Beeswarm plot of the #LA in endometrioma and endometriosis.

```{r}
#endometrioma vs Endometriosis LA by block
LA.extended.data = read_xlsx("data/2023-01-25 LA Summary MH v2.xlsx", sheet = 4)

#endometrioma vs Endometriosis LA by blocks stats
Endometrioma <- c(0,6,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,5,1,0,0,0,0,0,0,0)
Endometriossis<- c(0,0,0,0,2,0,0,2,3,0,4,8,10,7,7,9,0,0,1,1,11,0,5,0,0,0,0,0,0,0,0,2,2,3,4,0,0,1,5,0,13,1,3,4,0,2,0,1,5,2,1,1,0,1,1,1,1,1,0,1,1,0,0,1,3,0,0,0,0,0,0,0,0) 

t_test <- t.test(Endometrioma, Endometriossis)
t_test

p1 <- ggplot(LA.extended.data, aes(x=Category, y=Number_of_lymphoid_aggregates, color= Category)) + 
  scale_colour_manual(name="Category", values = c("Endometrioma" = "#7b3294", 
                                                    "Endometriosis" = "#d9f0d3")) + 
  geom_beeswarm(size=3, shape =21, fill= "black", stroke =1.75) + 
  geom_text(x=1, y=12, label=paste0("p-value = ", t_test$p.value), colour="black", family = "san", fontface = "plain", size =5) +
  ggtitle("Number of LA in each Endometrioma and Endometriosis Block") + 
  theme_classic(base_size = 20)

p1


#endometrioma vs Endometriosis LA by part
LA.simple.data = read_xlsx("data/2024-11-06_Combine_scGroup_withPriority2_analyze_only_endometriosis_positive_parts_deidentified_Rfile.xlsx", sheet = 5)
LA.simple.data <- LA.simple.data %>% filter(!grepl("No ", Category))
#endometrioma vs Endometriosis LA by part stats
#Endometrioma <- c(6,0,0,0,0,0,0,0,6,0,0,8,3,0,0,0,0,0,0,0,0,0,0,0,0,0,5)
#Endometriossis<- c(8,0,0,0,0,0,4,0,3,0,12,17,16,0,0,16,2,0,0,0,0,0,0,1,0,0,0,0,4,7,0,0,1,5,0,14,3,4,10,3,4,2,4,0,0,0,0,0,0,3,0,1,17,0,0,0,10,0,1,0,0,0,0,0,0,0,0,0,0,9,0,0,0,29,13,4,1,6,5,7,4,0,0,2,1,0,0,15,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,8,6,8,0,0,0,5,0,1,0,3,0,7,1,0,35,1,12,26,20,7,0,3,0,4,0,0,0,13,2,0,0,0,0,7,0,1,6,0,0,0,0,0,0,0,0,28,2,8,3,0,0,0,0) 
#wil_test <- wilcox.test(Number_of_lymphoid_aggregates ~ Category, data = LA.simple.data)
#wil_test <- wilcox.test(Endometrioma, Endometriossis)
#wil_test

wilcox.test(Number_of_lymphoid_aggregates ~ Category, data = LA.simple.data)

#p1 <- ggplot(LA.simple.data, aes(x=Category, y=Number_of_lymphoid_aggregates, color=Category)) + 
  #scale_colour_manual(name="Category", values = c("Endometrioma" = "#7b3294", 
   #                                                 "Endometriosis" = "#d9f0d3")) + 
  #geom_quasirandom(size=2.5) + 
  #geom_text(x=0.75, y=30, label="p-value = 0.2499", colour="black", family = "san", fontface = "plain", size =5) +
  #ggtitle("Number of LA in each Endometrioma and Endometriosis part") + 
  #theme_classic(base_size = 20)

p1 <- ggplot(LA.simple.data, aes(x=Category, y=Number_of_lymphoid_aggregates)) + 
  geom_quasirandom(size=2.5) + 
  geom_text(x=0.75, y=30, label="p-value = 0.2499", colour="black", family = "san", fontface = "plain", size =5) +
  ggtitle("Number of LA in each Endometrioma and Endometriosis part") + 
  theme_classic(base_size = 20)


p1


#Number of LA in Endometrioma and Endometriosis by part with patients labeled
LA.summarised <- LA.simple.data %>%
  select(`PAPER ID`, Number_of_lymphoid_aggregates, Category) %>% 
  group_by(`PAPER ID`, Category) %>% 
  summarise(total_lymph_agg = sum(Number_of_lymphoid_aggregates)) %>% 
  mutate(plotting_id = paste0(`PAPER ID`, Category))

LA.summarised$plotting_id <- factor(LA.summarised$plotting_id, levels = LA.summarised$plotting_id[order(LA.summarised$total_lymph_agg)])

wilcox.test(total_lymph_agg ~ Category, data = LA.summarised)

#ggplot(LA.summarised, aes(x=plotting_id, y=total_lymph_agg, color=Category)) + 
  #scale_colour_manual(name="Category", values = c("Endometrioma" = "#7b3294", 
                                                   "Endometriosis" = "#d9f0d3")) + 
  #geom_quasirandom(size=2.5) + 
  #scales= "free_x" means that it can ignore any point that is not in that category
  #facet_wrap(~Category, scales = "free_x") + 
   #geom_text(x=1, y=15, label="p-value = 0.0422", colour="black", family = "san", fontface = "plain", size =5) +
  #ggtitle("Number of LA in each Endometrioma and Endometriosis part") + 
  #theme_classic(base_size = 20) +
  #theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5)) +
  #geom_hline(yintercept = 0, linetype = 2) #+
  #scale_x_discrete(labels = setNames(LA.summarised$`PAPER ID`, LA.summarised$plotting_id))
  
  

ggplot(LA.summarised, aes(x=plotting_id, y=total_lymph_agg)) + 
  geom_quasirandom(size=2.5) + 
  #scales= "free_x" means that it can ignore any point that is not in that category
  facet_wrap(~Category, scales = "free_x") + 
  #geom_text(x=1, y=15, label="p-value = 0.001217", colour="black", family = "san", fontface = "plain", size =5) +
  ggtitle("Number of LA in each Endometrioma and Endometriosis in each patient") + 
  theme_classic(base_size = 20) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5)) +
  geom_hline(yintercept = 0, linetype = 2) +
  scale_x_discrete(labels = setNames(LA.summarised$`PAPER ID`, LA.summarised$plotting_id))
  


#Beeswarm plot of total La by patient

LA.summarised <- LA.simple.data %>%
  select(`PAPER ID`, Number_of_lymphoid_aggregates, Category) %>% 
  group_by(`PAPER ID`, Category) %>% 
  summarise(total_lymph_agg = sum(Number_of_lymphoid_aggregates)) %>% 
  mutate(plotting_id = paste0(`PAPER ID`, Category))

LA.summarised$plotting_id <- factor(LA.summarised$plotting_id, levels = LA.summarised$plotting_id[order(LA.summarised$total_lymph_agg)])

wilcox.test(total_lymph_agg ~ Category, data = LA.summarised)

ggplot(LA.summarised, aes(x=Category, y=total_lymph_agg)) + 
     geom_quasirandom(size=2.5) + 
     geom_text(x=1, y=60, label="p-value = 0.001217", colour="black", family = "san", fontface = "plain", size =5) +
     ggtitle("Number of LA in each Endometrioma and Endometriosis in patient") + 
     theme_classic(base_size = 20) +
     geom_hline(yintercept = 0, linetype = 2) 

```

## Fig 1C
```{r}
#Heatmap of peritoneal endometriosis lesions in parts
library(pheatmap)

la_summary = read_xlsx("data/2023-01-25 LA Summary MH v2.xlsx", sheet = 2)
la_summary[is.na(la_summary)] = -10

my_cluster_col = as.data.frame(la_summary[3])
my_cluster_col$Patient = as.character(my_cluster_col$Patient)
rownames(my_cluster_col) <- paste0("row_", seq(nrow(my_cluster_col)))
la_summary = la_summary[c(-8, -14:-41),c(-1:-3, -11:-31)]  #need to adjust rows depending on the samples being included.

test_labels <- as.matrix(la_summary)
rownames(la_summary) <- paste0("row_", seq(nrow(la_summary)))
test_labels[test_labels < 0] <- "\u2217\u2217"
test_labels[test_labels < 0] <- "\uFF0A"

p1 <- pheatmap(la_summary,
         cluster_cols = F,
         display_numbers = test_labels,
         fontsize_number=12,
         cellheight=20,
         annotation_row = my_cluster_col,
         clustering_distance_row = "euclidean",
         main = "12 samples euclidean")

p1


#or plot based on Beeswarm plot
LA.simple.data = read_xlsx("data/2023-01-25 LA Summary MH v2.xlsx", sheet = 5)
LA.simple.data.sel = LA.simple.data[c(-1,-2,-3, -5, -10, -23, -33, -36, -47), c("LA_Activity", "Category", "Number_of_lymphoid_aggregates")]

p1 <- ggplot(LA.simple.data.sel, aes(x=LA_Activity, y=Number_of_lymphoid_aggregates, color=LA_Activity)) + 
  #scale_colour_manual(name="Category", values = c("Endometrioma" = "#7b3294", 
   #                                                 "Endometriosis" = "#d9f0d3")) + 
  geom_beeswarm(size=2.5) + 
  #geom_text(x=1, y=15, label="p-value = 0.4654", colour="black", family = "san", fontface = "plain", size =5) +
  ggtitle("Number of LA in patients defined as High and low by part") + 
  theme_classic(base_size = 30)
  

p1

```






## Fig 1F
PCA based on the following cell proportion for each samples:

Follow the below steps for PCA analysis 

1. import Annotation tracker to add the LA_Status Epi_data tab - we need that col to label ggplot ( Master_Assigned_Cell_Type is already included in mate.data)
2. import `broom`package to parse PCA output from `prcomp`
3. Only meta.data from Seurat obj is needed for the PCA analysis
4. If the label on plot is too crowded. consider using `Patient.No.`

```{r}
d10x.seurat <- readRDS("data/fixed_annotated_aux.seurat.rds")@meta.data
excel_path <- "data/Annotation tracker.xlsx"
sheets <- readxl::excel_sheets(excel_path)
Annotation_tracker <- sheets %>% 
                      lapply(. , function(s){ read_excel(excel_path, sheet = s) %>% 
                                              mutate(across(where(is.character),str_trim))
                                              }) %>%
                      setNames(sheets)

Annotation_tracker$Epi_Data 
```


```{r}
immue_endo_t<- d10x.seurat@meta.data%>%
                  #filter(Tissue_LA_Activity == "Yes") %>%
                  #filter(Master_Assigned_Cell_Type) %>%
                  #filter(! is.na(Number_of_lymphoid_aggregates)) %>%
                  group_by(SampleName,Master_Assigned_Cell_Type) %>%
                  summarise(count=n()) %>%
                  mutate(prop= count/sum(count)*100)

anno_df <- d10x.seurat@meta.data  %>%
           left_join(Annotation_tracker$Epi_Data, by=c("Patient.No."="Patient_ID")) %>%
           dplyr::filter(LA_Status != "Exclude")  %>% 
           distinct(Patient.No.,LA_Status,SampleName) %>%
           column_to_rownames("SampleName")

heatmap_df <- immue_endo_t  %>% 
              dplyr::select(Master_Assigned_Cell_Type,prop,SampleName) %>%
              spread(Master_Assigned_Cell_Type, prop) %>%
              column_to_rownames("SampleName") %>%
              as.matrix() %>%
              replace(.,is.na(.),0) %>%
              t()

heatmap_df %>% 
t %>%
prcomp(scale = TRUE) %>%
augment(.) %>%
inner_join(anno_df %>% rownames_to_column(".rownames"), by=".rownames") %>%
  ggplot(aes(x=.fittedPC1, y=.fittedPC2,color=LA_Status,label=.rownames))+
  geom_point(size=5) +
  geom_text(nudge_x = 0.2, nudge_y = 0.2,angle=20,size=3) +
  labs(x="PC1",y="PC2") + 
  theme_half_open(12) 


```

# Final Fig pannel combine using cowplot

```{r}
fig1top <- plot_grid(fig1a, fig1b, labels = c("A", "B"), ncol = 2, label_size = 10, rel_widths = c(1, 1.15) )
fig1middle <- plot_grid(fig1c, fig1d, NULL, labels = c("C", "D"), ncol = 3, label_size = 10)
fig1bottom <- plot_grid(fig1e, fig1f, NULL, labels = c("E", "F"), ncol = 3, label_size = 10)
fig1 <- plot_grid(fig1top, fig1middle,fig1bottom ,ncol = 1, label_size = 10, align = 'v', hjust=-6, rel_heights = c(0.5, 0.25))
```

#save figures

```{r}
ggsave("data/fig1.pdf",fig1, width = 210, 
       height = 148.5, 
       units = "mm")
```


#DEG of B-cell in LA true vs false and volcano plot
```{r}
#DEG

Bcells<- subset(Final.Immune.Cells.PE, subset = Final_Immune_Cluster_ID %in% "B-cells")

Idents(Bcells) <- Bcells$has_LA #this is to say which columns to use

DEG <- FindMarkers(Bcells, ident.1 = "TRUE", ident.2 = "FALSE")

DEG <- DEG %>% rownames_to_column("genes")

#volcano plot
fc.p = 0.5
pv.p = 0.05
DEG$STATUS = "NOT.SIG"
DEG[DEG$avg_log2FC < -fc.p & DEG$p_val_adj < pv.p, ]$STATUS = "Down"
DEG[DEG$avg_log2FC > fc.p & DEG$p_val_adj < pv.p, ]$STATUS = "Up"

p1 <- ggplot(DEG, aes(x = avg_log2FC, y = -log10(p_val_adj))) +
  geom_point(aes(color = STATUS), cex = 1.45) +
  scale_color_manual(values = c("darkgreen", "grey", "red")) +
  ylim(0, 80) +
  theme_bw(base_size = 15) + 
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5)) +
  geom_text_repel(
    data          = subset(DEG, avg_log2FC > 0.5),
    aes(label = genes),
    segment.size  = 0.4,
    direction     = "y",
    segment.color = "grey50",
    hjust         = 0
  ) +
  geom_text_repel(
    data          = subset(DEG, avg_log2FC < -0.5),
    aes(label = genes),
    segment.size  = 0.4,
    direction     = "y",
    segment.color = "grey50",
    hjust         = 1
  ) +
  geom_hline(yintercept = -log10(pv.p), linetype="dotted", color = "black", size= 0.5) +
  geom_vline(xintercept = -fc.p, linetype="dotted", color = "black", size= 0.5) +
  geom_vline(xintercept = fc.p, linetype="dotted", color = "black", size= 0.5) #+
  #labs(title = paste0(feature, " : ", c, " (n=", nDEG, ") FC: ", fc.p, " pval = ", pv.p))

p1



#doing DEG and volcano plots based pseudobulk
library("SingleCellExperiment")

#first convert objects to singleCellExperiiment object
Bcells2 <- as.SingleCellExperiment(Final.Immune.Cells.PE)

#need library glmGampoi
BiocManager::install("glmGamPoi")
library("glmGamPoi")

#This is the grouping of cells
reduced_sce <- pseudobulk(Bcells2, group_by = vars(Patient.No., has_LA, SampleName, harmonized_major_plus_immune_cell_type), 
                          n_cells = n())

save("reduced_sce", file = "reduced_sce.rda")

# Remove NA's
reduced_sce <- reduced_sce[,!is.na(reduced_sce$SampleName)]

colData(reduced_sce)$Patient.No. <- factor(colData(reduced_sce)$Patient.No.)
colData(reduced_sce)$has_LA <- factor(colData(reduced_sce)$has_LA)

# Use DESeq2's size factor calculation procedure
fit <- glm_gp(reduced_sce, design = ~ has_LA + harmonized_major_plus_immune_cell_type + Patient.No., size_factor = "ratio", verbose = TRUE)

#to test the differential expression
es <- test_de(fit, contrast = cond(has_LA = "TRUE") - cond(has_LA = "FALSE"))
```


#pathway analysis
```{r}
BiocManager::install("clusterProfiler")

BiocManager::install("msigdbr")

library(clusterProfiler)
library(msigdbr)

m_df <- msigdbr(species = "Homo sapiens")%>%filter(str_detect(gs_name,"REACTOME"))%>%
    distinct(gs_name,gene_symbol)

temp <- DEG %>% mutate(keep = case_when(avg_log2FC > 0 & pct.1 > 0.8 ~ TRUE,
                                avg_log2FC < 0 & pct.2 > 0.8 ~ TRUE,
                                TRUE ~ FALSE)) %>%
  filter(keep, p_val_adj < 0.05) %>%
  mutate(STATUS = ifelse(avg_log2FC > 0, "Up", "Down"))

#temp<-DEG%>%filter(p_val_adj<0.05 & abs(avg_log2FC)>.5)
oraRes<-temp%>%split(temp$STATUS)%>%lapply(function(x) enricher(x$genes,TERM2GENE = m_df))
temp<-lapply(names(oraRes),function(x) 
  oraRes[[x]]%>%as.data.frame()%>%mutate(cluster=x))%>%bind_rows()


barplot(oraRes$Up)

dotplot(oraRes$Up)

dotplot(oraRes$Down)
```