---
title: "Figure 1"
author: "Pak Yu"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---
# Figure 1 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Library loading 


```{r ,message = FALSE}
library(Seurat)
library(tidyverse)
library(cowplot)
library(broom)
library(readxl)
library(ggplot2)
library(gdata)
library(ggbeeswarm)
library(beeswarm)
library(pheatmap)
```


```{r}
getwd()

setwd("/home/harom/Code/sc_endo_la_immune")
```


#Immune cells in all the classes
```{r}
#load Immune cells RDS file
Immune.Cells.TrueV2<- readRDS("data/Immune_Cell_annotation.rds")
```


#Getting the Fold change enrichment accross class
```{r}
table.class.cell = table(Immune.Cells.TrueV2@meta.data$Immune_Cluster_ID, Immune.Cells.TrueV2@meta.data$Class)
table.class.cell.perc = table.class.cell/rowSums(table.class.cell)

obs.class.perc = table(Immune.Cells.TrueV2@meta.data$Class)/length(Immune.Cells.TrueV2@meta.data$Class)
table.fc = table.class.cell
table.fc = is.na(table.fc)

for (class in colnames(table.class.cell.perc)) {
        for (cell.type in rownames(table.class.cell.perc)) {
                expected = table.class.cell.perc[cell.type, class]
                obs = obs.class.perc[class]
                if (obs > expected) {
                        ratio = -(obs/expected)
                } else {
                        ratio = expected/obs
                }
                #ratio = expected/obs
                
                table.fc[cell.type, class] = ratio
        }
}


quantile_breaks <- function(xs, n = 10) {
        breaks <- quantile(xs, probs = seq(0, 1, length.out = n))
        breaks[!duplicated(breaks)]
}
mat_breaks <- quantile_breaks(table.fc, n = 11)

table.fc = table.fc[,c("Endometrioma", "Endometriosis", "Eutopic Endometrium", "Unaffected ovary")]


pheatmap(t(table.fc), 
         display_numbers = signif(t(table.fc), 3),
         color             = colorRampPalette(c("#81B8D5", "white", "#E37F65"))(10),
         breaks            = mat_breaks,
         cluster_rows      = F,
         cluster_cols      = F,
         fontsize          = 14)
```


#looking at specific receptors and ligans in EnEpi and Central Memory in endometriosis
```{r}
sel <- subset(d10x.seurat, subset = Master_Assigned_Cell_Type %in% "EnS cells")

sel <- subset(sel, subset = Class %in% c("Peritoneal endometriosis", "Endometrioma"))

Immune.markers = read.delim("data/cell_chat_receptor_ligand_filtered.csv", sep = ",")
head(Immune.markers)

selected.features = unique(as.character(unique(Immune.markers$M2Macs)))


DotPlot(sel, assay = "RNA", features = c("IGF1", "THY1", "ANXA1", "C3", "IGF"), group.by = "Class") + labs(title = "Ligand Receptor in CLASS EnS cells") +
 theme_classic(base_size = 25) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.8, colour="black"), axis.text.y = element_text(color = "black")) +
  coord_flip() +
  scale_colour_gradient2(low = "#2166ac", mid = "#f7f7f7", high = "#b2182b")


sel <- subset(Endometriosis, subset = Immune_Cluster_ID %in% "Plasma cells")

sel <- subset(Immune.Cells.TrueV2, subset = Class %in% c("Peritoneal endometriosis", "Endometrioma"))

sel <- subset(sel, subset = Immune_Cluster_ID %in% "Plasma cells")

DotPlot(sel, assay = "RNA", features = c("IL1B", "NAMPT", "CD69", "GZMA", "HLA-B", "APP", "CLEC2B", "INSR"), group.by = "Class") + labs(title = "Ligand Recipient signals from Activated Plasma cell") +
 theme_classic(base_size = 25) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.8, colour="black"), axis.text.y = element_text(colour = "black")) +
  coord_flip() +
  scale_colour_gradient2(low = "#2166ac", mid = "#f7f7f7", high = "#b2182b")



Endometriosis <- subset(Immune.Cells.TrueV2, subset = Class %in% "Peritoneal endometriosis")

Idents(sel) <- "Immune_Cluster_ID"

DotPlot(sel, assay = "RNA", features = 
          c("NAMPT", "CD69", "GZMA", "HLA-B", "APP", "CLEC2B"), split.by = "Class", cols= "BrBG") + labs(title = "COX expression in Immune cells in endometriosis") +
 theme_classic(base_size = 20) #+
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.8)) +
  #coord_flip() +
  #scale_colour_gradient2(low = "#2166ac", mid = "#f7f7f7", high = "#b2182b")

```






















##Figure1A will be an HE of LA high with big aggregate and a small aggregate





## Fig 1B Beeswarm plot of the #LA in endometrioma and endometriosis.

```{r}
#endometrioma vs Endometriosis LA by block
LA.extended.data = read_xlsx("data/2023-01-25 LA Summary MH v2.xlsx", sheet = 4)

#endometrioma vs Endometriosis LA by blocks stats
Endometrioma <- c(0,6,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,5,1,0,0,0,0,0,0,0)
Endometriossis<- c(0,0,0,0,2,0,0,2,3,0,4,8,10,7,7,9,0,0,1,1,11,0,5,0,0,0,0,0,0,0,0,2,2,3,4,0,0,1,5,0,13,1,3,4,0,2,0,1,5,2,1,1,0,1,1,1,1,1,0,1,1,0,0,1,3,0,0,0,0,0,0,0,0) 

t_test <- t.test(Endometrioma, Endometriossis)
t_test

p1 <- ggplot(LA.extended.data, aes(x=Category, y=Number_of_lymphoid_aggregates, color= Category)) + 
  scale_colour_manual(name="Category", values = c("Endometrioma" = "#7b3294", 
                                                    "Endometriosis" = "#d9f0d3")) + 
  geom_beeswarm(size=3, shape =21, fill= "black", stroke =1.75) + 
  geom_text(x=1, y=12, label=paste0("p-value = ", t_test$p.value), colour="black", family = "san", fontface = "plain", size =5) +
  ggtitle("Number of LA in each Endometrioma and Endometriosis Block") + 
  theme_classic(base_size = 20)

p1


#endometrioma vs Endometriosis LA by part
LA.simple.data = read_xlsx("data/2023-01-25 LA Summary MH v2.xlsx", sheet = 5)

#endometrioma vs Endometriosis LA by part stats
Endometrioma <- c(6,0,0,0,0,6,0,0)
Endometriossis<- c(0,0,4,3,0,12,0,17,16,0,16,2,0,0,0,0,0,0,0,0,0,0,4,7,0,0,1,5,0,6,14,3,4,0,10,3,4,2,4,0,0,0,0,0,0) 

wil_test <- wilcox.test(Endometrioma, Endometriossis)

p1 <- ggplot(LA.simple.data, aes(x=Category, y=Number_of_lymphoid_aggregates, color=Category)) + 
  scale_colour_manual(name="Category", values = c("Endometrioma" = "#7b3294", 
                                                    "Endometriosis" = "#d9f0d3")) + 
  geom_beeswarm(size=2.5) + 
  geom_text(x=1, y=15, label="p-value = 0.4654", colour="black", family = "san", fontface = "plain", size =5) +
  ggtitle("Number of LA in each Endometrioma and Endometriosis part") + 
  theme_classic(base_size = 20)

p1
```

## Fig 1C
```{r}
#Heatmap of peritoneal endometriosis lesions in parts
library(pheatmap)

la_summary = read_xlsx("data/2023-01-25 LA Summary MH v2.xlsx", sheet = 2)
la_summary[is.na(la_summary)] = -10

my_cluster_col = as.data.frame(la_summary[3])
my_cluster_col$Patient = as.character(my_cluster_col$Patient)
rownames(my_cluster_col) <- paste0("row_", seq(nrow(my_cluster_col)))
la_summary = la_summary[c(-8, -14:-41),c(-1:-3, -11:-31)]  #need to adjust rows depending on the samples being included.

test_labels <- as.matrix(la_summary)
rownames(la_summary) <- paste0("row_", seq(nrow(la_summary)))
test_labels[test_labels < 0] <- "\u2217\u2217"
test_labels[test_labels < 0] <- "\uFF0A"

p1 <- pheatmap(la_summary,
         cluster_cols = F,
         display_numbers = test_labels,
         fontsize_number=12,
         cellheight=20,
         annotation_row = my_cluster_col,
         clustering_distance_row = "euclidean",
         main = "12 samples euclidean")

p1


#or plot based on Beeswarm plot
LA.simple.data = read_xlsx("data/2023-01-25 LA Summary MH v2.xlsx", sheet = 5)
LA.simple.data.sel = LA.simple.data[c(-1,-2,-3, -5, -10, -23, -33, -36, -47), c("LA_Activity", "Category", "Number_of_lymphoid_aggregates")]

p1 <- ggplot(LA.simple.data.sel, aes(x=LA_Activity, y=Number_of_lymphoid_aggregates, color=LA_Activity)) + 
  #scale_colour_manual(name="Category", values = c("Endometrioma" = "#7b3294", 
   #                                                 "Endometriosis" = "#d9f0d3")) + 
  geom_beeswarm(size=2.5) + 
  #geom_text(x=1, y=15, label="p-value = 0.4654", colour="black", family = "san", fontface = "plain", size =5) +
  ggtitle("Number of LA in patients defined as High and low by part") + 
  theme_classic(base_size = 30)
  

p1

```






## Fig 1F
PCA based on the following cell proportion for each samples:

Follow the below steps for PCA analysis 

1. import Annotation tracker to add the LA_Status Epi_data tab - we need that col to label ggplot ( Master_Assigned_Cell_Type is already included in mate.data)
2. import `broom`package to parse PCA output from `prcomp`
3. Only meta.data from Seurat obj is needed for the PCA analysis
4. If the label on plot is too crowded. consider using `Patient.No.`

```{r}
d10x.seurat <- readRDS("data/fixed_annotated_aux.seurat.rds")@meta.data
excel_path <- "data/Annotation tracker.xlsx"
sheets <- readxl::excel_sheets(excel_path)
Annotation_tracker <- sheets %>% 
                      lapply(. , function(s){ read_excel(excel_path, sheet = s) %>% 
                                              mutate(across(where(is.character),str_trim))
                                              }) %>%
                      setNames(sheets)

Annotation_tracker$Epi_Data 
```


```{r}
immue_endo_t<- d10x.seurat@meta.data%>%
                  #filter(Tissue_LA_Activity == "Yes") %>%
                  #filter(Master_Assigned_Cell_Type) %>%
                  #filter(! is.na(Number_of_lymphoid_aggregates)) %>%
                  group_by(SampleName,Master_Assigned_Cell_Type) %>%
                  summarise(count=n()) %>%
                  mutate(prop= count/sum(count)*100)

anno_df <- d10x.seurat@meta.data  %>%
           left_join(Annotation_tracker$Epi_Data, by=c("Patient.No."="Patient_ID")) %>%
           dplyr::filter(LA_Status != "Exclude")  %>% 
           distinct(Patient.No.,LA_Status,SampleName) %>%
           column_to_rownames("SampleName")

heatmap_df <- immue_endo_t  %>% 
              dplyr::select(Master_Assigned_Cell_Type,prop,SampleName) %>%
              spread(Master_Assigned_Cell_Type, prop) %>%
              column_to_rownames("SampleName") %>%
              as.matrix() %>%
              replace(.,is.na(.),0) %>%
              t()

heatmap_df %>% 
t %>%
prcomp(scale = TRUE) %>%
augment(.) %>%
inner_join(anno_df %>% rownames_to_column(".rownames"), by=".rownames") %>%
  ggplot(aes(x=.fittedPC1, y=.fittedPC2,color=LA_Status,label=.rownames))+
  geom_point(size=5) +
  geom_text(nudge_x = 0.2, nudge_y = 0.2,angle=20,size=3) +
  labs(x="PC1",y="PC2") + 
  theme_half_open(12) 


```

# Final Fig pannel combine using cowplot

```{r}
fig1top <- plot_grid(fig1a, fig1b, labels = c("A", "B"), ncol = 2, label_size = 10, rel_widths = c(1, 1.15) )
fig1middle <- plot_grid(fig1c, fig1d, NULL, labels = c("C", "D"), ncol = 3, label_size = 10)
fig1bottom <- plot_grid(fig1e, fig1f, NULL, labels = c("E", "F"), ncol = 3, label_size = 10)
fig1 <- plot_grid(fig1top, fig1middle,fig1bottom ,ncol = 1, label_size = 10, align = 'v', hjust=-6, rel_heights = c(0.5, 0.25))
```

#save figures

```{r}
ggsave("data/fig1.pdf",fig1, width = 210, 
       height = 148.5, 
       units = "mm")
```





