---
title: "Figure 2"
author: "Marcela Haro"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---
# Figure 2 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
getwd()
```




## Library loading 

```{r ,message = FALSE}

library(Seurat)
library(tidyverse)
library(cowplot)
library(broom)
library(ggplot2)
library(dplyr)
library(reshape)
library(clusterProfiler)
library(readxl)
library(ggrepel)
library(clustree)
library(pheatmap)
library(sankeyNetwork)
```


#Loading the scRNAseq data and loading into seurat object
```{r}
#reading the Seurat object RDS file
d10x.seurat <- readRDS("data/fixed_annotated_aux.seurat.rds")
```


```{r}
##To confirm annotation
table(d10x.seurat@meta.data$Master_Assigned_Cell_Type)
```



#adding more columns to meta data: based on pathology LA status classified patients as LA high or low, also have a colum for each tissue to say if it has any LAs(yes/no)
```{r}
#adding more columns to metadata about the LA immune activity at patient leve (High/Low); this is Yes/No LA activity for each tissue specimen (Yes/No).

##Note tissue activity need to be updated in the annoatation tracker file. Just use the immune_activity column that is coming from the patient_la_anno

excel_path <- "data/Annotation tracker.xlsx"
sheets <- readxl::excel_sheets(excel_path)
Annotation_tracker <- sheets %>% 
                      lapply(. , function(s){ read_excel(excel_path, sheet = s) %>%
                                              mutate(across(everything(),as.factor))}) %>%
                      setNames(sheets)
```

```{r}
Annotation_tracker$Clusters <- Annotation_tracker$Clusters %>% 
                            mutate(Master_Assigned_Cell_Type= if_else(is.na(Updated_Anno),Master_Assigned_Cell_Type,Updated_Anno) ) %>% 
                            dplyr::select(-Updated_Anno)

back_up_meta_table <- d10x.seurat@meta.data 

d10x.seurat@meta.data <- back_up_meta_table  %>%
              rownames_to_column() %>%
              left_join(Annotation_tracker$Major_groups,by=c("seurat_clusters"="Cluster_Number")) %>%
              left_join(Annotation_tracker$Patient_LA_Status_Tissue_Level, by=c("orig.ident"= "Orig.ident")) %>%
              left_join(Annotation_tracker$Epi_Data %>% mutate(Patient_ID = as.integer(Patient_ID)), by=c("Patient.No."="Patient_ID")) %>%
  column_to_rownames()

```



```{r}
#UMAPS of all the cells

#DimPlot(d10x.seurat, reduction = "umap", label =FALSE, raster = FALSE) ##This is a umap with all the clusters

cluster.cols = c("#67001f", "#a50026", "#d73027", "#f46d43", "#fdae61", "#fee090", "#ffffbf", "#a6dba0", "#1a9850", "#e0f3f8", "#abd9e9")

p1 <- DimPlot(object = d10x.seurat, pt.size = 0.1, group.by ="Master_Assigned_Cell_Type", cols = cluster.cols, label = T, raster=FALSE) +
  theme_classic(base_size = 20) ##this is a umap with the clusters grouped with their cell identity

p1

#pdf(file = "Plots/Endometriosis_LA_Immune/Original Cell Clusters.pdf", width = 15, height = 15) 
#p1
#dev.off()

FeaturePlot(d10x.seurat, features = c("KIT", "GNLY", "CD3D", "LYZ", "CD79A", "GZMB", "ACTA2", "DCN", "KRT18", "COL1A1", "EPCAM")) +
  theme_classic(base_size = 20)
```



#pull out Immune cells from the data set
```{r}
###pull out just the Immune cells
Immune.Cells <- subset(d10x.seurat, subset = Master_Assigned_Cell_Type %in% c("B/Plasma cells", "T/NK cells", "Myeloid cells", "Mast cells"))
```


```{r}
Immune.Cells.Subset <- Immune.Cells

##Subclustering just the Immune cells from LA high low
Immune.cells.subclustering <- FindNeighbors(Immune.Cells.Subset, dims = 1:20)
Immune.cells.subclustering <- FindClusters(Immune.cells.subclustering, resolution = 3)
Immune.cells.subclustering <- RunUMAP(Immune.cells.subclustering, dims = 1:20)

p1 <- DimPlot(Immune.cells.subclustering, reduction = "umap", label =  TRUE, raster = FALSE) + 
  theme_classic(base_size = 20) +
  ggtitle('UMAP of Immune cells subclusters Res 3')

p1

FeaturePlot(Immune.cells.subclustering, features = c("KIT", "GNLY", "CD3D", "LYZ", "CD79A", "GZMB", "ACTA2", "DCN", "KRT18", "COL1A1", "EPCAM")) +
  theme_classic(base_size = 20)

#removing levels that are empty (only keeping the values in the vectors I want) and getting the numb tisseue, n_patient, and n_cells
Immune.cells.subclustering@meta.data <- droplevels(Immune.cells.subclustering@meta.data)

Immune.cells.subclustering@meta.data %>% 
      group_by(seurat_clusters %>% as.character()) %>% 
     summarise(n_tissue=n_distinct(orig.ident), n_patient= n_distinct(Patient.No.), n=n()) %>%
      print(n=92)
```


```{r}
#Immune markers to identify immune cells
Immune.markers = read.delim("data/Immune_Marker.csv", sep = ",")
head(Immune.markers)

#selected.features = unique(c(as.character(unique(Immune.markers$Positive)), as.character(unique(Immune.markers$Negative))))
selected.features = unique(c(as.character(unique(Immune.markers$Positive))))
```

```{r}
p1 <-DotPlot(Immune.cells.subclustering, assay = "RNA", features = selected.features) + labs(title = "Identifying Immune cell Sub-cluster by Canonical Markers res3") +
 theme_classic(base_size = 20) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.8), axis.text.y = element_text(size = 9)) +
  coord_flip() +
  scale_colour_gradient2(low = "#2166ac", mid = "#f7f7f7", high = "#b2182b")

p1
```


#Adding the Epithelial and Mesenchymal cells to the Immune cells subsetting dot plot
```{r}
#To add cell types to the dot plot of the subclusters of Immune-cells
#need to add a new column in the d10xseurat.filt object to mark the cells that you wanna retrieve
d10x.seurat@meta.data$id.cells = rownames(d10x.seurat@meta.data)
d10x.seurat@meta.data$selected.cells = "No"
d10x.seurat@meta.data$subcluster.name=NA

id.cells.annot = rownames(Immune.cells.subclustering@meta.data)
d10x.seurat@meta.data$selected.cells[match(id.cells.annot, d10x.seurat@meta.data$id.cells)] = "Yes"
d10x.seurat@meta.data$subcluster.name[match(id.cells.annot, d10x.seurat@meta.data$id.cells)] = as.character(Immune.cells.subclustering@meta.data$seurat_clusters)
table(d10x.seurat@meta.data$selected.cells)
table(d10x.seurat@meta.data$subcluster.name)

d10x.seurat@meta.data$selected.cells[d10x.seurat@meta.data$active.cluster %in% c("Epithelial cells", "Mesenchymal")] = "Yes"
d10x.seurat@meta.data$selected.cells[d10x.seurat@meta.data$active.cluster == "Mesenchymal cells"] = "Yes"
#to add subcluster name
d10x.seurat@meta.data$subcluster.name[d10x.seurat@meta.data$active.cluster == "Epithelial cells"] = "Epithelial cells"
d10x.seurat@meta.data$subcluster.name[d10x.seurat@meta.data$active.cluster == "Mesenchymal cells"] = "Mesenchymal cells"
table(d10x.seurat@meta.data$selected.cells)
table(d10x.seurat@meta.data$subcluster.name)
#then subset the marked cells
sc.selected = subset(d10x.seurat, subset = selected.cells == "Yes")

DotPlot(sc.selected, group.by = "subcluster.name", assay = "RNA", features = selected.features) + labs(title = "Identifying Immune-cells subcluster by Canonical Markers res0.3") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.8, size = 8), axis.text.y = element_text(size = 8)) +
  coord_flip() +
  scale_colour_gradient2(low = "#2166ac", mid = "#f7f7f7", high = "#b2182b")
```



```{r}
Immune.subcluster.DEG <- FindAllMarkers(Immune.cells.subclustering, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)

write_csv(Immune.subcluster.DEG, "data/Immune_subcluster_DEG_list.csv")
```


#trying superCAT to see how cluster correlate together
```{r}
markers = read.delim("data/Immune_subcluster_DEG_list.csv", sep = ",")

ng = 100
p.v = 0.05
fc = 0
features.deg <- markers %>%
  filter(p_val_adj < p.v & avg_log2FC > fc) %>%
  group_by(cluster) %>%
  top_n(ng)

aux = DotPlot(Immune.cells.subclustering, features = unique(features.deg$gene))
aux.data = aux$data

data.avg.exp.scaled = data.frame(row.names = unique(features.deg$gene))

for (o in unique(aux.data$id) ) {
  aux.data.filt = aux.data[which(aux.data$id == o),]
  rownames(aux.data.filt) <- aux.data.filt$features.plot
  data.avg.exp.scaled = cbind(data.avg.exp.scaled, aux.data.filt[, 'avg.exp.scaled', drop=FALSE])
}

colnames(data.avg.exp.scaled) <- paste(unique(aux.data$id), " (n=", table(features.deg$cluster), ")", sep = "")
pheatmap(data.avg.exp.scaled, 
         show_rownames = F, 
         main = paste0("DEG expression signature clusters (n=", length(unique(features.deg$gene)), ")"), 
         cutree_cols = 10 )
```



#Figure 2A
#remove non-Immune cell clusters
```{r}
#Pull out True Immune cells.
Immune.Cells.True <- subset(Immune.cells.subclustering, subset = seurat_clusters %in% c(8, 12, 21, 23, 25, 36, 41, 43, 47, 53, 57:58, 60, 66, 68, 70, 75, 77:78, 81:82, 85, 87:91), invert = TRUE)

Immune.Cells.True.subsetting <- Immune.Cells.True

Immune.Cells.TrueV2 <- FindNeighbors(Immune.Cells.True.subsetting, dims = 1:20)
Immune.Cells.TrueV2 <- FindClusters(Immune.Cells.TrueV2, resolution = 0.6)
Immune.Cells.TrueV2 <- RunUMAP(Immune.Cells.TrueV2, dims = 1:20)

table(Immune.Cells.TrueV2@meta.data$seurat_clusters)

clustree(Immune.Cells.TrueV2)

p1 <- DimPlot(Immune.Cells.TrueV2, reduction = "umap", label =  TRUE, raster = FALSE) + 
  theme_classic(base_size = 20) +
  ggtitle('UMAP of Immune cells TrueV2 subclusters Res 0.6')

p1

#removing levels that are empty (only keeping the values in the vectors I want) and getting the numb tisseue, n_patient, and n_cells
Immune.Cells.TrueV2@meta.data <- droplevels(Immune.Cells.TrueV2@meta.data)

Immune.Cells.TrueV2@meta.data %>% 
      group_by(seurat_clusters %>% as.character()) %>% 
     summarise(n_tissue=n_distinct(orig.ident), n_patient= n_distinct(Patient.No.), n=n()) %>%
      print(n=29)


#Immune markers to identify immune cells
Immune.markers = read.delim("data/Tcell_NKTcells_NKcells_Immune_Markers.csv", sep = ",")
head(Immune.markers)

#selected.features = unique(c(as.character(unique(Immune.markers$Positive)), as.character(unique(Immune.markers$Negative))))
selected.features = unique(c(as.character(unique(Immune.markers$Positive))))



p1 <-DotPlot(Immune.Cells.TrueV2, assay = "RNA", features = selected.features) + labs(title = "Identifying Immune cell TrueV2 Sub-cluster by Canonical Markers res0.6") +
 theme_classic(base_size = 20) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.8), axis.text.y = element_text(size = 9)) +
  coord_flip() +
  scale_colour_gradient2(low = "#2166ac", mid = "#f7f7f7", high = "#b2182b")


p1
```

#combining clusters into their immune families
```{r}
Immune.Cells.TrueV2@meta.data <- Immune.Cells.TrueV2@meta.data  %>%
                    mutate(Parent_immune_cluster = case_when( 
                             seurat_clusters %in% c(0,1,2,3,5,6,7,8,9,10,12,14,21,23) ~ "T cells", 
                             seurat_clusters %in% c(15,16) ~ "NK cells",
                             seurat_clusters %in% c(4,11,20,22) ~ "Myeloid cells", 
                             seurat_clusters == 19 ~ "Mast cells",
                             seurat_clusters == 13 ~ "B cells",
                             seurat_clusters %in% c(17,18) ~ "Plasma cells"))
```


```{r}
markers <- FindAllMarkers(Immune.Cells.TrueV2, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)

write_csv(markers, "data/Immune_Cells_TrueV2_DEG_list.csv")
```

#trying superCAT to see how cluster correlate together of the Immune cells true cluster
```{r}
markers = read.delim("data/Immune_Cells_TrueV2_DEG_list.csv", sep = ",")

ng = 100
p.v = 0.05
fc = 0
features.deg <- markers %>%
  filter(p_val_adj < p.v & avg_log2FC > fc) %>%
  group_by(cluster) %>%
  top_n(ng)

aux = DotPlot(Immune.Cells.TrueV2, features = unique(features.deg$gene))
aux.data = aux$data

data.avg.exp.scaled = data.frame(row.names = unique(features.deg$gene))

for (o in unique(aux.data$id) ) {
  aux.data.filt = aux.data[which(aux.data$id == o),]
  rownames(aux.data.filt) <- aux.data.filt$features.plot
  data.avg.exp.scaled = cbind(data.avg.exp.scaled, aux.data.filt[, 'avg.exp.scaled', drop=FALSE])
}

colnames(data.avg.exp.scaled) <- paste(unique(aux.data$id), " (n=", table(features.deg$cluster), ")", sep = "")
pheatmap(data.avg.exp.scaled, 
         show_rownames = F, 
         main = paste0("DEG expression signature clusters (n=", length(unique(features.deg$gene)), ")"), 
         cutree_cols = 10 )
```

```{r}
Immune.Cells.TrueV2@meta.data <- droplevels(Immune.Cells.TrueV2@meta.data)


#proportion by patient
counts.prop = data.frame(table(Immune.Cells.TrueV2@meta.data$Patient.No., Immune.Cells.TrueV2@meta.data$seurat_clusters))
counts.prop.perc = group_by(counts.prop, Var2) %>% mutate(percent = Freq/sum(Freq))

head(counts.prop.perc)

aux = aggregate(counts.prop$Freq, by=list(Category=counts.prop$Var1), FUN=sum)
aux$Var2 = "Patient"
colnames(aux) <- c("Var1", "Freq", "Var2")
aux$percent = aux$Freq / sum(aux$Freq)
counts.prop.perc = rbind(as.data.frame(counts.prop.perc), aux)

col = c("#a50026", "#d73027", "#f46d43", "#fdae61", "#fee090", "#ffffbf",
        "#00441b", "#1b7837", "#5aae61", "#a6dba0", "#d9f0d3", "#f7f7f7",
        "#e0f3f8", "#abd9e9", "#74add1", "#4575b4", "#313695",
        "#40004b","#762a83", "#9970ab", "#c2a5cf")

bsize = 20
ggplot(counts.prop.perc, aes(x = Var2, y = percent, fill = Var1)) + 
  geom_bar(position="stack",stat = "identity", width=0.8) +
  scale_fill_manual(name="Patient", values = c("1" = col[1],
                                               "2" = col[2],
                                               "3" = col[3],
                                               "4" = col[4],
                                               "5" = col[5],
                                               "6" = col[6],
                                               "7" = col[7],
                                               "8" = col[8],
                                               "9"= col[9],
                                               "10" = col[10],
                                               "11"= col[11],
                                               "12" = col[12],
                                               "13" = col[13],
                                               "14" = col[14],
                                               "15" = col[15],
                                               "16" = col[16],
                                               "17" = col[17],
                                               "18" = col[18],
                                               "19" = col[19],
                                               "20"= col[20],
                                               "21" = col[21])) +
  scale_x_discrete(limits = rev(levels(counts.prop.perc$Var2))) +
  coord_flip() +
  theme_set(theme_classic(base_size = bsize)) +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 270, hjust = 1),
        axis.text.y = element_text(angle = 0, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none") +
  ggtitle('Proportion of Immune cells TrueV2 by patient res0.6')
```

```{r}
#proportions by LA tissue activity
Immune.Cells.TrueV2@meta.data <- droplevels(Immune.Cells.TrueV2@meta.data)

counts.prop = data.frame(table(Immune.Cells.TrueV2@meta.data$LA_Status, Immune.Cells.TrueV2@meta.data$seurat_clusters))
counts.prop.perc = group_by(counts.prop, Var2) %>% mutate(percent = Freq/sum(Freq))
head(counts.prop.perc)
aux = aggregate(counts.prop$Freq, by=list(Category=counts.prop$Var1), FUN=sum)
aux$Var2 = "Total"
colnames(aux) <- c("Var1", "Freq", "Var2")
aux$percent = aux$Freq / sum(aux$Freq)
counts.prop.perc = rbind(as.data.frame(counts.prop.perc), aux)

bsize = 20
p1 <- ggplot(counts.prop.perc, aes(x = Var2, y = percent, fill = Var1)) + 
  geom_bar(position="stack",stat = "identity", width=0.8) +
  scale_fill_manual(name="LA Group", values = c("High" = "#fc8d59", 
                                                    "Low" = "#91bfdb",
                                                  "Exclude" = "#8b0000",
                                                  "NA" = "#2f4f4f")) +
  scale_x_discrete(limits = rev(levels(counts.prop.perc$Var2))) +
  coord_flip() +
  theme_set(theme_classic(base_size = 30)) +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 270, hjust = 1),
        axis.text.y = element_text(angle = 0, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle('Proportion of Immune cells TrueV2 res0.6 in LA Status') +
  theme_classic(base_size = 20)

p1
```


```{r}
#proportions by Major class
Immune.Cells.TrueV2@meta.data <- droplevels(Immune.Cells.TrueV2@meta.data)

counts.prop = data.frame(table(Immune.Cells.TrueV2@meta.data$Major.Class, Immune.Cells.TrueV2@meta.data$seurat_clusters))
counts.prop.perc = group_by(counts.prop, Var2) %>% mutate(percent = Freq/sum(Freq))
head(counts.prop.perc)
aux = aggregate(counts.prop$Freq, by=list(Category=counts.prop$Var1), FUN=sum)
aux$Var2 = "Total"
colnames(aux) <- c("Var1", "Freq", "Var2")
aux$percent = aux$Freq / sum(aux$Freq)
counts.prop.perc = rbind(as.data.frame(counts.prop.perc), aux)

bsize = 20
p1 <- ggplot(counts.prop.perc, aes(x = Var2, y = percent, fill = Var1)) + 
  geom_bar(position="stack",stat = "identity", width=0.8) +
  scale_fill_manual(name="Major Class", values = c("Endometrioma" = "#7b3294", 
                      "Eutopic Endometrium" = "#c2a5cf", 
                      "Endometriosis" = "#d9f0d3", 
                      "No endometriosis detected" = "#a6dba0",
                      "Unaffected ovary" = "#008837")) +
  scale_x_discrete(limits = rev(levels(counts.prop.perc$Var2))) +
  coord_flip() +
  theme_set(theme_classic(base_size = 30)) +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 270, hjust = 1),
        axis.text.y = element_text(angle = 0, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle('Proportion of Immune cells TrueV2 res0.6 in Major class') +
  theme_classic(base_size = 20)

p1
```


#pull out just peritoneal endometriosis samples
```{r}
Endometriosis <- subset(Immune.Cells.TrueV2, subset = Class %in% "Peritoneal endometriosis")


#proportions by LA tissue activity
Endometriosis@meta.data <- droplevels(Endometriosis@meta.data)

counts.prop = data.frame(table(Endometriosis@meta.data$LA_Status, Endometriosis@meta.data$seurat_clusters))
counts.prop.perc = group_by(counts.prop, Var2) %>% mutate(percent = Freq/sum(Freq))
head(counts.prop.perc)
aux = aggregate(counts.prop$Freq, by=list(Category=counts.prop$Var1), FUN=sum)
aux$Var2 = "Total"
colnames(aux) <- c("Var1", "Freq", "Var2")
aux$percent = aux$Freq / sum(aux$Freq)
counts.prop.perc = rbind(as.data.frame(counts.prop.perc), aux)

bsize = 20
p1 <- ggplot(counts.prop.perc, aes(x = Var2, y = percent, fill = Var1)) + 
  geom_bar(position="stack",stat = "identity", width=0.8) +
  scale_fill_manual(name="LA Group", values = c("High" = "#fc8d59", 
                                                    "Low" = "#91bfdb",
                                                  "Exclude" = "#8b0000",
                                                  "NA" = "#2f4f4f")) +
  scale_x_discrete(limits = rev(levels(counts.prop.perc$Var2))) +
  coord_flip() +
  theme_set(theme_classic(base_size = 30)) +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 270, hjust = 1),
        axis.text.y = element_text(angle = 0, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle('Proportion of Immune cells TrueV2 res0.6 in LA Status') +
  theme_classic(base_size = 20)

p1

Endometriosis@meta.data %>% 
      group_by(seurat_clusters %>% as.character()) %>% 
     summarise(n_tissue=n_distinct(orig.ident), n_patient= n_distinct(Patient.No.), n=n()) %>%
      print(n=18)

```





## Fig 2D
PCA based on the following cell proportion for each samples:

Follow the below steps for PCA analysis 

1. import Annotation tracker to add the LA_Status Epi_data tab - we need that col to label ggplot ( Master_Assigned_Cell_Type is already included in mate.data)
2. import `broom`package to parse PCA output from `prcomp`
3. Only meta.data from Seurat obj is needed for the PCA analysis
4. If the label on plot is too crowded. consider using `Patient.No.`

```{r}
#Endo_table <- readRDS("data/fixed_annotated_aux.seurat.rds")@meta.data
excel_path <- "data/Annotation tracker.xlsx"
sheets <- readxl::excel_sheets(excel_path)
Annotation_tracker <- sheets %>% 
                      lapply(. , function(s){ read_excel(excel_path, sheet = s) %>% 
                                              mutate(across(where(is.character),str_trim))
                                              }) %>%
                      setNames(sheets)

Annotation_tracker$Epi_Data 
```


```{r}
immue_endo_t<- Endometriosis@meta.data %>%
                  #filter(LA_Status == c("High", "Low")) %>%
                  #filter(Master_Assigned_Cell_Type) %>%
                  #filter(! is.na(Number_of_lymphoid_aggregates)) %>%
                  group_by(SampleName,Parent_immune_cluster) %>%
                  #group_by(SampleName,Master_Assigned_Cell_Type) %>%
                  summarise(count=n()) %>%
                  mutate(prop= count/sum(count)*100)

anno_df <- Endometriosis@meta.data  %>%
           rownames_to_column() %>%
           dplyr::select(-rowname) %>%
           #left_join(Annotation_tracker$Epi_Data, by=c("Patient.No."="Patient_ID")) %>%
           #dplyr::filter(LA_Status != "Exclude")  %>% 
           distinct(Patient.No.,LA_Status,SampleName) %>%
           column_to_rownames("SampleName")

heatmap_df <- immue_endo_t  %>% 
              dplyr::select(Parent_immune_cluster,prop,SampleName) %>%
              spread(Parent_immune_cluster, prop) %>%
              column_to_rownames("SampleName") %>%
              as.matrix() %>%
              replace(.,is.na(.),0) %>%
              t()
write_csv(as.data.frame(heatmap_df),"data/heatmap_df.csv")

pc <- heatmap_df %>% 
t %>%
prcomp(scale = TRUE) 

## Variance Explain in each PC

pc %>%
 tidy("pcs") %>%
  ggplot(aes(x=PC,y=percent*100)) +
    geom_bar(stat="identity") +
    labs(y="Percent Variance Explained") +
    theme_classic()

# A tricky one to show - how much each clusters contribute to each PC (only show largest contributors)
var_explain <- varimax(pc$rotation)


pc %>%
augment(.) %>%
inner_join(anno_df %>% rownames_to_column(".rownames"), by=".rownames") %>%
  ggplot(aes(x=.fittedPC4, y=.fittedPC5,color=LA_Status,label=Patient.No.))+
  geom_point(size=5) +
  geom_text(nudge_x = 0.2, nudge_y = 0.2,angle=20,size=3) +
  labs(x="PC4",y="PC5") + 
  theme_half_open(12) 


```

#Sanky plot to compare resolution of 0.4 from previous analysis when pulling out Immune cells in LA group to resolution of 0.6 in current analysis

```{r}

patient_la_anno <- read_csv("data/endo_LA_Summary.csv",col_types = cols(patient_id = col_integer()))

Immune.Cells@meta.data <- Immune.Cells@meta.data  %>%
              rownames_to_column() %>%
              left_join(patient_la_anno %>% mutate(patient_id = as.integer(patient_id)), by=c("Patient.No."="patient_id")) %>%
  column_to_rownames()


#Pull out Immune cells from LA groups in tissue samples
Immune.Cells.LA <- subset(Immune.Cells, subset = immune_activity %in% c("High", "Low"))

p1 <- DimPlot(Immune.Cells.LA, pt.size = 0.1, group.by = "Master_Assigned_Cell_Type", label = T, raster=FALSE) +
  theme_classic(base_size = 20)

p1

pdf(file = "Plots/Endometriosis_LA_Immune/Immune Cell Clusters in LA Tissue Group.pdf", width = 15, height = 10) 
p1
dev.off()



#removing levels that are empty (only keeping the values in the vectors I want) and getting the numb tisseue, n_patient, and n_cells
Immune.Cells.LA@meta.data <- droplevels(Immune.Cells.LA@meta.data)

Immune.Cells.LA@meta.data %>% 
     group_by(Master_Assigned_Cell_Type) %>% 
     summarise(n_tissue=n_distinct(orig.ident), n_patient= n_distinct(Patient.No.), n=n())


```


```{r}
##Subclustering just the Immune cells from LA high low
Immune.cells.subclustering.res0.3 <- FindNeighbors(Immune.Cells.LA, dims = 1:20)
Immune.cells.subclustering.res0.3 <- FindClusters(Immune.cells.subclustering.res0.3, resolution = 0.3)
Immune.cells.subclustering.res0.3 <- RunUMAP(Immune.cells.subclustering.res0.3, dims = 1:20)

p1 <- DimPlot(Immune.cells.subclustering.res0.3, reduction = "umap", label =  TRUE) + 
  theme_classic(base_size = 20) +
  ggtitle('UMAP of LA tissue group Immune cells subclusters Res 0.3')

p1

p1 <-DotPlot(Immune.cells.subclustering.res0.3, assay = "RNA", features = selected.features) + labs(title = "Identifying LA tisue groups Immune cell Sub-cluster by Canonical Markers res0.3") +
 theme_classic(base_size = 20) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.8), axis.text.y = element_text(size = 9)) +
  coord_flip() +
  scale_colour_gradient2(low = "#2166ac", mid = "#f7f7f7", high = "#b2182b")

p1

```


```{r}
#Pull out True Immune cells.
Immune.Cells.True.res0.3 <- subset(Immune.cells.subclustering.res0.3, subset = seurat_clusters %in% c(5, 11, 12, 13, 17), invert = TRUE)

Immune.Cells.True.subsetting.LAgroup <- Immune.Cells.True.res0.3

Immune.Cells.TrueV2.LAgroup <- FindNeighbors(Immune.Cells.True.subsetting.LAgroup, dims = 1:20)
Immune.Cells.TrueV2.LAgroup <- FindClusters(Immune.Cells.TrueV2.LAgroup, resolution = 0.4)
Immune.Cells.TrueV2.LAgroup <- RunUMAP(Immune.Cells.TrueV2.LAgroup, dims = 1:20)

table(Immune.Cells.TrueV2.LAgroup@meta.data$seurat_clusters)


p1 <- DimPlot(Immune.Cells.TrueV2.LAgroup, reduction = "umap", label =  TRUE) + 
  theme_classic(base_size = 20) +
  ggtitle('UMAP of only True LA Tissue Group Immune cells subclusters Res0.4')

p1

p1 <- DotPlot(Immune.Cells.TrueV2.LAgroup, assay = "RNA", features = selected.features) + labs(title = "Identifying Immune cell cluster by Canonical Markers res2") +
  theme_classic(base_size = 20) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.8, size = 8), axis.text.y = element_text(size = 8)) +
  coord_flip() +
  scale_colour_gradient2(low = "#2166ac", mid = "#f7f7f7", high = "#b2182b")

p1

Endometriosis.LAgroup <- subset(Immune.Cells.TrueV2.LAgroup, subset = Class %in% "Peritoneal endometriosis")

Endometriosis.LAgroup@meta.data %>% 
      group_by(seurat_clusters %>% as.character()) %>% 
     summarise(n_tissue=n_distinct(orig.ident), n_patient= n_distinct(Patient.No.), n=n()) %>%
      print(n=18)

```



```{r}
#Sankey
Endometriosis.LAgroup@meta.data %>%
rownames_to_column(var = "cell") %>%
dplyr::select(cell,seurat_clusters) %>%
inner_join( Endometriosis@meta.data %>%
            rownames_to_column(var = "cell") %>% 
            dplyr::select(cell,seurat_clusters)
            , by = "cell") %>%
  group_by( seurat_clusters.x, seurat_clusters.y)%>%
  summarise(n=n())%>%
  spread(seurat_clusters.y, seurat_clusters.x) 
#View
  
data = as.data.frame.matrix(table(Endometriosis.LAgroup@meta.data$seurat_clusters, Endometriosis@meta.data$seurat_clusters))
data
library(networkD3)
library(dplyr)
library(tidyverse)
# I need a long format
data.long.class <- data %>%
  rownames_to_column %>%
  gather(key = 'key', value = 'value', -rowname) %>%
  filter(value > 0)
colnames(data.long.class) <- c("source", "target", "value")
data.long.class$target <- paste(data.long.class$target, " ", sep="")
# From these flows we need to create a node data frame: it lists every entities involved in the flow
nodes <- data.frame(name=c(as.character(data.long.class$source), as.character(data.long.class$target)) %>% unique())
# With networkD3, cnodes <- data.frame(name=c(as.character(data.long.class$source), as.character(data.long.class$target)) %>% unique())
#onnection must be provided using id, not using real name like in the links dataframe.. So we need to reformat it.
data.long.class$IDsource=match(data.long.class$source, nodes$name)-1
data.long.class$IDtarget=match(data.long.class$target, nodes$name)-1
# prepare colour scale
ColourScal ='d3.scaleOrdinal() .range(["#FDE725FF","#B4DE2CFF","#6DCD59FF","#35B779FF","#1F9E89FF","#26828EFF","#31688EFF","#3E4A89FF","#482878FF","#440154FF"])'
sankeyNetwork(Links = data.long.class, Nodes = nodes,
              Source = "IDsource", Target = "IDtarget",
              Value = "value", NodeID = "name",
              sinksRight=T, colourScale=ColourScal, nodeWidth=40, fontSize=20, nodePadding=20)
```

