---
title: "Figure 2"
author: "Marcela Haro"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---
# Figure 2 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
getwd()
```




## Library loading 

```{r ,message = FALSE}

library(Seurat)
library(tidyverse)
library(cowplot)
library(broom)
library(ggplot2)
library(dplyr)
library(reshape)
library(clusterProfiler)
library(readxl)
library(ggrepel)
library(clustree)
library(pheatmap)
```


#Loading the scRNAseq data and loading into seurat object
```{r}
#reading the Seurat object RDS file
d10x.seurat <- readRDS("/media/Data01/abraaodem/Code/AtlasEndometriosis/rds/aux.seurat.rds")
```


```{r}
##To confirm annotation
table(d10x.seurat@meta.data$active.cluster)
```



#adding more columns to meta data: based on pathology LA status classified patients as LA high or low, also have a colum for each tissue to say if it has any LAs(yes/no)
```{r}
#adding more columns to metadata about the LA immune activity at patient leve (High/Low); this is Yes/No LA activity for each tissue specimen (Yes/No).

##Note tissue activity need to be updated in the annoatation tracker file. Just use the immune_activity column that is coming from the patient_la_anno

excel_path <- "data/Annotation tracker.xlsx"
sheets <- readxl::excel_sheets(excel_path)
Annotation_tracker <- sheets %>% 
                      lapply(. , function(s){ read_excel(excel_path, sheet = s) %>%
                                              mutate(across(everything(),as.factor))}) %>%
                      setNames(sheets)
```

```{r}
Annotation_tracker$Clusters <- Annotation_tracker$Clusters %>% 
                            mutate(Master_Assigned_Cell_Type= if_else(is.na(Updated_Anno),Master_Assigned_Cell_Type,Updated_Anno) ) %>% 
                            dplyr::select(-Updated_Anno)

back_up_meta_table <- d10x.seurat@meta.data 

d10x.seurat@meta.data <- back_up_meta_table  %>%
              rownames_to_column() %>%
              left_join(Annotation_tracker$Clusters,by=c("SCT_snn_res.3"="Cluster_Number")) %>%
              left_join(Annotation_tracker$tissue_sample_meta, by=c("Anatomic.annotation" = "Anatomic.annotation","orig.ident"= "Orig.ident" )) %>%
              left_join(Annotation_tracker$Epi_Data %>% mutate(Patient_ID = as.integer(Patient_ID)), by=c("Patient.No."="Patient_ID")) %>%
  column_to_rownames()
```


```{r}
##To confirm annotation
table(d10x.seurat@meta.data$Master_Assigned_Cell_Type)
```



```{r}
#UMAPS of all the cells

#DimPlot(d10x.seurat, reduction = "umap", label =FALSE, raster = FALSE) ##This is a umap with all the clusters

cluster.cols = c("#67001f", "#a50026", "#d73027", "#f46d43", "#fdae61", "#fee090", "#ffffbf", "#a6dba0", "#1a9850", "#e0f3f8", "#abd9e9")

p1 <- DimPlot(object = d10x.seurat, pt.size = 0.1, group.by ="Master_Assigned_Cell_Type", cols = cluster.cols, label = T, raster=FALSE) +
  theme_classic(base_size = 20) ##this is a umap with the clusters grouped with their cell identity

p1

#pdf(file = "Plots/Endometriosis_LA_Immune/Original Cell Clusters.pdf", width = 15, height = 15) 
#p1
#dev.off()
```



#pull out Immune cells from the data set
```{r}
###pull out just the Immune cells
Immune.Cells <- subset(d10x.seurat, subset = Master_Assigned_Cell_Type %in% c("B/Plasma cells", "T/NK cells", "Myeloid cells", "Mast cells"))
```


```{r}
Immune.Cells.Subset <- Immune.Cells

##Subclustering just the Immune cells from LA high low
Immune.cells.subclustering <- FindNeighbors(Immune.Cells.Subset, dims = 1:20)
Immune.cells.subclustering <- FindClusters(Immune.cells.subclustering, resolution = 1)
Immune.cells.subclustering <- RunUMAP(Immune.cells.subclustering, dims = 1:20)

p1 <- DimPlot(Immune.cells.subclustering, reduction = "umap", label =  TRUE, raster = FALSE) + 
  theme_classic(base_size = 20) +
  ggtitle('UMAP of Immune cells subclusters Res 1')

p1


#removing levels that are empty (only keeping the values in the vectors I want) and getting the numb tisseue, n_patient, and n_cells
Immune.cells.subclustering@meta.data <- droplevels(Immune.cells.subclustering@meta.data)

Immune.cells.subclustering@meta.data %>% 
      group_by(seurat_clusters %>% as.character()) %>% 
     summarise(n_tissue=n_distinct(orig.ident), n_patient= n_distinct(Patient.No.), n=n()) %>%
      print(n=43)
```

```{r}
#Immune markers to identify immune cells
Immune.markers = read.delim("data/Immune_Marker.csv", sep = ",")
head(Immune.markers)

#selected.features = unique(c(as.character(unique(Immune.markers$Positive)), as.character(unique(Immune.markers$Negative))))
selected.features = unique(c(as.character(unique(Immune.markers$Positive))))
```

```{r}
p1 <-DotPlot(Immune.cells.subclustering, assay = "RNA", features = selected.features) + labs(title = "Identifying Immune cell Sub-cluster by Canonical Markers res1") +
 theme_classic(base_size = 20) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.8), axis.text.y = element_text(size = 9)) +
  coord_flip() +
  scale_colour_gradient2(low = "#2166ac", mid = "#f7f7f7", high = "#b2182b")

p1
```


```{r}
Immune.subcluster.DEG <- FindAllMarkers(Immune.cells.subclustering, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)

write_csv(Immune.subcluster.DEG, "data/Immune_subcluster_DEG_list.csv")
```


#trying superCAT to see how cluster correlate together
```{r}
markers = read.delim("data/Immune_subcluster_DEG_list.csv", sep = ",")

ng = 100
p.v = 0.05
fc = 0
features.deg <- markers %>%
  filter(p_val_adj < p.v & avg_log2FC > fc) %>%
  group_by(cluster) %>%
  top_n(ng)

aux = DotPlot(Immune.cells.subclustering, features = unique(features.deg$gene))
aux.data = aux$data

data.avg.exp.scaled = data.frame(row.names = unique(features.deg$gene))

for (o in unique(aux.data$id) ) {
  aux.data.filt = aux.data[which(aux.data$id == o),]
  rownames(aux.data.filt) <- aux.data.filt$features.plot
  data.avg.exp.scaled = cbind(data.avg.exp.scaled, aux.data.filt[, 'avg.exp.scaled', drop=FALSE])
}

colnames(data.avg.exp.scaled) <- paste(unique(aux.data$id), " (n=", table(features.deg$cluster), ")", sep = "")
pheatmap(data.avg.exp.scaled, 
         show_rownames = F, 
         main = paste0("DEG expression signature clusters (n=", length(unique(features.deg$gene)), ")"), 
         cutree_cols = 10 )
```


#remove non-Immune cell clusters
```{r}
#Pull out True Immune cells.
Immune.Cells.True <- subset(Immune.cells.subclustering, subset = seurat_clusters %in% c(11, 17, 18, 22:24, 26, 31:35, 37, 38, 40:42), invert = TRUE)

Immune.Cells.True.subsetting <- Immune.Cells.True

Immune.Cells.TrueV2 <- FindNeighbors(Immune.Cells.True.subsetting, dims = 1:20)
Immune.Cells.TrueV2 <- FindClusters(Immune.Cells.TrueV2, resolution = 1)
Immune.Cells.TrueV2 <- RunUMAP(Immune.Cells.TrueV2, dims = 1:20)

table(Immune.Cells.TrueV2@meta.data$seurat_clusters)

clustree(Immune.Cells.TrueV2)

p1 <- DimPlot(Immune.Cells.TrueV2, reduction = "umap", label =  TRUE, raster = FALSE) + 
  theme_classic(base_size = 20) +
  ggtitle('UMAP of Immune cells TrueV2 subclusters Res 1')

p1

#removing levels that are empty (only keeping the values in the vectors I want) and getting the numb tisseue, n_patient, and n_cells
Immune.Cells.TrueV2@meta.data <- droplevels(Immune.Cells.TrueV2@meta.data)

Immune.Cells.TrueV2@meta.data %>% 
      group_by(seurat_clusters %>% as.character()) %>% 
     summarise(n_tissue=n_distinct(orig.ident), n_patient= n_distinct(Patient.No.), n=n()) %>%
      print(n=35)

p1 <-DotPlot(Immune.Cells.TrueV2, assay = "RNA", features = selected.features) + labs(title = "Identifying Immune cell TrueV2 Sub-cluster by Canonical Markers res1") +
 theme_classic(base_size = 20) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.8), axis.text.y = element_text(size = 9)) +
  coord_flip() +
  scale_colour_gradient2(low = "#2166ac", mid = "#f7f7f7", high = "#b2182b")

p1
```



```{r}
Immune.Cells.TrueV2.DEG <- FindAllMarkers(Immune.Cells.TrueV2, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)

write_csv(Immune.Cells.TrueV2.DEG, "data/Immune_Cells_TrueV2_DEG_list.csv")
```

#trying superCAT to see how cluster correlate together of the Immune cells true cluster
```{r}
markers = read.delim("data/Immune_Cells_TrueV2_DEG_list.csv", sep = ",")

ng = 100
p.v = 0.05
fc = 0
features.deg <- markers %>%
  filter(p_val_adj < p.v & avg_log2FC > fc) %>%
  group_by(cluster) %>%
  top_n(ng)

aux = DotPlot(Immune.Cells.TrueV2, features = unique(features.deg$gene))
aux.data = aux$data

data.avg.exp.scaled = data.frame(row.names = unique(features.deg$gene))

for (o in unique(aux.data$id) ) {
  aux.data.filt = aux.data[which(aux.data$id == o),]
  rownames(aux.data.filt) <- aux.data.filt$features.plot
  data.avg.exp.scaled = cbind(data.avg.exp.scaled, aux.data.filt[, 'avg.exp.scaled', drop=FALSE])
}

colnames(data.avg.exp.scaled) <- paste(unique(aux.data$id), " (n=", table(features.deg$cluster), ")", sep = "")
pheatmap(data.avg.exp.scaled, 
         show_rownames = F, 
         main = paste0("DEG expression signature clusters (n=", length(unique(features.deg$gene)), ")"), 
         cutree_cols = 10 )
```

```{r}
Immune.Cells.TrueV2@meta.data <- droplevels(Immune.Cells.TrueV2@meta.data)


#proportion by patient
counts.prop = data.frame(table(Immune.Cells.TrueV2@meta.data$Patient.No., Immune.Cells.TrueV2@meta.data$seurat_clusters))
counts.prop.perc = group_by(counts.prop, Var2) %>% mutate(percent = Freq/sum(Freq))

head(counts.prop.perc)

aux = aggregate(counts.prop$Freq, by=list(Category=counts.prop$Var1), FUN=sum)
aux$Var2 = "Patient"
colnames(aux) <- c("Var1", "Freq", "Var2")
aux$percent = aux$Freq / sum(aux$Freq)
counts.prop.perc = rbind(as.data.frame(counts.prop.perc), aux)

col = c("#a50026", "#d73027", "#f46d43", "#fdae61", "#fee090", "#ffffbf",
        "#00441b", "#1b7837", "#5aae61", "#a6dba0", "#d9f0d3", "#f7f7f7",
        "#e0f3f8", "#abd9e9", "#74add1", "#4575b4", "#313695",
        "#40004b","#762a83", "#9970ab", "#c2a5cf")

bsize = 20
ggplot(counts.prop.perc, aes(x = Var2, y = percent, fill = Var1)) + 
  geom_bar(position="stack",stat = "identity", width=0.8) +
  scale_fill_manual(name="Patient", values = c("1" = col[1],
                                               "2" = col[2],
                                               "3" = col[3],
                                               "4" = col[4],
                                               "5" = col[5],
                                               "6" = col[6],
                                               "7" = col[7],
                                               "8" = col[8],
                                               "9"= col[9],
                                               "10" = col[10],
                                               "11"= col[11],
                                               "12" = col[12],
                                               "13" = col[13],
                                               "14" = col[14],
                                               "15" = col[15],
                                               "16" = col[16],
                                               "17" = col[17],
                                               "18" = col[18],
                                               "19" = col[19],
                                               "20"= col[20],
                                               "21" = col[21])) +
  scale_x_discrete(limits = rev(levels(counts.prop.perc$Var2))) +
  coord_flip() +
  theme_set(theme_classic(base_size = bsize)) +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 270, hjust = 1),
        axis.text.y = element_text(angle = 0, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none") +
  ggtitle('Proportion of Immune cells TrueV2 by patient res0.4')
```

```{r}
#proportions by LA tissue activity
Immune.Cells.TrueV2@meta.data <- droplevels(Immune.Cells.TrueV2@meta.data)

counts.prop = data.frame(table(Immune.Cells.TrueV2@meta.data$LA_Status, Immune.Cells.TrueV2@meta.data$seurat_clusters))
counts.prop.perc = group_by(counts.prop, Var2) %>% mutate(percent = Freq/sum(Freq))
head(counts.prop.perc)
aux = aggregate(counts.prop$Freq, by=list(Category=counts.prop$Var1), FUN=sum)
aux$Var2 = "Total"
colnames(aux) <- c("Var1", "Freq", "Var2")
aux$percent = aux$Freq / sum(aux$Freq)
counts.prop.perc = rbind(as.data.frame(counts.prop.perc), aux)

bsize = 20
p1 <- ggplot(counts.prop.perc, aes(x = Var2, y = percent, fill = Var1)) + 
  geom_bar(position="stack",stat = "identity", width=0.8) +
  scale_fill_manual(name="LA Group", values = c("High" = "#fc8d59", 
                                                    "Low" = "#91bfdb",
                                                  "Exclude" = "#8b0000",
                                                  "NA" = "#2f4f4f")) +
  scale_x_discrete(limits = rev(levels(counts.prop.perc$Var2))) +
  coord_flip() +
  theme_set(theme_classic(base_size = 30)) +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 270, hjust = 1),
        axis.text.y = element_text(angle = 0, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle('Proportion of Immune cells TrueV2 res0.4 in LA Status') +
  theme_classic(base_size = 20)

p1
```


```{r}
#proportions by Major class
Immune.Cells.TrueV2@meta.data <- droplevels(Immune.Cells.TrueV2@meta.data)

counts.prop = data.frame(table(Immune.Cells.TrueV2@meta.data$Major.Class, Immune.Cells.TrueV2@meta.data$seurat_clusters))
counts.prop.perc = group_by(counts.prop, Var2) %>% mutate(percent = Freq/sum(Freq))
head(counts.prop.perc)
aux = aggregate(counts.prop$Freq, by=list(Category=counts.prop$Var1), FUN=sum)
aux$Var2 = "Total"
colnames(aux) <- c("Var1", "Freq", "Var2")
aux$percent = aux$Freq / sum(aux$Freq)
counts.prop.perc = rbind(as.data.frame(counts.prop.perc), aux)

bsize = 20
p1 <- ggplot(counts.prop.perc, aes(x = Var2, y = percent, fill = Var1)) + 
  geom_bar(position="stack",stat = "identity", width=0.8) +
  scale_fill_manual(name="Major Class", values = c("Endometrioma" = "#7b3294", 
                      "Eutopic Endometrium" = "#c2a5cf", 
                      "Endometriosis" = "#d9f0d3", 
                      "No endometriosis detected" = "#a6dba0",
                      "Unaffected ovary" = "#008837")) +
  scale_x_discrete(limits = rev(levels(counts.prop.perc$Var2))) +
  coord_flip() +
  theme_set(theme_classic(base_size = 30)) +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 270, hjust = 1),
        axis.text.y = element_text(angle = 0, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle('Proportion of Immune cells TrueV2 res0.4 in Major class') +
  theme_classic(base_size = 20)

p1
```


#pull out just peritoneal endometriosis samples
```{r}
Endometriosis <- subset(Immune.Cells.TrueV2, subset = Class %in% "Peritoneal endometriosis")

#proportions by LA tissue activity
Endometriosis@meta.data <- droplevels(Endometriosis@meta.data)

counts.prop = data.frame(table(Endometriosis@meta.data$LA_Status, Endometriosis@meta.data$seurat_clusters))
counts.prop.perc = group_by(counts.prop, Var2) %>% mutate(percent = Freq/sum(Freq))
head(counts.prop.perc)
aux = aggregate(counts.prop$Freq, by=list(Category=counts.prop$Var1), FUN=sum)
aux$Var2 = "Total"
colnames(aux) <- c("Var1", "Freq", "Var2")
aux$percent = aux$Freq / sum(aux$Freq)
counts.prop.perc = rbind(as.data.frame(counts.prop.perc), aux)

bsize = 20
p1 <- ggplot(counts.prop.perc, aes(x = Var2, y = percent, fill = Var1)) + 
  geom_bar(position="stack",stat = "identity", width=0.8) +
  scale_fill_manual(name="LA Group", values = c("High" = "#fc8d59", 
                                                    "Low" = "#91bfdb",
                                                  "Exclude" = "#8b0000",
                                                  "NA" = "#2f4f4f")) +
  scale_x_discrete(limits = rev(levels(counts.prop.perc$Var2))) +
  coord_flip() +
  theme_set(theme_classic(base_size = 30)) +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 270, hjust = 1),
        axis.text.y = element_text(angle = 0, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle('Proportion of Immune cells TrueV2 res0.4 in LA Status') +
  theme_classic(base_size = 20)

p1

Endometriosis@meta.data %>% 
      group_by(seurat_clusters %>% as.character()) %>% 
     summarise(n_tissue=n_distinct(orig.ident), n_patient= n_distinct(Patient.No.), n=n()) %>%
      print(n=18)

```



#NOt going to use below ananlysis so skip

#lets see what happens if we pull out the B-cell clusters
```{r}
B.Cells <- subset(Immune.Cells.TrueV2, subset = seurat_clusters %in% c(8, 13))


#subcluster B-cells
B.cells.subclustering <- FindNeighbors(B.Cells, dims = 1:20)
B.cells.subclustering <- FindClusters(B.cells.subclustering, resolution = 0.3)
B.cells.subclustering <- RunUMAP(B.cells.subclustering, dims = 1:20)

p1 <- DimPlot(B.cells.subclustering, reduction = "umap", label =  TRUE, raster = FALSE) + 
  theme_classic(base_size = 20) +
  ggtitle('UMAP of B-cells subclusters Res 0.3')

p1

#removing levels that are empty (only keeping the values in the vectors I want) and getting the numb tisseue, n_patient, and n_cells
B.cells.subclustering@meta.data <- droplevels(B.cells.subclustering@meta.data)

B.cells.subclustering@meta.data %>% 
      group_by(seurat_clusters %>% as.character()) %>% 
     summarise(n_tissue=n_distinct(orig.ident), n_patient= n_distinct(Patient.No.), n=n()) %>%
      print(n=17)

#Immune markers to identify B-cells
Bcells.markers = read.delim("data/Bcells_PlasmaCell_Markers.csv", sep = ",")
head(Bcells.markers)

selected.features = unique(c(as.character(unique(Bcells.markers$Positive))))

p1 <-DotPlot(B.cells.subclustering, assay = "RNA", features = selected.features) + labs(title = "Identifying Immune cell TrueV2 Sub-cluster by Canonical Markers res0.3") +
 theme_classic(base_size = 20) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.8), axis.text.y = element_text(size = 9)) +
  coord_flip() +
  scale_colour_gradient2(low = "#2166ac", mid = "#f7f7f7", high = "#b2182b")

p1

```


#remove non-Bcell cluster

```{r}
B.Cells.True <- subset(B.cells.subclustering, subset = seurat_clusters %in% c(10), invert=TRUE)

B.cell.subset <- B.Cells.True


#subcluster B-cells
B.cells.True.subclustering <- FindNeighbors(B.cell.subset, dims = 1:20)
B.cells.True.subclustering <- FindClusters(B.cells.True.subclustering, resolution = 0.5)
B.cells.True.subclustering <- RunUMAP(B.cells.True.subclustering, dims = 1:20)

p1 <- DimPlot(B.cells.True.subclustering, reduction = "umap", label =  TRUE, raster = FALSE) + 
  theme_classic(base_size = 20) +
  ggtitle('UMAP of B-cells True subclusters Res 0.5')

p1

#removing levels that are empty (only keeping the values in the vectors I want) and getting the numb tisseue, n_patient, and n_cells
B.cells.True.subclustering@meta.data <- droplevels(B.cells.True.subclustering@meta.data)

B.cells.True.subclustering@meta.data %>% 
      group_by(seurat_clusters %>% as.character()) %>% 
     summarise(n_tissue=n_distinct(orig.ident), n_patient= n_distinct(Patient.No.), n=n()) %>%
      print(n=17)

#Immune markers to identify B-cells
Bcells.markers = read.delim("data/Bcells_PlasmaCell_Markers.csv", sep = ",")
head(Bcells.markers)

selected.features = unique(c(as.character(unique(Bcells.markers$Positive))))

p1 <-DotPlot(B.cells.True.subclustering, assay = "RNA", features = selected.features) + labs(title = "Identifying B-cells true Sub-cluster by Canonical Markers res0.5") +
 theme_classic(base_size = 20) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.8), axis.text.y = element_text(size = 9)) +
  coord_flip() +
  scale_colour_gradient2(low = "#2166ac", mid = "#f7f7f7", high = "#b2182b")

p1


clustree(B.cells.True.subclustering)
```


```{r}
B.cells.True.subclustering@meta.data <- droplevels(B.cells.True.subclustering@meta.data)


#proportion by patient
counts.prop = data.frame(table(B.cells.True.subclustering@meta.data$Patient.No., B.cells.True.subclustering@meta.data$seurat_clusters))
counts.prop.perc = group_by(counts.prop, Var2) %>% mutate(percent = Freq/sum(Freq))

head(counts.prop.perc)

aux = aggregate(counts.prop$Freq, by=list(Category=counts.prop$Var1), FUN=sum)
aux$Var2 = "Patient"
colnames(aux) <- c("Var1", "Freq", "Var2")
aux$percent = aux$Freq / sum(aux$Freq)
counts.prop.perc = rbind(as.data.frame(counts.prop.perc), aux)

col = c("#a50026", "#d73027", "#f46d43", "#fdae61", "#fee090", "#ffffbf",
        "#00441b", "#1b7837", "#5aae61", "#a6dba0", "#d9f0d3", "#f7f7f7",
        "#e0f3f8", "#abd9e9", "#74add1", "#4575b4", "#313695",
        "#40004b","#762a83", "#9970ab", "#c2a5cf")

bsize = 20
ggplot(counts.prop.perc, aes(x = Var2, y = percent, fill = Var1)) + 
  geom_bar(position="stack",stat = "identity", width=0.8) +
  scale_fill_manual(name="Patient", values = c("1" = col[1],
                                               "2" = col[2],
                                               "3" = col[3],
                                               "4" = col[4],
                                               "5" = col[5],
                                               "6" = col[6],
                                               "7" = col[7],
                                               "8" = col[8],
                                               "9"= col[9],
                                               "10" = col[10],
                                               "11"= col[11],
                                               "12" = col[12],
                                               "13" = col[13],
                                               "14" = col[14],
                                               "15" = col[15],
                                               "16" = col[16],
                                               "17" = col[17],
                                               "18" = col[18],
                                               "19" = col[19],
                                               "20"= col[20],
                                               "21" = col[21])) +
  scale_x_discrete(limits = rev(levels(counts.prop.perc$Var2))) +
  coord_flip() +
  theme_set(theme_gray(base_size = bsize)) +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 270, hjust = 1),
        axis.text.y = element_text(angle = 0, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none") +
  ggtitle('Proportion of B-cells True by patient res0.5')
```


```{r}
#proportions by LA tissue activity
B.cells.True.subclustering@meta.data <- droplevels(B.cells.True.subclustering@meta.data)

counts.prop = data.frame(table(B.cells.True.subclustering@meta.data$LA_Status, B.cells.True.subclustering@meta.data$seurat_clusters))
counts.prop.perc = group_by(counts.prop, Var2) %>% mutate(percent = Freq/sum(Freq))
head(counts.prop.perc)
aux = aggregate(counts.prop$Freq, by=list(Category=counts.prop$Var1), FUN=sum)
aux$Var2 = "Total"
colnames(aux) <- c("Var1", "Freq", "Var2")
aux$percent = aux$Freq / sum(aux$Freq)
counts.prop.perc = rbind(as.data.frame(counts.prop.perc), aux)

bsize = 20
p1 <- ggplot(counts.prop.perc, aes(x = Var2, y = percent, fill = Var1)) + 
  geom_bar(position="stack",stat = "identity", width=0.8) +
  scale_fill_manual(name="LA Group", values = c("High" = "#fc8d59", 
                                                    "Low" = "#91bfdb")) +
  scale_x_discrete(limits = rev(levels(counts.prop.perc$Var2))) +
  coord_flip() +
  theme_set(theme_classic(base_size = 30)) +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 270, hjust = 1),
        axis.text.y = element_text(angle = 0, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle('Proportion of Immune cells in LA tissue activity') +
  theme_classic(base_size = 20)

p1
```

