---
title: "Dataset introduction"
---


```{r}
library(tidyverse)
library(cowplot)
library(ggplot2)
library(ComplexHeatmap)
library(circlize)
library(ggpubr)

files<-list.files()
files<-files[str_detect(files,"bgclus")]
files<-files[-1]
metadata<-lapply(files,data.table::fread)%>%bind_rows()

domcol <- c( "gray13", "gold2", "plum4", "darkorange1", "lightskyblue2", "firebrick", "burlywood3", "gray51", "springgreen4", "lightpink2", "deepskyblue4", "lightsalmon2", "mediumpurple4", "orange", "maroon", "yellow3", "brown4", "yellow4", "sienna4", "chocolate")
names(domcol)<-1:20
slidecol<-sample(domcol,length(unique(metadata$Run_Tissue_name)))
names(slidecol)<-unique(metadata$Run_Tissue_name)
classcols = c( "Eutopic endometrium" = "#c2a5cf", 
              "Peritoneal endometriosis" = "#d9f0d3")
cellcols<-c("#8DB600","#fdae61","#a95f96","#0067A5","#604E97",
            "#f46d43","hotpink","#BE0032","cornflowerblue","black")
names(cellcols)<-c("B cells","Plasma cells","CD4 T cells","CD8 T cells",
                   "NK cells","Myeloid cells","unknown","Epithelial","Stroma","low_signal")

```
# Filter/QC
```{r}
temp<-metadata%>%group_by(Run_Tissue_name)%>%
  summarise(thres=quantile(nCount_RNA,c(0.05)))
metadata<-metadata%>%left_join(temp)
metadata<-metadata%>%
  mutate(qc_type=ifelse(nCount_RNA>thres,
                        cl_corrected,"low_signal"))

temp<-metadata%>%count(Run_Tissue_name,fov,qc_type)%>%
  group_by(Run_Tissue_name,fov)%>%mutate(prop=100*n/sum(n))%>%
  filter(qc_type=="low_signal")
i<-temp%>%filter(prop>=25)%>%select(1:3)
colnames(i)[3]<-"fov_type"
metadata<-metadata%>%left_join(i)
metadata<-metadata%>%
  mutate(fov_type=ifelse(is.na(fov_type),qc_type,fov_type))%>%
  select(-qc_type)

metadata<-metadata%>%filter(!fov_type%in%c("low_signal","unknown"))
```
# Cell type markers
```{r}
markers<-read_tsv("cosmx_markers.tsv")
markers<-markers%>%separate_rows(Positive,sep = ",")%>%
  distinct(Positive)%>%unlist()
markers<-markers[c(2,1,5,4,8,7,9,12,10,16,18,3,19,22,23)]

pe<-vroom::vroom("/media/Lawrenson_Lab_NAS/uthscsa/group_data/CosMx_temp/PE.csv")%>%
  filter(cell_ID%in%metadata$cell_ID)
mat<-vroom::vroom("/media/Lawrenson_Lab_NAS/uthscsa/group_data/CosMx_temp/endometrium.csv")%>%filter(cell_ID%in%metadata$cell_ID)
mat<-mat%>%bind_rows(pe)%>%column_to_rownames("cell_ID")%>%t()
negavg<-colMeans(mat[c("Ms IgG1","Rb IgG"),])
normi<-t(t(mat)-negavg)
normi<-asinh(mat/5)
normi<-sva::ComBat(dat = normi,batch = str_remove(colnames(normi),"c_.+"),
                     BPPARAM = BiocParallel::MulticoreParam(workers = 10))

temp<-metadata%>%select(cell_ID,fov_type)
avged<-split(temp,by = "fov_type")%>%
  sapply(function(x) rowMeans(normi[,x$cell_ID]))

pC<-avged%>%t()%>%scale()%>%t()%>%as.data.frame()%>%
    rownames_to_column("probe")%>%filter(probe%in%markers)%>%
    pivot_longer(-probe,names_to = "celltype",values_to = "z")%>%
    filter(celltype!="unknown")%>%
    group_by(celltype)%>%
    mutate(maxz=z/max(abs(z)),
           probe=factor(probe,levels=markers),
           celltype=factor(celltype,levels=names(cellcols)))%>%
    ungroup()%>%
    ggplot(aes(x=probe,y=celltype,fill=maxz))+geom_tile()+
    #scale_fill_gradient2(low = "blue",mid = "white",high = "red")+
    scale_fill_viridis_c(option = "D",direction = -1)+
    theme_minimal()+
    theme(axis.text.x = element_text(angle=45,vjust = 1,hjust = 1))

pC1<-temp%>%count(fov_type)%>%
  mutate(fov_type=factor(fov_type,levels = names(cellcols)))%>%
  ggplot(aes(x=n,y=fov_type))+geom_col()+theme_classic()+scale_x_log10()+        theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.spacing = unit(.01, "lines"),
        plot.margin=grid::unit(c(0,0,0,-5),"mm"),
        axis.text.x = element_text(angle=45,vjust = 1,hjust = 1))+
  labs(x="Cells",y="")
plot_grid(pC+theme(legend.position = "left"),pC1,
          align = 'h',axis = "tb",rel_widths = c(.6,.4))

```

#   Cell composition contrast
```{r}
library(limma)
library(speckle)

temp<-metadata%>%distinct(Class,Run_Tissue_name)
props<-getTransformedProps(metadata$fov_type,
                           metadata$Run_Tissue_name)
props<-props$TransformedProps[,temp$Run_Tissue_name]
designM<-model.matrix(~0+Class,data=temp)#groups around replicates=samplename
colnames(designM)<-str_replace(colnames(designM)," ","_")
fit<-lmFit(props,designM)
cont<-makeContrasts(ClassPeritoneal_endometriosis-ClassEutopic_endometrium,levels = designM)
fit<-contrasts.fit(fit,cont)
fit <- eBayes(fit, robust=TRUE)
res<-topTable(fit,number = nrow(props),coef = 1)
res<-res%>%rownames_to_column("cell_type")

## scRNA
library(Seurat)

sc<-readRDS("data/fixed_annotated_aux.seurat.rds")
sc<-subset(sc,cells = which(sc$Class%in%c("Eutopic endometrium","Peritoneal endometriosis")&
                            str_detect(sc$harmonized_major_plus_immune_cell_type,
                                 "Excl|Ery",negate=T)))
#fix labels
sc@meta.data<-sc@meta.data%>%
  mutate(celltype=case_when(str_detect(harmonized_major_cell_type,"Epi|Meso")~"Epithelial",
                            str_detect(harmonized_major_plus_immune_cell_type,"CD8|Gamma")~"CD8 T cells",
                            str_detect(harmonized_major_plus_immune_cell_type,"Th.+ T-cells")~"CD4 T cells",
                            str_detect(harmonized_major_cell_type,"EnS|Smooth|Mesen|Endot")~"Stroma",
                            TRUE~harmonized_major_cell_type))
sc@meta.data<-sc@meta.data%>%
  mutate(nosub=str_remove(harmonized_major_plus_immune_cell_type," \\(.+"))%>%
  mutate(nosub=str_replace(nosub,"Memmo","Memo")%>%str_remove("Early "))

aux_table<-sc@meta.data
temp<-aux_table%>%distinct(Class,SampleName)

props<-getTransformedProps(aux_table$celltype,
                           aux_table$SampleName)
props<-props$TransformedProps[,temp$SampleName]
designM<-model.matrix(~0+Class,data=temp)#groups around replicates=samplename
colnames(designM)<-str_replace(colnames(designM)," ","_")

fit<-lmFit(props,designM)
cont<-makeContrasts(ClassPeritoneal_endometriosis-ClassEutopic_endometrium,levels = designM)
fit<-contrasts.fit(fit,cont)
fit <- eBayes(fit, robust=TRUE)
resc<-topTable(fit,number = nrow(props),coef = 1)
resc<-resc%>%rownames_to_column("cell_type")

temp<-res%>%mutate(CosMx_LFC=logFC,
                   CosMx_P=adj.P.Val)%>%
  select(cell_type,CosMx_LFC,CosMx_P)
temp<-resc%>%mutate(RNA_LFC=logFC,RNA_P=adj.P.Val)%>%
  select(cell_type,RNA_LFC,RNA_P)%>%inner_join(temp)
temp$max_P<-apply(temp[,c(3,5)],1,max)
pE<-temp%>%ggplot(aes(x=RNA_LFC,y=CosMx_LFC,size=max_P))+
  geom_point()+geom_text_repel(aes(label=cell_type))+
  scale_size(range=c(6,3),breaks = c(0,.05,.5,1))+
  theme_linedraw()+geom_vline(xintercept = 0)+
  geom_hline(yintercept = 0)+
  labs(x="RNA LFC",y="protein LFC",size="Max. P.")+
  theme(legend.position = "bottom",panel.grid = element_blank())

temp<-metadata%>%count(Class,fov_type,name = "CosMx_n")%>%
  mutate(celltype=fov_type)%>%select(-fov_type)
temp<-aux%>%count(Class,celltype,name = "RNA_n")%>%inner_join(temp)%>%
  pivot_longer(RNA_n:CosMx_n,names_to = "platform",values_to = "n")
temp<-temp%>%group_by(Class,platform)%>%
  summarise(n=sum(n),celltype="Total")%>%bind_rows(temp)
pE1<-temp%>%mutate(celltype=factor(celltype,
  levels=c(names(cellcols)[c(1,2,4,3,8,9,6,5)],"Total")))%>%
  ggplot(aes(x=n,y=celltype,fill=Class))+
  geom_bar(stat="identity",position = "fill")+
  facet_grid(~platform)+scale_fill_manual(values=classcols)+
  theme_classic()+theme(strip.background = element_blank(),
                        legend.position = "bottom")+
  labs(x="Proportion")
plot_grid(pE,pE1,rel_widths = c(.4,.6))
```
# RNA difference
```{r}
alt_cellcols<-CellChat::scPalette(length(unique(sc$nosub)))
#alt_cellcols<-grDevices::colorRampPalette(alt_cellcols)(30)
names(alt_cellcols)<-sort(unique(aux_table$nosub))
alt_cellcols["B-cells"]<-cellcols["B cells"]
alt_cellcols["Plasmablast/Plasma cells"]<-cellcols["Plasma cells"]

#weird <- scales::trans_new("signed_sqrt",
#       transform=function(x) sign(x)*sqrt(abs(x)),
#       inverse=function(x) sign(x)*x**2)
temp<-aux_table%>%count(nosub,Class)
i<-temp%>%filter(Class=="Peritoneal endometriosis")%>%
  arrange(n)%>%distinct(nosub)%>%unlist()
pF<-temp%>%mutate(nosub=factor(nosub,levels=i),
                  Class=factor(Class,levels=rev(names(classcols))))%>%
  ggplot(aes(x=n,y=Class,fill=nosub))+geom_col()+
  facet_grid(rows=vars(nosub),switch="y")+
  scale_x_log10()+scale_fill_manual(values=alt_cellcols)+
  theme_classic()+theme(legend.position = "n",strip.background =element_blank(),
                        strip.text.y.left = element_text(angle=0,vjust = .5,hjust = 1),
                        panel.spacing = unit(.02, "lines"),axis.text.y = element_blank(),
                        axis.line.y = element_blank(),axis.ticks.y = element_blank())+
  labs(x="Cells",y="")
```
## Subsets 
```{r}
tcells<-subset(sc,cells=which(str_detect(sc$celltype,"T cells")))
tcells<-RunUMAP(tcells,repulsion.strength = 5,dims = 1:15)
tcoords<-tcells@reductions$umap@cell.embeddings%>%as.data.frame()%>%
  rownames_to_column("barcode")
tcoords%>%write_csv("RNA_Tcells_umap.csv")
nont<-subset(sc,cells = which(str_detect(sc$celltype,"B|Mast|NK|Plas|My")))
nont<-RunUMAP(nont,dims = 1:15)
nontcoords<-nont@reductions$umap@cell.embeddings%>%
  as.data.frame()%>%rownames_to_column("barcode")
nontcoors%>%write_csv("RNA_nonTcells_umap.csv")
sc<-subset(sc,subset=celltype%in%c("Stroma","Epithelial"))
sc<-RunUMAP(sc,dims = 1:15)
nonimmcoords<-sc@reductions$umap@cell.embeddings%>%
  as.data.frame()%>%rownames_to_column("barcode")
nonimmcoords%>%data.table::fwrite("RNA_nonImmune_umap.csv")
coords<-tcoords%>%mutate(ct="T-cells")%>%
  bind_rows(nonimmcoords%>%mutate(ct="non Immune"))%>%
  bind_rows(nontcoords%>%mutate(ct="non T-cells"))
coords<-aux_table%>%select(barcode,nosub,Class)%>%
  right_join(coords)
coords<-coords%>%group_by(nosub)%>%mutate(center1=median(umap_1),
                                          center2=median(umap_2))

i<-temp%>%distinct(ct,nosub)%>%
  mutate(ct=factor(ct,levels=c("non Immune","non T-cells","T-cells")))%>%
  arrange(ct,nosub)
alt_cellcols<-alt_cellcols[i$nosub]
pF1<-coords%>%
    mutate(nosub=factor(nosub,levels=names(alt_cellcols)))%>%
    ggplot(aes(x=umap_1,y=umap_2,col=nosub))+
    geom_point(size=.1)+geom_text_repel(data=coords%>%
                                          distinct(nosub,center1,center2),
                                        aes(x=center1,y=center2,label=nosub),
                                        col="black")+
    scale_color_manual(values=alt_cellcols)+
    guides(col=guide_legend(override.aes = list(size=5),ncol=1))+
    facet_grid(~ct)+theme_light()+
    theme(strip.background = element_blank(),axis.text.x = element_blank(),
          legend.position = "n",plot.margin=grid::unit(c(0,1,-5,0),"mm"))+
  labs(col="cell type",x="")
pF2<-coords%>%ggplot(aes(x=umap_1,y=umap_2))+
    stat_density_2d(aes(fill = ..level..), geom = "polygon")+
    facet_grid(rows=vars(Class),cols=vars(ct))+
    scale_fill_viridis_c(option = "B",direction = -1,labels=c("Min","Max"),breaks=c(0,.01))+
    theme_light()+theme(strip.background = element_blank(),
                        strip.text.y.right = element_text(colour ="black"))+
    labs(x="UMAP 1",y="UMAP 2")
```
# CellChat
```{r}
library(CellChat)
library(ggrepel)

#names(alt_cellcols)<-sort(unique(aux_table$harmonized_major_plus_immune_cell_type))

cc_res<-readRDS("docs/CellChat_nosubPE_Eu.rds")
compareInteractions(cc_res[[1]], show.legend = F, group = c(1,2))

cc_res$`Eutopic endometrium`<-netAnalysis_computeCentrality(cc_res$`Eutopic endometrium`)
cc_res$`Peritoneal endometriosis`<-netAnalysis_computeCentrality(cc_res$`Peritoneal endometriosis`)

temp<-lapply(names(cc_res)[-1],function(x)
  rowSums(cc_res[[x]]@net$count)+colSums(cc_res[[x]]@net$count))
names(temp)<-names(cc_res)[-1]
temp<-lapply(names(temp), function(x) 
  temp[[x]]%>%as.data.frame()%>%mutate(Class=x)%>%
    rownames_to_column("ct"))%>%bind_rows()
colnames(temp)[2]<-"total_interactions"

pG<-temp%>%filter(ct!="Erythrocytes")%>%
    mutate(ct=factor(ct,levels=rev(levels(pF$data$nosub))))%>%
    ggplot(aes(y=ct,x=total_interactions,fill=Class))+
    geom_bar(stat="identity",position="fill")+
    geom_vline(xintercept = .5)+
    theme_classic()+theme(axis.line.y = element_blank(),
                          legend.position = "none")+
    scale_fill_manual(values=classcols)+
    labs(y="",x="Total interaction")

cts<-unique(temp$ct)
#netVisual_diffInteraction() not working
pH<-netVisual_diffInteraction(cc_res[[1]],
                              comparison = c(1,2),
                              targets.use = "B-cells",
                              weight.scale = T,
                              remove.isolate = T,
                              sources.use=cts[cts!="Erythrocytes"])

pH1<-netVisual_diffInteraction(cc_res[[1]],
                              comparison = c(1,2),
                              sources.use = "B-cells",
                              weight.scale = T,
                              remove.isolate = T,
                              targets.use=cts[cts!="Erythrocytes"])

```
# CosMx neighbors
```{r}
temp<-edges%>%count(Class,type_to,type_from)
p0<-temp%>%filter(type_to==type_from&type_to!="")%>%
  ggplot(aes(x=Class,y=n,col=type_from,group=type_from))+
  geom_point()+geom_line()+
  scale_color_manual(values = cellcols)+scale_y_log10()+
  theme_light()+ggtitle("Homotypic interactions change")+
  theme(axis.text.x = element_text(angle=45,vjust = 1,hjust = 1))
  labs(y="Interactions",col="",x="")

temp<-temp%>%filter(type_to!=""&type_from!="")%>%
  pivot_wider(names_from = type_to,values_from = n)%>%
  mutate(across(3:11,function(x) replace_na(x,0)))
temp1<-split(temp,f = temp$Class)%>%lapply(function(x) 
  x%>%mutate(across(3:11,function(y) 
    100*y/sum(y))))
cl<-temp1$`Eutopic endometrium`%>%select(-c(1,2))%>%t()%>%dist()%>%hclust(method = "complete")
temp1<-temp1%>%bind_rows()
temp1<-temp1[c(cl$order,9+cl$order),]
hm<-temp1%>%select(-c(1,2))%>%t()%>%
  Heatmap(name="%",column_split=temp1$Class,column_labels = temp1$type_from,
          col=colorRamp2(c(60,40,15,5,1,0),
                         c("#14081D","#651B79","#de7065ff","#FFC28B","#FCFDBF","white")),
          cluster_columns = cl,show_row_dend = F,row_names_side = "left",
          show_heatmap_legend = T,column_names_rot = 45)
hm<-grid.grabExpr(draw(hm))

plot_grid(hm,p0,align = "h",axis = "tb")


#partA<-plot_grid(pC+theme(axis.text = element_text(size=10),legend.position = "left",plot.margin=grid::unit(c(1,10,0,-3), "mm")),pC1+theme(axis.text = element_text(size=10),plot.margin=grid::unit(c(0,0,0,-15), "mm")),align="h",axis="tb",rel_widths = c(.8,.2),labels=c('B',''))
#partB<-plot_grid(pE+theme(axis.text = element_text(size=10)),pE1+theme(axis.text = element_text(size=10)),align="h",axis="tb",labels=c('C','D'))
#partA<-plot_grid(partA,partB,nrow = 2,rel_heights = c(.4,.6))
#partB<-plot_grid(pF1,pF2,nrow=2,align='v',axis="lr",rel_heights = c(.35,.65))
#partC<-plot_grid(pF,pG+theme(axis.text.y = element_blank(),axis.ticks.y =element_blank(),plot.margin=grid::unit(c(0,0,0,-10), "mm")),rel_widths = c(.7,.3),align = "h",axis = "tb")
#partC<-plot_grid(partC,NULL,nrow=2,rel_heights = c(.6,.4))
pdf("temp.pdf",width = 13,height = 11)
plot_grid(plot_grid(NULL,partA),plot_grid(partB,partC,rel_widths = c(.6,.4)),nrow=2)
dev.off()

save(hm0,p1,lgd0,pB,pB1,pC,pC1,pE,pH1,pI,
     p0,file="docs/Fig1.RData")

```

<<<<<<< HEAD











#Stattistics for LA High vs Low
```{r}
##stage comparision in Low vs High (not including patient 1 and 2 that have only endometrioma)
Metadata.Stats = read_xlsx("data/LA_Supp_organization.xlsx", sheet = 1)
counts = table(Metadata.Stats$Group[-c(1:2)], Metadata.Stats$Stage.Group[-c(1:2)])
Chisq <- chisq.test(counts)
Chisq

##Low vs High mestral cycle comparision pch = circles, 21= is the type symbol (21 =circle)
Metadata.Stats = read_xlsx("data/LA_Supp_organization.xlsx", sheet = 1)
counts = table(Metadata.Stats$Group[-c(1:2)], Metadata.Stats$Menstral.Cycle.Phase[-c(1:2)])
Chisq <- chisq.test(counts)
Chisq


##Low vs High Age comparision 
Metadata.Stats = read_xlsx("data/LA_Supp_organization.xlsx", sheet = 1)
Metadata.Stats = Metadata.Stats[-c(1:2), c("Group", "Age")]

Low <- c(32, 41, 31)
High <- c(32, 24, 40, 42, 28, 21, 37, 46, 36)

wilcox.test(Low, High)

p1 <- ggplot(Metadata.Stats, aes(x=as.character(Group), y=Age)) +
  geom_boxplot(width = 0.5) + 
  geom_point(pch = 21) +
  theme_classic(base_size = 20) +
  theme(axis.text.x=element_text(angle = -45, vjust = 0.5, colour = "black"), axis.text.y = element_text(colour = "black")) +
  geom_text(x=0.75, y=45, label= "p-value = 1", size = 10)
  
p1


##Low vs High Deep vs superficial comparision by tissue
Metadata.Stats = read_xlsx("data/LA_Supp_organization.xlsx", sheet = 2)
counts = table(Metadata.Stats$Group, Metadata.Stats$Deep_superficial)
Chisq <- chisq.test(counts)
Chisq



##Low vs High mutations status comparision by tissue
Metadata.Stats = read_xlsx("data/LA_Supp_organization.xlsx", sheet = 2)
counts = table(Metadata.Stats$Group[-c(3, 12, 16, 18, 20:22, 28)], Metadata.Stats$ARID1A_Mut[-c(3, 12, 16, 18, 20:22, 28)])
Chisq <- chisq.test(counts)
Chisq


counts = table(Metadata.Stats$Group[-c(3:4, 6:8, 11:14, 16, 18:23, 28)], Metadata.Stats$KRAS_Mut[-c(3:4, 6:8, 11:14, 16, 18:23, 28)])
Chisq <- chisq.test(counts)
Chisq

```













#looking at specific receptors and ligans in EnEpi and Central Memory in endometriosis
```{r}
sel <- subset(d10x.seurat, subset = Master_Assigned_Cell_Type %in% "EnS cells")

sel <- subset(sel, subset = Class %in% c("Peritoneal endometriosis", "Endometrioma"))

Immune.markers = read.delim("data/cell_chat_receptor_ligand_filtered.csv", sep = ",")
head(Immune.markers)

selected.features = unique(as.character(unique(Immune.markers$M2Macs)))


DotPlot(sel, assay = "RNA", features = c("IGF1", "THY1", "ANXA1", "C3", "IGF"), group.by = "Class") + labs(title = "Ligand Receptor in CLASS EnS cells") +
 theme_classic(base_size = 25) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.8, colour="black"), axis.text.y = element_text(color = "black")) +
  coord_flip() +
  scale_colour_gradient2(low = "#2166ac", mid = "#f7f7f7", high = "#b2182b")


sel <- subset(Endometriosis, subset = Immune_Cluster_ID %in% "Plasma cells")

sel <- subset(Immune.Cells.TrueV2, subset = Class %in% c("Peritoneal endometriosis", "Endometrioma"))

sel <- subset(sel, subset = Immune_Cluster_ID %in% "Plasma cells")

DotPlot(sel, assay = "RNA", features = c("IL1B", "NAMPT", "CD69", "GZMA", "HLA-B", "APP", "CLEC2B", "INSR"), group.by = "Class") + labs(title = "Ligand Recipient signals from Activated Plasma cell") +
 theme_classic(base_size = 25) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.8, colour="black"), axis.text.y = element_text(colour = "black")) +
  coord_flip() +
  scale_colour_gradient2(low = "#2166ac", mid = "#f7f7f7", high = "#b2182b")



Endometriosis <- subset(Immune.Cells.TrueV2, subset = Class %in% "Peritoneal endometriosis")

Idents(sel) <- "Immune_Cluster_ID"

DotPlot(sel, assay = "RNA", features = 
          c("NAMPT", "CD69", "GZMA", "HLA-B", "APP", "CLEC2B"), split.by = "Class", cols= "BrBG") + labs(title = "COX expression in Immune cells in endometriosis") +
 theme_classic(base_size = 20) #+
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.8)) +
  #coord_flip() +
  #scale_colour_gradient2(low = "#2166ac", mid = "#f7f7f7", high = "#b2182b")

```






















##Figure1A will be an HE of LA high with big aggregate and a small aggregate





## Fig 1B Beeswarm plot of the #LA in endometrioma and endometriosis.

```{r}
#endometrioma vs Endometriosis LA by block
LA.extended.data = read_xlsx("data/2023-01-25 LA Summary MH v2.xlsx", sheet = 4)

#endometrioma vs Endometriosis LA by blocks stats
Endometrioma <- c(0,6,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,5,1,0,0,0,0,0,0,0)
Endometriossis<- c(0,0,0,0,2,0,0,2,3,0,4,8,10,7,7,9,0,0,1,1,11,0,5,0,0,0,0,0,0,0,0,2,2,3,4,0,0,1,5,0,13,1,3,4,0,2,0,1,5,2,1,1,0,1,1,1,1,1,0,1,1,0,0,1,3,0,0,0,0,0,0,0,0) 

t_test <- t.test(Endometrioma, Endometriossis)
t_test

p1 <- ggplot(LA.extended.data, aes(x=Category, y=Number_of_lymphoid_aggregates, color= Category)) + 
  scale_colour_manual(name="Category", values = c("Endometrioma" = "#7b3294", 
                                                    "Endometriosis" = "#d9f0d3")) + 
  geom_beeswarm(size=3, shape =21, fill= "black", stroke =1.75) + 
  geom_text(x=1, y=12, label=paste0("p-value = ", t_test$p.value), colour="black", family = "san", fontface = "plain", size =5) +
  ggtitle("Number of LA in each Endometrioma and Endometriosis Block") + 
  theme_classic(base_size = 20)

p1


#endometrioma vs Endometriosis LA by part
LA.simple.data = read_xlsx("data/2024-11-06_Combine_scGroup_withPriority2_analyze_only_endometriosis_positive_parts_deidentified_Rfile.xlsx", sheet = 5)
LA.simple.data <- LA.simple.data %>% filter(!grepl("No ", Category))
#endometrioma vs Endometriosis LA by part stats
#Endometrioma <- c(6,0,0,0,0,0,0,0,6,0,0,8,3,0,0,0,0,0,0,0,0,0,0,0,0,0,5)
#Endometriossis<- c(8,0,0,0,0,0,4,0,3,0,12,17,16,0,0,16,2,0,0,0,0,0,0,1,0,0,0,0,4,7,0,0,1,5,0,14,3,4,10,3,4,2,4,0,0,0,0,0,0,3,0,1,17,0,0,0,10,0,1,0,0,0,0,0,0,0,0,0,0,9,0,0,0,29,13,4,1,6,5,7,4,0,0,2,1,0,0,15,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,8,6,8,0,0,0,5,0,1,0,3,0,7,1,0,35,1,12,26,20,7,0,3,0,4,0,0,0,13,2,0,0,0,0,7,0,1,6,0,0,0,0,0,0,0,0,28,2,8,3,0,0,0,0) 
#wil_test <- wilcox.test(Number_of_lymphoid_aggregates ~ Category, data = LA.simple.data)
#wil_test <- wilcox.test(Endometrioma, Endometriossis)
#wil_test

wilcox.test(Number_of_lymphoid_aggregates ~ Category, data = LA.simple.data)

#p1 <- ggplot(LA.simple.data, aes(x=Category, y=Number_of_lymphoid_aggregates, color=Category)) + 
  #scale_colour_manual(name="Category", values = c("Endometrioma" = "#7b3294", 
   #                                                 "Endometriosis" = "#d9f0d3")) + 
  #geom_quasirandom(size=2.5) + 
  #geom_text(x=0.75, y=30, label="p-value = 0.2499", colour="black", family = "san", fontface = "plain", size =5) +
  #ggtitle("Number of LA in each Endometrioma and Endometriosis part") + 
  #theme_classic(base_size = 20)

p1 <- ggplot(LA.simple.data, aes(x=Category, y=Number_of_lymphoid_aggregates)) + 
  geom_quasirandom(size=2.5) + 
  geom_text(x=0.75, y=30, label="p-value = 0.2499", colour="black", family = "san", fontface = "plain", size =5) +
  ggtitle("Number of LA in each Endometrioma and Endometriosis part") + 
  theme_classic(base_size = 20)


p1


#Number of LA in Endometrioma and Endometriosis by part with patients labeled
LA.summarised <- LA.simple.data %>%
  select(`PAPER ID`, Number_of_lymphoid_aggregates, Category) %>% 
  group_by(`PAPER ID`, Category) %>% 
  summarise(total_lymph_agg = sum(Number_of_lymphoid_aggregates)) %>% 
  mutate(plotting_id = paste0(`PAPER ID`, Category))

LA.summarised$plotting_id <- factor(LA.summarised$plotting_id, levels = LA.summarised$plotting_id[order(LA.summarised$total_lymph_agg)])

wilcox.test(total_lymph_agg ~ Category, data = LA.summarised)

#ggplot(LA.summarised, aes(x=plotting_id, y=total_lymph_agg, color=Category)) + 
  #scale_colour_manual(name="Category", values = c("Endometrioma" = "#7b3294", 
                                                   "Endometriosis" = "#d9f0d3")) + 
  #geom_quasirandom(size=2.5) + 
  #scales= "free_x" means that it can ignore any point that is not in that category
  #facet_wrap(~Category, scales = "free_x") + 
   #geom_text(x=1, y=15, label="p-value = 0.0422", colour="black", family = "san", fontface = "plain", size =5) +
  #ggtitle("Number of LA in each Endometrioma and Endometriosis part") + 
  #theme_classic(base_size = 20) +
  #theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5)) +
  #geom_hline(yintercept = 0, linetype = 2) #+
  #scale_x_discrete(labels = setNames(LA.summarised$`PAPER ID`, LA.summarised$plotting_id))
  
  

ggplot(LA.summarised, aes(x=plotting_id, y=total_lymph_agg)) + 
  geom_quasirandom(size=2.5) + 
  #scales= "free_x" means that it can ignore any point that is not in that category
  facet_wrap(~Category, scales = "free_x") + 
  #geom_text(x=1, y=15, label="p-value = 0.001217", colour="black", family = "san", fontface = "plain", size =5) +
  ggtitle("Number of LA in each Endometrioma and Endometriosis in each patient") + 
  theme_classic(base_size = 20) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5)) +
  geom_hline(yintercept = 0, linetype = 2) +
  scale_x_discrete(labels = setNames(LA.summarised$`PAPER ID`, LA.summarised$plotting_id))
  


#Beeswarm plot of total La by patient

LA.summarised <- LA.simple.data %>%
  select(`PAPER ID`, Number_of_lymphoid_aggregates, Category) %>% 
  group_by(`PAPER ID`, Category) %>% 
  summarise(total_lymph_agg = sum(Number_of_lymphoid_aggregates)) %>% 
  mutate(plotting_id = paste0(`PAPER ID`, Category))

LA.summarised$plotting_id <- factor(LA.summarised$plotting_id, levels = LA.summarised$plotting_id[order(LA.summarised$total_lymph_agg)])

wilcox.test(total_lymph_agg ~ Category, data = LA.summarised)

ggplot(LA.summarised, aes(x=Category, y=total_lymph_agg)) + 
     geom_quasirandom(size=2.5) + 
     geom_text(x=1, y=60, label="p-value = 0.001217", colour="black", family = "san", fontface = "plain", size =5) +
     ggtitle("Number of LA in each Endometrioma and Endometriosis in patient") + 
     theme_classic(base_size = 20) +
     geom_hline(yintercept = 0, linetype = 2) 

```

## Fig 1C
```{r}
#Heatmap of peritoneal endometriosis lesions in parts
library(pheatmap)

la_summary = read_xlsx("data/2023-01-25 LA Summary MH v2.xlsx", sheet = 2)
la_summary[is.na(la_summary)] = -10

my_cluster_col = as.data.frame(la_summary[3])
my_cluster_col$Patient = as.character(my_cluster_col$Patient)
rownames(my_cluster_col) <- paste0("row_", seq(nrow(my_cluster_col)))
la_summary = la_summary[c(-8, -14:-41),c(-1:-3, -11:-31)]  #need to adjust rows depending on the samples being included.

test_labels <- as.matrix(la_summary)
rownames(la_summary) <- paste0("row_", seq(nrow(la_summary)))
test_labels[test_labels < 0] <- "\u2217\u2217"
test_labels[test_labels < 0] <- "\uFF0A"

p1 <- pheatmap(la_summary,
         cluster_cols = F,
         display_numbers = test_labels,
         fontsize_number=12,
         cellheight=20,
         annotation_row = my_cluster_col,
         clustering_distance_row = "euclidean",
         main = "12 samples euclidean")

p1


#or plot based on Beeswarm plot
LA.simple.data = read_xlsx("data/2023-01-25 LA Summary MH v2.xlsx", sheet = 5)
LA.simple.data.sel = LA.simple.data[c(-1,-2,-3, -5, -10, -23, -33, -36, -47), c("LA_Activity", "Category", "Number_of_lymphoid_aggregates")]

p1 <- ggplot(LA.simple.data.sel, aes(x=LA_Activity, y=Number_of_lymphoid_aggregates, color=LA_Activity)) + 
  #scale_colour_manual(name="Category", values = c("Endometrioma" = "#7b3294", 
   #                                                 "Endometriosis" = "#d9f0d3")) + 
  geom_beeswarm(size=2.5) + 
  #geom_text(x=1, y=15, label="p-value = 0.4654", colour="black", family = "san", fontface = "plain", size =5) +
  ggtitle("Number of LA in patients defined as High and low by part") + 
  theme_classic(base_size = 30)
  

p1

```






## Fig 1F
PCA based on the following cell proportion for each samples:

Follow the below steps for PCA analysis 

1. import Annotation tracker to add the LA_Status Epi_data tab - we need that col to label ggplot ( Master_Assigned_Cell_Type is already included in mate.data)
2. import `broom`package to parse PCA output from `prcomp`
3. Only meta.data from Seurat obj is needed for the PCA analysis
4. If the label on plot is too crowded. consider using `Patient.No.`

```{r}
d10x.seurat <- readRDS("data/fixed_annotated_aux.seurat.rds")@meta.data
excel_path <- "data/Annotation tracker.xlsx"
sheets <- readxl::excel_sheets(excel_path)
Annotation_tracker <- sheets %>% 
                      lapply(. , function(s){ read_excel(excel_path, sheet = s) %>% 
                                              mutate(across(where(is.character),str_trim))
                                              }) %>%
                      setNames(sheets)

Annotation_tracker$Epi_Data 
```


```{r}
immue_endo_t<- d10x.seurat@meta.data%>%
                  #filter(Tissue_LA_Activity == "Yes") %>%
                  #filter(Master_Assigned_Cell_Type) %>%
                  #filter(! is.na(Number_of_lymphoid_aggregates)) %>%
                  group_by(SampleName,Master_Assigned_Cell_Type) %>%
                  summarise(count=n()) %>%
                  mutate(prop= count/sum(count)*100)

anno_df <- d10x.seurat@meta.data  %>%
           left_join(Annotation_tracker$Epi_Data, by=c("Patient.No."="Patient_ID")) %>%
           dplyr::filter(LA_Status != "Exclude")  %>% 
           distinct(Patient.No.,LA_Status,SampleName) %>%
           column_to_rownames("SampleName")

heatmap_df <- immue_endo_t  %>% 
              dplyr::select(Master_Assigned_Cell_Type,prop,SampleName) %>%
              spread(Master_Assigned_Cell_Type, prop) %>%
              column_to_rownames("SampleName") %>%
              as.matrix() %>%
              replace(.,is.na(.),0) %>%
              t()

heatmap_df %>% 
t %>%
prcomp(scale = TRUE) %>%
augment(.) %>%
inner_join(anno_df %>% rownames_to_column(".rownames"), by=".rownames") %>%
  ggplot(aes(x=.fittedPC1, y=.fittedPC2,color=LA_Status,label=.rownames))+
  geom_point(size=5) +
  geom_text(nudge_x = 0.2, nudge_y = 0.2,angle=20,size=3) +
  labs(x="PC1",y="PC2") + 
  theme_half_open(12) 


```

# Final Fig pannel combine using cowplot

```{r}
fig1top <- plot_grid(fig1a, fig1b, labels = c("A", "B"), ncol = 2, label_size = 10, rel_widths = c(1, 1.15) )
fig1middle <- plot_grid(fig1c, fig1d, NULL, labels = c("C", "D"), ncol = 3, label_size = 10)
fig1bottom <- plot_grid(fig1e, fig1f, NULL, labels = c("E", "F"), ncol = 3, label_size = 10)
fig1 <- plot_grid(fig1top, fig1middle,fig1bottom ,ncol = 1, label_size = 10, align = 'v', hjust=-6, rel_heights = c(0.5, 0.25))
```

#save figures

```{r}
ggsave("data/fig1.pdf",fig1, width = 210, 
       height = 148.5, 
       units = "mm")
```


#DEG of B-cell in LA true vs false and volcano plot
```{r}
#DEG

Bcells<- subset(Final.Immune.Cells.PE, subset = Final_Immune_Cluster_ID %in% "B-cells")

Idents(Bcells) <- Bcells$has_LA #this is to say which columns to use

DEG <- FindMarkers(Bcells, ident.1 = "TRUE", ident.2 = "FALSE")

DEG <- DEG %>% rownames_to_column("genes")

#volcano plot
fc.p = 0.5
pv.p = 0.05
DEG$STATUS = "NOT.SIG"
DEG[DEG$avg_log2FC < -fc.p & DEG$p_val_adj < pv.p, ]$STATUS = "Down"
DEG[DEG$avg_log2FC > fc.p & DEG$p_val_adj < pv.p, ]$STATUS = "Up"

p1 <- ggplot(DEG, aes(x = avg_log2FC, y = -log10(p_val_adj))) +
  geom_point(aes(color = STATUS), cex = 1.45) +
  scale_color_manual(values = c("darkgreen", "grey", "red")) +
  ylim(0, 80) +
  theme_bw(base_size = 15) + 
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5)) +
  geom_text_repel(
    data          = subset(DEG, avg_log2FC > 0.5),
    aes(label = genes),
    segment.size  = 0.4,
    direction     = "y",
    segment.color = "grey50",
    hjust         = 0
  ) +
  geom_text_repel(
    data          = subset(DEG, avg_log2FC < -0.5),
    aes(label = genes),
    segment.size  = 0.4,
    direction     = "y",
    segment.color = "grey50",
    hjust         = 1
  ) +
  geom_hline(yintercept = -log10(pv.p), linetype="dotted", color = "black", size= 0.5) +
  geom_vline(xintercept = -fc.p, linetype="dotted", color = "black", size= 0.5) +
  geom_vline(xintercept = fc.p, linetype="dotted", color = "black", size= 0.5) #+
  #labs(title = paste0(feature, " : ", c, " (n=", nDEG, ") FC: ", fc.p, " pval = ", pv.p))

p1



#doing DEG and volcano plots based pseudobulk
library("SingleCellExperiment")

#first convert objects to singleCellExperiiment object
Bcells2 <- as.SingleCellExperiment(Final.Immune.Cells.PE)

#need library glmGampoi
BiocManager::install("glmGamPoi")
library("glmGamPoi")

#This is the grouping of cells
reduced_sce <- pseudobulk(Bcells2, group_by = vars(Patient.No., has_LA, SampleName, harmonized_major_plus_immune_cell_type), 
                          n_cells = n())

save("reduced_sce", file = "reduced_sce.rda")

# Remove NA's
reduced_sce <- reduced_sce[,!is.na(reduced_sce$SampleName)]

colData(reduced_sce)$Patient.No. <- factor(colData(reduced_sce)$Patient.No.)
colData(reduced_sce)$has_LA <- factor(colData(reduced_sce)$has_LA)

# Use DESeq2's size factor calculation procedure
fit <- glm_gp(reduced_sce, design = ~ has_LA + harmonized_major_plus_immune_cell_type + Patient.No., size_factor = "ratio", verbose = TRUE)

#to test the differential expression
es <- test_de(fit, contrast = cond(has_LA = "TRUE") - cond(has_LA = "FALSE"))
```


#pathway analysis
```{r}
BiocManager::install("clusterProfiler")

BiocManager::install("msigdbr")

library(clusterProfiler)
library(msigdbr)

m_df <- msigdbr(species = "Homo sapiens")%>%filter(str_detect(gs_name,"REACTOME"))%>%
    distinct(gs_name,gene_symbol)

temp <- DEG %>% mutate(keep = case_when(avg_log2FC > 0 & pct.1 > 0.8 ~ TRUE,
                                avg_log2FC < 0 & pct.2 > 0.8 ~ TRUE,
                                TRUE ~ FALSE)) %>%
  filter(keep, p_val_adj < 0.05) %>%
  mutate(STATUS = ifelse(avg_log2FC > 0, "Up", "Down"))

#temp<-DEG%>%filter(p_val_adj<0.05 & abs(avg_log2FC)>.5)
oraRes<-temp%>%split(temp$STATUS)%>%lapply(function(x) enricher(x$genes,TERM2GENE = m_df))
temp<-lapply(names(oraRes),function(x) 
  oraRes[[x]]%>%as.data.frame()%>%mutate(cluster=x))%>%bind_rows()


barplot(oraRes$Up)

dotplot(oraRes$Up)

dotplot(oraRes$Down)
```
=======
>>>>>>> 6c51002ddffc1b651b8965feed359556bbf00675
