---
title: "myeloid_story_LA_paper"
author: "Raghad Said"
date: "2026-06-25"
---

```{r}
library(tidyverse)
library(Seurat)
library(dplyr)
library(ggplot2)

#load dataset
endo<- readRDS("/media/cedars/caipirinha/speed/share/fixed_annotated_aux.seurat.rds")
#eutopic and pe only
eut_ect<-subset(endo, subset = Class %in% c("Eutopic endometrium", "Peritoneal endometriosis"))

```

```{r}
sub_myeloid<-subset(eut_ect, subset=harmonized_major_cell_type=="Myeloid cells")
sub_myeloid<- FindClusters(sub_myeloid, resolution = 0.2)

markers <- list(
#  "Myeloid" = c("CD68", "ITGAM", "CD14", "CD33", "CSF1R", "AIF1"),
  "Myeloid" = c("CD68", "CD33"),
  "M1" = c("IL1B","TNF","CXCL8","CXCL10"),
  "M2" = c("MRC1","CD163","FOLR2","CCL18"),
  "MDSC" = c("S100A8","S100A9","OLR1","FCN1","PTGS2"),
  "Resident" = c("C1QA","C1QB","C1QC","SELENOP"),
  "APC" = c("HLA-DRA","CD74","FCER1A","CLEC10A"))
DotPlot(
    sub_myeloid,
    features = markers,
    group.by = "SCT_snn_res.0.2") +
    RotatedAxis()+
    theme(axis.text.x = element_text(angle=45,vjust = 1,hjust = 1))+
    scale_colour_gradient2(low = "#2166ac", mid = "#f7f7f7", high = "#b2182b")

#updating cluster names
sub_myeloid@meta.data<-sub_myeloid@meta.data%>%
  mutate(myeloid_cl=case_when(SCT_snn_res.0.2==0~"M1 macrophages",
        SCT_snn_res.0.2==1~"M2 macrophages",
        SCT_snn_res.0.2==2~"DCs",
        SCT_snn_res.0.2==3~"MDSCs",
        TRUE~"undefined myeloid"))

sub_myeloid@meta.data%>%
  select(SCT_snn_res.0.2,myeloid_cl)%>%
  rownames_to_column("barcode")%>%
  write_csv("temp_myeloid_new_labels.csv")

```

```{r}
#inflammatory score
# Inflammatory gene module
inflam_genes <- list(c(
  "IL1B", "CXCL8", "PTGS2", "NLRP3", "CCL20",
  "PTX3", "TNFAIP3", "NFKBIA", "CEBPB"
))

# Calculate module score
m1 <- AddModuleScore(
  m1,
  features = inflam_genes,
  name = "Inflammatory",
  nbin = 10
)

# Average score per patient
score_df <- m1@meta.data %>%
  group_by(`Patient.No.`, has_LA) %>%
  summarise(
    score = mean(Inflammatory1, na.rm = TRUE),
    .groups = "drop"
  )

# Plot
ggplot(score_df, aes(has_LA, score, color = has_LA)) +
  geom_boxplot(fill = NA, outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 2, alpha = 0.8) +
  scale_color_manual(values = c("LA-" = "#7B6FC2",
                                "LA+" = "#9ACD32")) +
  theme_classic() +
  labs(x = NULL, y = "Inflammatory activation score") +
  NoLegend()

# Statistics
wilcox.test(score ~ has_LA, data = score_df)