---
title: "scDrugPrio"
---


```{r}
library(scDrugPrio)# annoying to install
#library(STRINGdb)
library(tidyverse)
#library(biomaRt)
#library(Seurat)
library(igraph)

load_sample_data()#to get drugbank data
```

# get public data
```{r}
drug_bank_example_data<-as.data.frame(drug_bank_example_data)
drug_bank_example_data<-drug_bank_example_data%>%
  filter(drug_action=="Humans"&str_detect(status,"approved"))
drug_bank_example_data<-drug_bank_example_data%>%
  filter(str_detect(status,"withdrawn",negate = T))

#string_db <- STRINGdb$new(
#  version          = "12.0",
#  species          = 9606,
#  score_threshold  = 900,#max confidence
#  input_directory  = "",
#  protocol         = "https")
#string_db
#edges<-get.edgelist(string_db$graph)
#edges<-as.data.frame(edges)
#only 1 result when using this network: insulin-HLADQA2(MDSCs)

ensembl <- useEnsembl(biomart = "genes", dataset = "hsapiens_gene_ensembl")
bm <- getBM(attributes = c('entrezgene_id', 'external_gene_name'), mart = ensembl)
temp<-lit_ppi%>%as.data.frame()%>%
  mutate(Protein_A=as.numeric(Protein_A))%>%
  left_join(bm,by=c("Protein_A"="entrezgene_id"))%>%
    mutate(nAname=external_gene_name)%>%
  dplyr::select(-external_gene_name)
temp<-temp%>%
  mutate(Protein_B=as.numeric(Protein_B))%>%
  left_join(bm,by=c("Protein_B"="entrezgene_id"))%>%
    mutate(nBname=external_gene_name)%>%
  dplyr::select(-external_gene_name)%>%
  drop_na()%>%dplyr::select(nAname:nBname)%>%
  distinct()%>%as.matrix()
ppin<-temp%>%ppin_formatting()
#write_csv(as.data.frame(ppin),file = "docs/edges.temp")


#Select unique drug target combinations found in the PPIN's LCC of
g<-ppin%>%graph_from_data_frame(directed = F)
nodes<-V(g)$name
#edges%>%as.data.frame()%>%write_csv("docs/PPIN.temp")
drug_bank_example_data<-drug_bank_example_data%>%filter(gene_symbol%in%nodes)
drug_target_matrix <- create_drug_target_matrix(drugID = drug_bank_example_data[,1], 
                                                target = drug_bank_example_data[,6])
#sum(colSums(is.na(drug_target_matrix))<nrow(drug_target_matrix))==ncol(drug_target_matrix)
drug_target_matrix <- prepare_drug_target_matrix_for_network_distance_calculation(ppin,
                                                                                  drug_target_matrix, 
                                                                                  file_name = "scdrug_in_ppin", 
                                                                                  out_dir = "docs")
#[1] "n of unique proteins/genes in PPIN: n =  17345"
#[1] "n unique drugs included in analysis: n = 94"
#[1] "n drugs with unique drug target combinations: n = 74"
#[1] "n unique drugs that can be extracted from UNIQUE_DRUG_TARGET_COMBINATION_scdrug_in_ppin.txt using SAME_DRUG_TARGETS_scdrug_in_ppin.txt: n = 94"
#[1] "Drug matrix = FORMATTED"
same_drugs <- read.table(file = "docs/SAME_DRUG_TARGETS_scdrug_in_ppin.txt", sep="\t", header = T, stringsAsFactors = F)

```

# focus on your data
```{r}
sc<-readRDS("data/fixed_annotated_aux.seurat.rds")
sc<-subset(sc,cells = which(sc$Class=="Peritoneal endometriosis"&
              str_detect(sc$harmonized_major_plus_immune_cell_type,"^Ex|Ery",negate = T)))
gc()
sc$has_LA<-ifelse(sc$LA_per_Sample>0,"LA+","LA-")
sc@meta.data<-sc@meta.data%>%
  mutate(nosub=str_remove(harmonized_major_plus_immune_cell_type," \\(.+"))%>%
  mutate(nosub=str_replace(nosub,"Memmo","Memo")%>%str_remove("Early "))

temp<-read_csv("data/myeloid_subgroups.csv")
aux_table<-sc@meta.data%>%rownames_to_column("barcode")
aux_table<-temp%>%dplyr::select(barcode,myeloid_cl)%>%
    right_join(aux_table)%>%
    mutate(nosub=ifelse(!is.na(myeloid_cl),
                        myeloid_cl,nosub))%>%
    dplyr::select(-myeloid_cl)
aux_table<-aux_table[order(match(aux_table$barcode,colnames(sc))),]
sc<-AddMetaData(sc,metadata = aux_table,col.name = "nosub")
```

## prep for NicheNet
```{r}
#gsea_bp<-read_csv("docs/has_laDEG-nosubbulk-GSEA-BPsimpl.csv")
#gsea_bp<-gsea_bp%>%separate_rows(core_enrichment,sep = '/')%>%
#  mutate(gene=core_enrichment)%>%distinct(ct,gene)
deg<-data.table::fread("docs/hasLA_DEG_sc_newMyeloid_MAST.gz")
degs.clusters.h.vs.s<-deg%>%filter(abs(avg_log2FC)>log(1.5)&p_val_adj<0.05)%>% 
  mutate(ct=str_replace_all(ct," |-|\\/","_"))%>%
  arrange(desc(p_val_adj))%>%distinct(gene,ct)%>%
  group_by(ct)%>%mutate(index=1:n())%>%
  pivot_wider(id_cols = index,names_from = ct,values_from = gene)%>%
  dplyr::select(-index)

cell_IDs<-sc@meta.data%>%split(f = sc@meta.data$nosub)%>%
  lapply(function(x) rownames(x))
names(cell_IDs)<-str_replace_all(names(cell_IDs)," |-|\\/","_")
bg <- background_genes_NicheNet(data = as.matrix(sc@assays$RNA$counts),
                                cell_IDs = cell_IDs)
#[1] "n genes expressed 10781"
#[1] "n genes expressed 11917"
#[1] "n genes expressed 9129"
#[1] "n genes expressed 9574"
#[1] "n genes expressed 12065"
#[1] "n genes expressed 14457"
#[1] "n genes expressed 12149"
#[1] "n genes expressed 9434"
#[1] "n genes expressed 9807"
#[1] "n genes expressed 10681"
#[1] "n genes expressed 11532"
#[1] "n genes expressed 8979"
#[1] "n genes expressed 10859"
#[1] "n genes expressed 12983"
#[1] "n genes expressed 14006"
#[1] "n genes expressed 10592"
#[1] "n genes expressed 11295"
#[1] "n genes expressed 9362"
#[1] "n genes expressed 12260"
#[1] "n genes expressed 8514"
#[1] "n genes expressed 9730"
bg<-bg[rowSums(is.na(bg))<ncol(bg),]
#write_csv(as.data.frame(bg),"docs/scDrug_bg.temp")

all_ligand_activity <- NicheNet_ligand_activity_analysis(degs = as.matrix(degs.clusters.h.vs.s),
                                                         background_genes = bg, 
                                                         out_dir = "docs/",
                                                         cores = 8)
#function ran but failed to re-load output files
files<-list.files("docs",full.names = T)
files<-files[str_detect(files,"ligand_act")]
files<-files[str_detect(files,"vity.txt",negate = T)]
all_ligand_activity<-lapply(files,read_tsv)
all_ligand_activity<-lapply(1:length(files), 
                            function(x) all_ligand_activity[[x]]%>%
                              mutate(interaction=str_remove_all(files[x],".+vity_|.txt")))%>%
  bind_rows()
all_ligand_activity<-all_ligand_activity%>%
  separate(interaction,into = c("Sender","Target"),sep = "_to_")
```
## Centrality
```{r}
intercellular_centrality <- NicheNet_cell_type_centrality(all_ligand_activity, out_dir = "docs")

intra_cent <- intracellular_drug_centrality(ppin = ppin,
                                     drug_target_matrix = drug_target_matrix,
                                     degs =as.matrix(degs.clusters.h.vs.s),
                                     file_name = "intracellular_centrality",
                                     centrality_alg = "eigenvector centralities",
                                     out_dir = "docs")
#[1] "The following clusters / cell types have no LCC: Plasmablast_Plasma_cells"

# Only NK, Th17 and MDSCs have centrality=0
```

## Network distance calculation
```{r}
for(i in 1:ncol(degs.clusters.h.vs.s)){ # For every cell types DEGs
  # calculate average closest distances between all drugs and the degs
  # if possible on your machine this should be parallelized (cores > 1), though the function will be memory heavy
  average_closest_distance_network_drug_screening(ppin = ppin,
                                               drug_target_matrix = drug_target_matrix,
                                               disease_genes = degs.clusters.h.vs.s[,i],
                                               file_name = colnames(degs.clusters.h.vs.s)[i],
                                               cores = 8,
                                               out_dir = "docs")
}
lf  <- list.files(path = "docs", pattern = "drug-disease_closest_distances_vs_random_bin_adjusted__")
separate_unique_drug_target_combinations_into_individual_drugs(files = lf,
  remove_part_of_file_name_for_output = "drug-disease_closest_distances_vs_random_bin_adjusted__",
  same_drugs = same_drugs,
  disease_specific_drugs = RA_drugs,
  output_file_name_add_on = "INDIVIDUAL_DRUGS_",
  "docs",
  out_dir = "docs")

files <- list.files("docs",
                         pattern = "drug-disease_closest_distance", 
                         full.names = T)
#fc_evaluation <- pharma_effect_on_fold_change(drug_dist_files = drug_dists, deg_files = deg_files, pharma_effect = drug_bank_example_data[,c(1,6:7)], save_names = save_name, out_dir = "Sample_output/FC_criteria_checking")
#manul pharme effect instead
drug_dists<-lapply(files,read_tsv)
drug_dists<-lapply(1:length(files),function(x) 
  drug_dists[[x]]%>%
    mutate(Cluster=str_remove_all(files[x],
                                  ".+__|.txt")))%>%bind_rows()
drug_dists<-drug_bank_example_data%>%
  dplyr::select(drugID,gene_symbol,target_organism)%>%
  right_join(drug_dists,by = c("drugID"="Drug"))
drug_dists<-drug_dists%>%
  mutate(pharmacological_effect=paste0(gene_symbol," (",target_organism,")"))%>%
  dplyr::select(-target_organism)%>%distinct()
temp<-deg%>%filter(abs(avg_log2FC)>log(1.5)&pct.1>=.1)%>%
  mutate(Cluster=ct,gene_symbol=gene,
                   targeted_degs=ifelse(avg_log2FC<0,
                                        paste0(gene," (down)"),
                                        paste0(gene," (up)")))%>%
  select(Cluster,gene_symbol,targeted_degs)
drug_dists<-drug_dists%>%left_join(temp)
drug_dists<-drug_dists%>%group_by(Cluster,drugID)%>%
  mutate(n_targeted_degs=sum(!is.na(targeted_degs)),
         targeted_degs=ifelse(n_targeted_degs==0,
                              "[]",targeted_degs[!is.na(targeted_degs)]))
drug_dists<-drug_dists%>%group_by(Cluster,drugID)%>%
  mutate(targeted_degs=ifelse(n_targeted_degs==0,"[]",
                              paste(unique(targeted_degs),
                                    collapse = ", ")),
         pharmacological_effect=ifelse(n()>1,
                paste(unique(pharmacological_effect),collapse=", "),
                pharmacological_effect))%>%
  select(-gene_symbol)%>%distinct()
```
# Plot
```{r}
library(ggrepel)
i<-drug_dists%>%ungroup()%>%filter(dc<=1&zc<1.64&n_targeted_degs>0)%>%
  separate_rows(targeted_degs,sep = ", ")%>%
  mutate(targeted_degs=str_remove(targeted_degs," .+"))%>%
  distinct(targeted_degs,Cluster)
#pseudobulk expression because 2 cell types can have opposite z scores
pA<-DotPlot(sc,
            features = i$targeted_degs,
            group.by = "has_LA")
i<-pA$data%>%filter(id=="LA+")%>%arrange(avg.exp*sign(avg.exp.scaled))
pA<-pA$data%>%
  mutate(features.plot=factor(features.plot,levels=i$features.plot))%>%
  ggplot(aes(x=id,y=features.plot,col=avg.exp.scaled,size=pct.exp))+
  geom_point()+
  scale_colour_gradient2(low = "#2166ac", mid = "#f7f7f7", high = "#b2182b")+
  scale_size(range=c(1,5),limits = c(0.15,63))+theme_classic()+labs(x="Status",y="",col="z",size="%")

#antibody targets
j<-c(CD25="IL2RA",CD41="ITGA2B",CD11a="ITGAL",CD20="MS4A1","ERBB2","EGFR","CD33","CD52","ITGA4","CD3D")
pB<-DotPlot(sc,
            features = j,
            group.by = "has_LA")
pB<-pB$data%>%mutate(features.plot=ifelse(feature.groups!="",
                                          as.character(feature.groups),
                                          as.character(features.plot)))
j<-pB%>%filter(id=="LA+")%>%arrange(avg.exp*sign(avg.exp.scaled))
pB<-pB%>%mutate(features.plot=factor(features.plot,levels=j$features.plot))%>%
    ggplot(aes(x=id,y=features.plot,col=avg.exp.scaled,size=pct.exp))+
    geom_point()+
    scale_colour_gradient2(low = "#2166ac", mid = "#f7f7f7", high = "#b2182b")+
    scale_size(range=c(1,5),limits = c(0.15,63))+theme_classic()+labs(x="Status",y="",col="z",size="%")

i<-drug_dists%>%ungroup()%>%filter(dc<=1&zc<1.64&n_targeted_degs>0)%>%distinct(drugID)
i<-intra_cent%>%as.data.frame()%>%rownames_to_column("drugID")%>%
  filter(drugID%in%i$drugID)%>%select(drugID,intracellular_centrality_mean)
i<-drug_bank_example_data%>%distinct(drugID,drug_name)%>%inner_join(i)
pC<-i%>%ggplot(aes(x='A',y=intracellular_centrality_mean))+
  geom_boxplot()+geom_text_repel(data=i%>%
                                   filter(intracellular_centrality_mean>0.03),
                                 aes(labels=drug_name))+
  theme_classic()+labs(x="",y="Avg. intracellular centrality")