---
title: "scDrugPrio"
---


```{r}
library(scDrugPrio)# annoying to install
#library(STRINGdb)
library(tidyverse)
#library(biomaRt)
#library(Seurat)
library(igraph)

load_sample_data()#to get drugbank data
```

# get public data
```{r}
drug_bank_example_data<-as.data.frame(drug_bank_example_data)
drug_bank_example_data<-drug_bank_example_data%>%
  filter(drug_action=="Humans"&str_detect(status,"approved"))
drug_bank_example_data<-drug_bank_example_data%>%
  filter(str_detect(status,"withdrawn",negate = T))

#string_db <- STRINGdb$new(
#  version          = "12.0",
#  species          = 9606,
#  score_threshold  = 900,#max confidence
#  input_directory  = "",
#  protocol         = "https")
#string_db
#edges<-get.edgelist(string_db$graph)
#edges<-as.data.frame(edges)
#only 1 result when using this network: insulin-HLADQA2(MDSCs)

ensembl <- useEnsembl(biomart = "genes", dataset = "hsapiens_gene_ensembl")
bm <- getBM(attributes = c('entrezgene_id', 'external_gene_name'), mart = ensembl)
temp<-lit_ppi%>%as.data.frame()%>%
  mutate(Protein_A=as.numeric(Protein_A))%>%
  left_join(bm,by=c("Protein_A"="entrezgene_id"))%>%
    mutate(nAname=external_gene_name)%>%
  dplyr::select(-external_gene_name)
temp<-temp%>%
  mutate(Protein_B=as.numeric(Protein_B))%>%
  left_join(bm,by=c("Protein_B"="entrezgene_id"))%>%
    mutate(nBname=external_gene_name)%>%
  dplyr::select(-external_gene_name)%>%
  drop_na()%>%dplyr::select(nAname:nBname)%>%
  distinct()%>%as.matrix()
ppin<-temp%>%ppin_formatting()
#write_csv(as.data.frame(ppin),file = "docs/edges.temp")


#Select unique drug target combinations found in the PPIN's LCC of
g<-ppin%>%graph_from_data_frame(directed = F)
nodes<-V(g)$name
#edges%>%as.data.frame()%>%write_csv("docs/PPIN.temp")
drug_bank_example_data<-drug_bank_example_data%>%filter(gene_symbol%in%nodes)
drug_target_matrix <- create_drug_target_matrix(drugID = drug_bank_example_data[,1], 
                                                target = drug_bank_example_data[,6])
#sum(colSums(is.na(drug_target_matrix))<nrow(drug_target_matrix))==ncol(drug_target_matrix)
drug_target_matrix <- prepare_drug_target_matrix_for_network_distance_calculation(ppin,
                                                                                  drug_target_matrix, 
                                                                                  file_name = "scdrug_in_ppin", 
                                                                                  out_dir = "docs")
#[1] "n of unique proteins/genes in PPIN: n =  17345"
#[1] "n unique drugs included in analysis: n = 94"
#[1] "n drugs with unique drug target combinations: n = 74"
#[1] "n unique drugs that can be extracted from UNIQUE_DRUG_TARGET_COMBINATION_scdrug_in_ppin.txt using SAME_DRUG_TARGETS_scdrug_in_ppin.txt: n = 94"
#[1] "Drug matrix = FORMATTED"
same_drugs <- read.table(file = "docs/SAME_DRUG_TARGETS_scdrug_in_ppin.txt", sep="\t", header = T, stringsAsFactors = F)

```

# focus on your data
```{r}
sc<-readRDS("data/fixed_annotated_aux.seurat.rds")
sc<-subset(sc,cells = which(sc$Class%in%c("Peritoneal endometriosis","Eutopic endometrium")&
              str_detect(sc$harmonized_major_plus_immune_cell_type,"^Exc|Ery",negate = T)))
gc()
sc$has_LA<-ifelse(sc$LA_per_Sample>0,"LA+","LA-")
sc@meta.data<-sc@meta.data%>%
  mutate(nosub=str_remove(harmonized_major_plus_immune_cell_type," \\(.+"))%>%
  mutate(nosub=str_replace(nosub,"Memmo","Memo"))

temp<-read_csv("data/myeloid_subgroups.csv")
aux_table<-sc@meta.data%>%rownames_to_column("barcode")
aux_table<-temp%>%dplyr::select(barcode,myeloid_cl)%>%
    right_join(aux_table)%>%
    mutate(nosub=ifelse(!is.na(myeloid_cl),
                        myeloid_cl,nosub))%>%
    dplyr::select(-myeloid_cl)
aux_table<-aux_table[order(match(aux_table$barcode,colnames(sc))),]
sc<-AddMetaData(sc,metadata = aux_table,col.name = "nosub")

pos<-subset(sc,cells = which(sc$Class=="Eutopic endometrium"|sc$has_LA=="LA+"))
neg<-subset(sc,cells = which(sc$Class=="Eutopic endometrium"|sc$has_LA=="LA-"))
cts<-unique(pos$nosub)
pos<-AggregateExpression(pos,assays = "RNA",return.seurat = T,
    group.by = c("SampleName","nosub","Class"))
neg<-AggregateExpression(neg,assays = "RNA",return.seurat = T,
    group.by = c("SampleName","nosub","Class"))
pos<-lapply(cts, function(x) subset(pos,subset =nosub==x))
neg<-lapply(cts, function(x) subset(neg,subset =nosub==x))
names(neg)<-names(pos)<-cts
degpos<-lapply(pos,function(x) 
  FindMarkers(x,group.by="Class",ident.1="Peritoneal endometriosis",
              ident.2="Eutopic endometrium",test.use="DESeq2"))
degneg<-lapply(neg,function(x) 
  FindMarkers(x,group.by="Class",ident.1="Peritoneal endometriosis",
              ident.2="Eutopic endometrium",test.use="DESeq2"))

degpos<-lapply(names(degpos),function(x) 
  degpos[[x]]%>%rownames_to_column("gene")%>%mutate(ct=x,status="LA+"))%>%bind_rows()
degneg<-lapply(names(degneg),function(x) 
  degneg[[x]]%>%rownames_to_column("gene")%>%mutate(ct=x,status="LA-"))%>%bind_rows()
bulkdeg<-degpos%>%bind_rows(degneg)
bulkdeg%>%data.table::fwrite("docs/DESEq_perLA_vsEu.gz")
```

## prep for NicheNet
```{r}
degs.pos<-bulkdeg%>%filter(status=="LA+"&abs(avg_log2FC)>log(1.5)&p_val_adj<0.05)%>% 
  mutate(ct=str_replace_all(ct," |-|\\/","_"))%>%
  arrange(desc(p_val_adj))%>%distinct(gene,ct)%>%
  group_by(ct)%>%mutate(index=1:n())%>%
  pivot_wider(id_cols = index,names_from = ct,values_from = gene)%>%
  dplyr::select(-index)
degs.neg<-bulkdeg%>%filter(status=="LA-"&abs(avg_log2FC)>log(1.5)&p_val_adj<0.05)%>% 
  mutate(ct=str_replace_all(ct," |-|\\/","_"))%>%
  arrange(desc(p_val_adj))%>%distinct(gene,ct)%>%
  group_by(ct)%>%mutate(index=1:n())%>%
  pivot_wider(id_cols = index,names_from = ct,values_from = gene)%>%
  dplyr::select(-index)

mat_pos<-do.call(cbind,lapply(pos, function(x) as.matrix(x@assays$RNA$counts)))
mat_neg<-do.call(cbind,lapply(neg, function(x) as.matrix(x@assays$RNA$counts)))

cell_IDs<-lapply(pos,function(x) colnames(x))
names(cell_IDs)<-str_replace_all(names(cell_IDs)," |-|\\/","_")
bg_pos <- background_genes_NicheNet(data = mat_pos,
                                cell_IDs = cell_IDs)
#~20k
bg_pos<-bg_pos[rowSums(is.na(bg_pos))<ncol(bg_pos),]
write_csv(as.data.frame(bg_pos),"docs/pb_perLA/LApos/bg.temp")
cell_IDs<-lapply(neg,function(x) colnames(x))
names(cell_IDs)<-str_replace_all(names(cell_IDs)," |-|\\/","_")
bg_neg <- background_genes_NicheNet(data = mat_neg,
                                cell_IDs = cell_IDs)
#~20k
bg_neg<-bg_neg[rowSums(is.na(bg_neg))<ncol(bg_neg),]
write_csv(as.data.frame(bg_neg),"docs/pb_perLA/LAneg/bg.temp")


all_ligand_pos <- NicheNet_ligand_activity_analysis(degs = as.matrix(degs.pos),
                            background_genes = bg_pos, 
                            out_dir = "docs/pb_perLA/LApos",
                            cores = 8,
                            ligand_target = as.matrix(readRDS("../../Downloads/ligand_target_matrix.rds")),
                            lr_network = as.matrix(readRDS("../../Downloads/lr_network.rds")))
all_ligand_neg <- NicheNet_ligand_activity_analysis(degs = as.matrix(degs.neg),
                            background_genes = bg_neg, 
                            out_dir = "docs/pb_perLA/LAneg",
                            cores = 8,
                            ligand_target = as.matrix(readRDS("../../Downloads/ligand_target_matrix.rds")),
                            lr_network = as.matrix(readRDS("../../Downloads/lr_network.rds")))


#function ran but failed to re-load output files
files<-list.files("docs/pb_perLA/LApos",full.names = T)
files<-files[str_detect(files,"ligand_act")]
#files<-files[str_detect(files,"vity.txt",negate = T)]
all_ligand_pos<-lapply(files,read_tsv)
all_ligand_pos<-lapply(1:length(files), 
                            function(x) all_ligand_pos[[x]]%>%
                              mutate(interaction=str_remove_all(files[x],".+vity_|.txt")))%>%
  bind_rows()
all_ligand_pos<-all_ligand_pos%>%
  separate(interaction,into = c("Sender","Target"),sep = "_to_")

files<-list.files("docs/pb_perLA/LAneg",full.names = T)
files<-files[str_detect(files,"ligand_act")]
#files<-files[str_detect(files,"vity.txt",negate = T)]
all_ligand_neg<-lapply(files,read_tsv)
all_ligand_neg<-lapply(1:length(files), 
                            function(x) all_ligand_neg[[x]]%>%
                              mutate(interaction=str_remove_all(files[x],".+vity_|.txt")))%>%
  bind_rows()
all_ligand_neg<-all_ligand_neg%>%
  separate(interaction,into = c("Sender","Target"),sep = "_to_")
#unique(all_ligand_neg$test_ligand)[!unique(all_ligand_neg$test_ligand)%in%unique(all_ligand_pos$test_ligand)]
# [1] "PTN"     "CCL3"    "THBS1"   "OMG"     "JAG1"    "CD24"    "MIF"     "AREG"    "ADCYAP1" "CCL14"   "SDK2"    "GRN"    
#[13] "SELE"    "DLL4"    "PDGFB"   "PCDH1"   "APOA1"   "CEACAM1" "ESAM"    "NGF"     "THBS2"   "RARRES2" "REN"     "SEMA3F" 
#[25] "SCG2"    "TNFSF14"
#unique(all_ligand_pos$test_ligand)[!unique(all_ligand_pos$test_ligand)%in%unique(all_ligand_neg$test_ligand)]
# [1] "CD200"   "NPNT"    "NTN4"    "SPN"     "FGF14"   "MPZ"     "IL15"    "NDP"     "CDH6"    "EDN3"    "TNFSF10" "CADM4"  
#[13] "CCL8"    "RSPO3"   "CXCL8"   "GDF11"   "FGF16"   "HBEGF"   "WNT11"   "HSPG2"   "CDH13"   "BMP5"
```
## Centrality
```{r}
inter_pos <- NicheNet_cell_type_centrality(all_ligand_pos, out_dir = "docs/pb_perLA/LApos")
intra_pos <- intracellular_drug_centrality(ppin = ppin,
                                     drug_target_matrix = drug_target_matrix,
                                     degs =as.matrix(degs.pos),
                                     file_name = "intracellular_centrality",
                                     centrality_alg = "eigenvector centralities",
                                     out_dir = "docs/pb_perLA/LApos")
#[1] "The following clusters / cell types have too few genes to have a LCC: M1_macrophages, Effector_Memory_CD8_T_cells, Mast_cells, Gamma_Delta_T_cells, M2_macrophages, Resident_Memory_CD8_T_cells, NK_cells"
#[1] "The following clusters / cell types have no LCC: EnEpi_cells, Mesothelial_cells, Smooth_muscle_cells, Effector_CD8_T_cells, DCs"

inter_neg <- NicheNet_cell_type_centrality(all_ligand_pos, out_dir = "docs/pb_perLA/LAneg")
intra_neg <- intracellular_drug_centrality(ppin = ppin,
                                     drug_target_matrix = drug_target_matrix,
                                     degs =as.matrix(degs.neg),
                                     file_name = "intracellular_centrality",
                                     centrality_alg = "eigenvector centralities",
                                     out_dir = "docs/pb_perLA/LAneg")
#[1] "The following clusters / cell types have too few genes to have a LCC: Plasmablast_Plasma_cells, Th17_T_cells, MDSCs"
#[1] "The following clusters / cell types have no LCC: M1_macrophages, Resident_Memory_CD8_T_cells, EnS_cells, Effector_CD8_T_cells, Mast_cells, Endothelial_cells, Gamma_Delta_T_cells, EnEpi_cells, M2_macrophages, Mesothelial_cells, DCs, Effector_Memory_CD8_T_cells"
```

## Network distance calculation
```{r}
for(i in 1:ncol(degs.pos)){ # For every cell types DEGs
  # calculate average closest distances between all drugs and the degs
  # if possible on your machine this should be parallelized (cores > 1), though the function will be memory heavy
  average_closest_distance_network_drug_screening(ppin = ppin,
                                               drug_target_matrix = drug_target_matrix,
                                               disease_genes = degs.pos[,i],
                                               file_name = colnames(degs.pos)[i],
                                               cores = 10,
                                               out_dir = "docs/pb_perLA/LApos")
}
lf  <- list.files(path = "docs/pb_perLA/LApos",
                  pattern = "drug-disease_closest_distances_vs_random_bin_adjusted__")
separate_unique_drug_target_combinations_into_individual_drugs(files = lf,
  remove_part_of_file_name_for_output = "drug-disease_closest_distances_vs_random_bin_adjusted__",
  same_drugs = same_drugs,
  disease_specific_drugs = c("DB00026"),
  output_file_name_add_on = "INDIVIDUAL_DRUGS_",
  "docs/pb_perLA/LApos",
  out_dir = "docs/pb_perLA/LApos")

files <- list.files("docs/pb_perLA/LApos",
                         pattern = "drug-disease_closest_distance", 
                         full.names = T)
#fc_evaluation <- pharma_effect_on_fold_change(drug_dist_files = drug_dists, deg_files = deg_files, pharma_effect = drug_bank_example_data[,c(1,6:7)], save_names = save_name, out_dir = "Sample_output/FC_criteria_checking")
#manul pharme effect instead
drug_dists<-lapply(files,read_tsv)
drug_dists<-lapply(1:length(files),function(x) 
  drug_dists[[x]]%>%
    mutate(Cluster=str_remove_all(files[x],
                                  ".+__|.txt")))%>%bind_rows()
drug_dists<-drug_bank_example_data%>%
  dplyr::select(drugID,gene_symbol,target_organism)%>%
  right_join(drug_dists,by = c("drugID"="Drug"))
drug_dists<-drug_dists%>%
  mutate(pharmacological_effect=paste0(gene_symbol," (",target_organism,")"))%>%
  dplyr::select(-target_organism)%>%distinct()
temp<-bulkdeg%>%filter(status=="LA+"&abs(avg_log2FC)>log(1.5)&p_val_adj<0.05)%>%
  mutate(Cluster=ct,gene_symbol=gene,
                   targeted_degs=ifelse(avg_log2FC<0,
                                        paste0(gene," (down)"),
                                        paste0(gene," (up)")))%>%
  select(Cluster,gene_symbol,targeted_degs)
drug_dists<-drug_dists%>%left_join(temp)
drug_dists<-drug_dists%>%group_by(Cluster,drugID)%>%
  mutate(n_targeted_degs=sum(!is.na(targeted_degs)),
         targeted_degs=ifelse(n_targeted_degs==0,
                              "[]",targeted_degs[!is.na(targeted_degs)]))
drug_dists<-drug_dists%>%group_by(Cluster,drugID)%>%
  mutate(targeted_degs=ifelse(n_targeted_degs==0,"[]",
                              paste(unique(targeted_degs),
                                    collapse = ", ")),
         pharmacological_effect=ifelse(n()>1,
                paste(unique(pharmacological_effect),collapse=", "),
                pharmacological_effect))%>%
  select(-gene_symbol)%>%distinct()

drug_dist_filtered<-drug_dists%>%filter(P<0.05)
drug_rank <- final_drug_prioritization(fc_evaluation = as.matrix(drug_dist_filtered), 
                                       pos_DrugID = 1,
                                       pos_clusterID = 8, 
                                       keep = c(1,2,7:10), #drugID,n_drug_targets,pharmacological_effect
                                       inter_cent = inter_pos[5,], 
                                       intra_cent = intra_pos[,ncol(intra_pos)],
                                       file_name = "FINAL_drug_ranking", 
                                       out_dir = "docs/pb_perLA/LApos")
drug_rank<-drug_bank_example_data%>%distinct(drugID,drug_name)%>%inner_join(as.data.frame(drug_rank))
write_excel_csv(drug_rank,"docs/pb_perLA/LApos/FINAL_drug_ranking.txt")

```
# Plot
```{r}
library(ggVennDiagram)
drug_rank<-drug_rank_pos%>%mutate(from="LA+")%>%
  bind_rows(drug_rank_neg%>%mutate(from="LA-"))
i<-drug_rank%>%distinct(drug_name,from)
i<-i%>%split(f = i$from)%>%lapply(function(x) x$drug_name)
p0<-ggVennDiagram(i)

drug_rank<-drug_rank%>%separate_rows(pharmacological_effect,sep = ", ")%>%
  mutate(gene=str_remove(pharmacological_effect," .+"))
i<-drug_rank%>%distinct(gene)%>%unlist()
temp<-DotPlot(subset(sc,subset=Class!="Eutopic endometrium"),
              features = i,
              group.by = "has_LA")
i<-temp$data%>%group_by(features.plot)%>%
  slice_max(avg.exp.scaled)%>%
  mutate(gene=features.plot)%>%
  ungroup()%>%select(gene,id)
i<-drug_rank%>%inner_join(i)%>%
  distinct(drug_name,pharmacological_effect,id)
i<-i%>%group_by(drug_name)%>%add_count(id,name = "hg_per_group")%>%
  mutate(no_annot=sum(str_detect(pharmacological_effect,"NA")),
         total=n())%>%
  mutate(target_group=ifelse(total-no_annot==1,
        as.character(id[str_detect(pharmacological_effect,"NA",negate = T)]),
        as.character(id[hg_per_group==max(hg_per_group)])))%>%
  distinct(drug_name,target_group)
i<-i%>%add_count(drug_name)%>%
  mutate(target_group=ifelse(n>1,"both",target_group))%>%
  select(-n)%>%distinct()
drug_rank<-drug_rank%>%left_join(i)
drug_rank<-drug_rank%>%
  mutate(target_group=ifelse(str_detect(drug_name,
                             "L-|As|Biot|balam|Coag|Strept"),
                             "likely inactive",target_group))
i<-drug_rank%>%count(from,target_group)
p1<-ggplot(i,aes(axis1 = from, axis2 = target_group, y = n,fill=from)) +
    geom_alluvium(alpha = 0.6) +
    geom_stratum(width = 1/2, color = "grey60") +
    geom_text(stat = "stratum",
               aes(label = after_stat(stratum)),
               size = 4)+
    scale_x_discrete(limits = c("Origin", "Manual annotation"),
                     expand = c(0.1, 0.1)) +
    scale_fill_manual(values=c("#5E4FA2","#8DB600","#BDBABB"))+
    theme_sankey()+theme(legend.position = "n")


drug_rank<-drug_rank%>%group_by(drug_name)%>%
  mutate(pharmacological_effect=paste(pharmacological_effect,collapse = ", "))
drug_rank%>%
  distinct(drug_name,pharmacological_effect,rank,target_group)%>%
  arrange(target_group,rank)%>%view()#manually select among these

#select targets from the top 26 drugs
i<-list("LA+"=c("Alteplase",#Tenecteplase
                "Ibritumomab tiuxetan",#Tositumomab,Rituximab
                "Etanercept"),#Adalimumab
        "LA-"=c("Octreotide","Cetuximab","Abciximab"),
        "both"=c("Collagenase clostridium histolyticum",
                 "Botulinum toxin type A","Ademetionine"))
i<-lapply(1:3,function(x) 
  as.data.frame(cbind(drug_name=i[[x]],group=names(i)[x])))%>%bind_rows()
i<-i%>%inner_join(drug_rank)%>%
    distinct(group,drug_name,pharmacological_effect)%>%
    mutate(pharmacological_effect=str_remove(pharmacological_effect," .+"))%>%
    distinct()
i[1,3]<-"PLAUR"#target with > expr%
i[4,3]<-"SSTR2"#target with > expr%
i[6,3]<-"ITGA2B"#target with > expr%
i<-temp$data%>%
    inner_join(i,by=c("features.plot"="pharmacological_effect"))

p2<-i%>%mutate(dt=paste0(drug_name," - ",features.plot))%>%
  ggplot(aes(x=id,y=dt,col=avg.exp.scaled,size=pct.exp))+geom_point()+
  facet_wrap(~group,scales = "free_y",space="free_y")+
  scale_colour_gradient2(low = "#2166ac", mid = "#f7f7f7", high = "#b2182b")+
  theme_classic()+theme(strip.text.y = element_text(angle=0,hjust = 0),
                          strip.background = element_blank(),
                          panel.spacing = unit(1, "mm"))+
  labs(x="Status",y="Drug - target",col="z",size='%')