---
title: "Cell Communcaiton "
author: "Pak Yu"
date: "2/10/2023"
editor_options: 
  chunk_output_type: console
output: html_document
---

# Loading cell_chat_data library

```{r}
library(cell_chat_data)
library(patchwork)
library(cowplot)
library(grid)
library(NMF)
library(ggalluvial)
library(ComplexHeatmap)
```

#Loading Seruat Obj to cell_chat_data

cell and cell interactions < 10 are filtered

```{r}
cell_chat_data <- readRDS("data/CellChat_endo.rds")
#cell_chat_data <- aggregateNet(cell_chat_data)

```

#Aggregated cell-cell communication network

```{r}
group_size <- cell_chat_data@idents %>% table %>% as.numeric()
par(mfrow = c(1,2), xpd=TRUE)
netVisual_circle(cell_chat_data@net$count, vertex.weight = group_size, weight.scale = T, label.edge= F, title.name = "Number of interactions")
netVisual_circle(cell_chat_data@net$weight, vertex.weight = group_size, weight.scale = T, label.edge= F, title.name = "Interacion weights/strength")


```



```{r}
par(mfrow = c(3,4), xpd=TRUE)
tbl <- cell_chat_data@net$weight
cell_group_interaction_plot <- lapply(1:nrow(tbl),function(i){
    max_weight <- max(tbl)
    print(rownames(tbl)[i])
    tbl[-i,] <- 0
    netVisual_circle(tbl, vertex.weight = group_size, weight.scale = T, edge.weight.max =max_weight , title.name = rownames(tbl)[i])
})
rm(tbl)
```

cowplo::plot_grid doesn't play nice with  recordedplot
```{r}
par(mfrow = c(3,4), xpd=TRUE)

```

#Singal role analysis - need specific pathway

```{r}
cell_chat_data <- netAnalysis_computeCentrality(cell_chat_data, slot.name = "netP")
netAnalysis_signalingRole_network(cell_chat_data, signaling = pathways.show, width = 8, height = 2.5, font.size = 10)
```

```{r}
ht1 <- netAnalysis_signalingRole_heatmap(cell_chat_data, pattern = "outgoing") 
ht2 <- netAnalysis_signalingRole_heatmap(cell_chat_data, pattern = "incoming")
#  grid.grabExpr(.)
draw(ht1+ht2,heatmap_height=unit(30,"cm"))
```



#global communication patterns

## outgoing 
3 retains the best  
```{r,include=FALSE}
selectK(cell_chat_data, pattern = "outgoing")
```

Heatmap color error - disable for now

```{r}
nPatterns = 3
cell_chat_data <- identifyCommunicationPatterns(cell_chat_data, pattern = "outgoing", k = nPatterns, heatmap.show = FALSE)
netAnalysis_river(cell_chat_data, pattern = "outgoing")
outgoing_dot <- netAnalysis_dot(cell_chat_data, pattern = "outgoing") + coord_flip()
```

##incoming

```{r, include=FALSE}
selectK(cell_chat_data, pattern = "incoming")
```

##outgoing

```{r}
nPatterns = 3
cell_chat_data <- identifyCommunicationPatterns(cell_chat_data, pattern = "incoming", k = nPatterns, heatmap.show = FALSE)
netAnalysis_river(cell_chat_data, pattern = "incoming")
incoming_dot <-netAnalysis_dot(cell_chat_data, pattern = "incoming") + coord_flip()
```

```{r}
plot_grid(incoming_dot,outgoing_dot)
```


#Manifold and classification learning analysis of signaling networks

```{r}
cell_chat_data <- computeNetSimilarity(cell_chat_data, type = "functional")
cell_chat_data <- netEmbedding(cell_chat_data, type = "functional")
cell_chat_data <- netClustering(cell_chat_data, type = "functional")
netVisual_embedding(cell_chat_data, type = "functional", label.size = 3.5)
```

# Exporting cell-cell communication 

```{r}
subsetCommunication(cell_chat_data) %>% write_csv("data/cell_chat_db_communication.csv")
subsetCommunication(cell_chat_data,slot.name = "netP") %>% write_csv("data/cell_chat_db_path_communication.csv")
```

```{r}
pattern_show <- "BAFF"
netAnalysis_signalingRole_scatter(cell_chat_data,signaling = pattern_show)
```


##Sanity Check 
```{r, includes=FALSE}
library(Seurat)
library(SeuratDisk)
endo_cluster_loom <-  Connect(filename = "data/aux_seurat.loom", mode ="r")
#BAFF_gene <- c("TNFRSF13B", "TNFRSF13C" , "TNFRSF17"))
endo_cluster <- endo_cluster_loom %>% as.Seurat()
endo_cluster_loom$close_all()

```



